### 2020-08-11 r-package/tests_marcus/rjava_tests.Rmd
Adding functionality to r5r_core: building R5 network.dat file

---
title: "Package Tests"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("rJava")
```

## Testing Java / R Integration

```{r}
.jinit()
.jaddClassPath(path = "../../JAR/R5.jar")
.jaddClassPath(path = "../../JAR/r5r_core.jar")
.jclassPath()
```

```{r}
r5r_core <- .jnew("R5RCore", here::here("inst/extdata/"))
```

```{r}
r5r_core
```

### 2020-08-13 r-package/tests_marcus/rjava_tests.Rmd
Added loadDestinationPointsFromCsv method to r5r api, and succesfully called it from R
Added single trip planning to r5r_core Tested single trip planning from R using .jcall

@@
 library("rJava")
+library("jdx")
+library("sf")
+library("dplyr")
+library("ggplot2")
+```
+
+```{r}
+poa_boundary <- st_read(here::here("inst/extdata/portoalegre_boundary.geojson"))
 ```
@@
 .jaddClassPath(path = "../../JAR/R5.jar")
-.jaddClassPath(path = "../../JAR/r5r_core.jar")
+.jaddClassPath(path = "../../java-api/out/artifacts/r5r_core/r5r_core.jar")
 .jclassPath()
@@
 ```{r}
-r5r_core
+.jcall(r5r_core, returnSig = "V", 
+       method = "loadDestinationPointsFromCsv", here::here("inst/extdata/", "poa_hexgrid.csv"))
 ```
+
+Estação Farrapos
+Lat: -29.997611
+Lon: -51.197720
+
+Shopping Praia de Belas
+Lat: -30.048951
+Lon: -51.229533
+
+Function Signature
+public void planSingleTrip(double fromLat, double fromLon, double toLat, double toLon,
+                               String[] directModes, String[] transitModes, String date, String departureTime) {
+
+```{r}
+fromLat <- -29.997611
+fromLon <- -51.197720
+toLat <- -30.048951
+toLon <- -51.229533
+trip_date <- "2019-04-08"
+departure_time <- "14:00:00"
+
+.jcall(r5r_core, returnSig = "V", 
+       method = "planSingleTrip", fromLat, fromLon, toLat, toLon, "WALK", "BUS", trip_date, departure_time)
+```
+
+```{r}
+path_options <- r5r_core$getPathOptionsTable()
+path_options_df <-convertToR(path_options)
+```
+
+```{r}
+path_options_df %>% 
+  filter(option <= 4) %>%
+  mutate(geometry = st_as_sfc(geometry)) %>%
+  st_sf(crs = 4326) %>%
+  ggplot() +
+  geom_sf(data=poa_boundary) +
+  geom_sf(aes(geometry=geometry, colour = mode)) +
+  scale_colour_brewer(palette = "Set1") +
+  facet_wrap(~option, ncol = 4) +
+  theme(legend.position = "bottom")
+```

### 2020-08-13 java-api/src/R5RCore.java

@@ -1,15 +1,31 @@
 import com.conveyal.r5.analyst.FreeFormPointSet;
+import com.conveyal.r5.analyst.cluster.AnalysisTask;
+import com.conveyal.r5.analyst.cluster.RegionalTask;
+import com.conveyal.r5.api.ProfileResponse;
+import com.conveyal.r5.api.util.*;
 import com.conveyal.r5.kryo.KryoNetworkSerializer;
+import com.conveyal.r5.point_to_point.builder.PointToPointQuery;
 import com.conveyal.r5.transit.TransportNetwork;
+import com.conveyal.r5.transit.TripPattern;
 import com.conveyal.r5.util.CSVInputStreamProvider;
+import org.locationtech.jts.geom.Coordinate;
 
 import java.io.File;
 import java.io.IOException;
+import java.text.DateFormat;
+import java.text.ParseException;
+import java.text.SimpleDateFormat;
+import java.time.LocalDate;
+import java.time.LocalTime;
+import java.util.*;
+
+import static com.conveyal.r5.streets.VertexStore.FIXED_FACTOR;
 
 public class R5RCore {
 
     private TransportNetwork transportNetwork;
     private FreeFormPointSet destinationPointSet;
+    private LinkedHashMap<String, Object> pathOptionsTable;
 
     public R5RCore(String dataFolder) {
         File file = new File(dataFolder + "network.dat");
@@ -42,4 +58,168 @@ public class R5RCore {
         }
     }
 
+    public void planMultipleTrips(double[] fromLat, double[] fromLon, double[] toLat, double[] toLon,
+                                  String directModes, String transitModes, String date, String departureTime) {
+
+    }
+
+    public void planSingleTrip(double fromLat, double fromLon, double toLat, double toLon,
+                               String directModes, String transitModes, String date, String departureTime) throws ParseException {
+        AnalysisTask request = new RegionalTask();
+        request.zoneId = transportNetwork.getTimeZone();
+        request.fromLat = fromLat;
+        request.fromLon = fromLon;
+        request.toLat = toLat;
+        request.toLon = toLon;
+        request.streetTime = 60;
+        request.computePaths = true;
+        request.computeTravelTimeBreakdown = true;
+        request.directModes = EnumSet.of(LegMode.WALK, LegMode.BICYCLE, LegMode.CAR);
+
+        request.transitModes = EnumSet.of(TransitModes.BUS);
[…]

### 2020-08-13 r-package/tests_marcus/rjava_tests.Rmd

--- 2020-08-11 r-package/tests_marcus/rjava_tests.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
-library("rJava")
+library("sf")
+library("tidyverse")
+```
+
+```{r}
+poa_boundary <- st_read("/Users/marcussaraiva/Repos/data_r5r/portoalegre_boundary.geojson")
 ```
@@
 ```{r}
-.jinit()
-.jaddClassPath(path = "../../JAR/R5.jar")
-.jaddClassPath(path = "../../JAR/r5r_core.jar")
-.jclassPath()
+r5_path <- "/Users/marcussaraiva/Repos/r5r/JAR/"
+data_path <- "/Users/marcussaraiva/Repos/data_r5r/"
+
+r5r_core <- r5_setup(r5_path, data_path)
 ```
@@
 ```{r}
-r5r_core <- .jnew("R5RCore", here::here("inst/extdata/"))
+# .jcall(r5r_core, returnSig = "V", 
+       # method = "loadDestinationPointsFromCsv", here::here("inst/extdata/", "poa_hexgrid.csv"))
+```
+
+Estação Farrapos
+Lat: -29.997611
+Lon: -51.197720
+
+Shopping Praia de Belas
+Lat: -30.048951
+Lon: -51.229533
+
+Function Signature
+public void planSingleTrip(double fromLat, double fromLon, double toLat, double toLon,
+                               String[] directModes, String[] transitModes, String date, String departureTime) {
+
+```{r}
+fromLat <- -29.997611
+fromLon <- -51.197720
+toLat <- -30.048951
+toLon <- -51.229533
+trip_date <- "2019-05-20"
+departure_time <- "14:00:00"
+street_time = 15L
+
+paths_bus_df <- r5_plan_trip(r5r_core, fromLat, fromLon, toLat, toLon, "WALK", "BUS", 
+                        trip_date, departure_time, max_street_time = 30L)
+paths_active_df <- r5_plan_trip(r5r_core, fromLat, fromLon, toLat, toLon, c("WALK", "BICYCLE"), "", 
+                           trip_date, departure_time, max_street_time = 45L)
+
+paths_car_df <- r5_plan_trip(r5r_core, fromLat, fromLon, toLat, toLon, c("CAR"), "", 
+                           trip_date, departure_time, max_street_time = 45L)
 ```
@@
 ```{r}
-r5r_core
+paths_active_df
+paths_car_df
 ```
+
+```{r}
+paths_bus_df %>%
+  ggplot() +
+  # geom_sf(data=poa_boundary) +
+  geom_sf(aes(geometry=geometry, colour = mode, linetype=factor(segment))) +
+  scale_colour_brewer(palette = "Set1") +
+  facet_wrap(~option, ncol = 4) +
+  theme(legend.position = "bottom")
+```
+
+## Teste Bike
+
+Porto Alegre, rota ao longo da Av. Ramiro Barcellos.
+Rota mais direta possui alta declividade, com alternativas mais longas porém mais planas.
+
+```{r}
+fromLat <- -30.040177
+fromLon <- -51.208234
+toLat <- -30.022159
+toLon <- -51.212495
+
+paths_bike_df <- r5_plan_trip(r5r_core, fromLat, fromLon, toLat, toLon, c("BICYCLE"), "", 
+                           trip_date, departure_time, max_street_time = 45L)
+```
+
+```{r}
+paths_bike_df %>%
+  ggplot() + 
+  # geom_sf(data=poa_boundary) + 
+  geom_sf()
+```
+
+### Travel Times
+
+Shopping Praia de Belas
+Lat: -30.048951
+Lon: -51.229533
+
+```{r}
+fromId = "praia_de_belas"
+fromLat <- -30.048951
+fromLon <- -51.229533
+trip_date <- "2019-05-20"
+departure_time <- "14:00:00"
+street_time <- 30L
+trip_time <- 120L
+
+destinations_df <- read_csv(paste0(data_path, "poa_hexgrid.csv"))
+
+travel_times <- r5_travel_times_from_origin(r5r_core, fromId, fromLat, fromLon, 
+                                            destinations = destinations_df, 
+                                            direct_modes = "WALK", transit_modes = "BUS", 
+                                            trip_date, departure_time, street_time, trip_time)
+```
+
+```{r}
+travel_times
+```

### 2020-08-14 r-package/tests_marcus/rjava_tests.Rmd
Implemented travel time matrix calculation

--- 2020-08-11 r-package/tests_marcus/rjava_tests.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
-library("rJava")
-```
-
-## Testing Java / R Integration
-
-```{r}
-.jinit()
-.jaddClassPath(path = "../../JAR/R5.jar")
-.jaddClassPath(path = "../../JAR/r5r_core.jar")
-.jclassPath()
+library("sf")
+library("tidyverse")
 ```
@@
 ```{r}
-r5r_core <- .jnew("R5RCore", here::here("inst/extdata/"))
+poa_boundary <- st_read("/Users/marcussaraiva/Repos/data_r5r/portoalegre_boundary.geojson")
+```
+
+## Loading Java Libraries
+
+```{r include=FALSE}
+r5_path <- "/Users/marcussaraiva/Repos/r5r/JAR/"
+data_path <- "/Users/marcussaraiva/Repos/data_r5r/"
+
+r5r_core <- r5_setup(r5_path, data_path)
+```
+
+## Querying detailed itineraries
+
+Estação Farrapos
+Lat: -29.997611
+Lon: -51.197720
+
+Shopping Praia de Belas
+Lat: -30.048951
+Lon: -51.229533
+
+Function Signature
+public void planSingleTrip(double fromLat, double fromLon, double toLat, double toLon,
+                               String[] directModes, String[] transitModes, String date, String departureTime) {
+
+```{r include=FALSE}
+fromLat <- -29.997611
+fromLon <- -51.197720
+toLat <- -30.048951
+toLon <- -51.229533
+trip_date <- "2019-05-20"
+departure_time <- "14:00:00"
+street_time = 15L
+
+paths_bus_df <- r5_plan_single_trip(r5r_core, fromLat, fromLon, toLat, toLon, c("WALK", "BICYCLE", "CAR"), "BUS", 
+                        trip_date, departure_time, max_street_time = 30L)
 ```
@@
 ```{r}
-r5r_core
+paths_bus_df
 ```
+
+```{r}
+paths_bus_df %>%
+  ggplot() +
+  # geom_sf(data=poa_boundary) +
+  geom_sf(aes(geometry=geometry, colour = factor(segment), linetype=mode)) +
+  scale_colour_brewer(palette = "Set1") +
+  facet_wrap(~option, ncol = 5) +
+  theme(legend.position = "bottom")
+```
+
+## Teste Bike
+
+Porto Alegre, rota ao longo da Av. Ramiro Barcellos.
+Rota mais direta possui alta declividade, com alternativas mais longas porém mais planas.
+
+```{r}
+fromLat <- -30.040177
+fromLon <- -51.208234
+toLat <- -30.022159
+toLon <- -51.212495
+
+paths_bike_df <- r5_plan_single_trip(r5r_core, fromLat, fromLon, toLat, toLon, c("BICYCLE"), "", 
+                                     trip_date, departure_time, max_street_time = 45L)
+```
+
+```{r}
+paths_bike_df %>%
+  ggplot() + 
+  geom_sf(data=poa_boundary) +
+  geom_sf()
+```
+
+### Travel Times from Single Origin
+
+Shopping Praia de Belas
+Lat: -30.048951
+Lon: -51.229533
+
+```{r include=FALSE}
+fromId = "praia_de_belas"
+fromLat <- -30.048951
+fromLon <- -51.229533
+trip_date <- "2019-05-20"
+departure_time <- "14:00:00"
+street_time <- 30L
+trip_time <- 120L
+
+destinations_df <- read_csv(paste0(data_path, "poa_hexgrid.csv"))
+
+travel_times <- r5_travel_times_from_origin(r5r_core, fromId, fromLat, fromLon, 
+                                            destinations = destinations_df, 
+                                            direct_modes = "WALK", transit_modes = "BUS", 
+                                            trip_date, departure_time, street_time, trip_time)
+```
+
+```{r}
+travel_times
+```
+
+### Travel Time Matrix
+
+```{r include=FALSE}
+trip_date <- "2019-05-20"
+departure_time <- "14:00:00"
+street_time <- 30L
+trip_time <- 120L
+
+origins_df <- read_csv(paste0(data_path, "poa_hexgrid.csv"))  %>% sample_n(20)
+destinations_df <- read_csv(paste0(data_path, "poa_hexgrid.csv")) %>% sample_n(100)
+
+travel_times <- r5_travel_time_matrix(r5r_core, origins = origins_df, destinations = destinations_df, 
+                                      direct_modes = "WALK", transit_modes = "BUS", 
+                                      trip_date, departure_time, street_time, trip_time)
+
+```
+
+```{r}
+travel_times
+```

### 2020-08-14 java-api/src/R5RCore.java

@@ -10,24 +10,21 @@ import com.conveyal.r5.kryo.KryoNetworkSerializer;
 import com.conveyal.r5.point_to_point.builder.PointToPointQuery;
 import com.conveyal.r5.transit.TransportNetwork;
 import com.conveyal.r5.transit.TripPattern;
-import com.conveyal.r5.util.CSVInputStreamProvider;
-import org.locationtech.jts.geom.Coordinate;
-
 import java.io.File;
 import java.io.IOException;
 import java.text.DateFormat;
 import java.text.ParseException;
 import java.text.SimpleDateFormat;
 import java.time.LocalDate;
-import java.time.LocalTime;
 import java.util.*;
+import java.util.stream.Collectors;
 
 import static com.conveyal.r5.streets.VertexStore.FIXED_FACTOR;
 
 public class R5RCore {
 
     private TransportNetwork transportNetwork;
-    private LinkedHashMap<String, Object> pathOptionsTable;
+//    private LinkedHashMap<String, Object> pathOptionsTable;
 
     public R5RCore(String dataFolder) {
         File file = new File(dataFolder + "network.dat");
@@ -65,7 +62,7 @@ public class R5RCore {
 
     }
 
-    public void planSingleTrip(double fromLat, double fromLon, double toLat, double toLon,
+    public LinkedHashMap<String, Object> planSingleTrip(double fromLat, double fromLon, double toLat, double toLon,
                                String directModes, String transitModes, String date, String departureTime,
                                int maxStreetTime) throws ParseException {
         AnalysisTask request = new RegionalTask();
@@ -112,7 +109,9 @@ public class R5RCore {
         ProfileResponse response = query.getPlan(request);
 
         if (!response.getOptions().isEmpty()) {
-            buildPathOptionsTable(response.getOptions());
+            return buildPathOptionsTable(response.getOptions());
+        } else {
+            return null;
         }
     }
 
@@ -123,7 +122,7 @@ public class R5RCore {
         return (int) ((date.getTime() - reference.getTime()) / 1000L);
     }
 
-    private void buildPathOptionsTable(List<ProfileOption> pathOptions) {
+    private LinkedHashMap<String, Object> buildPathOptionsTable(List<ProfileOption> pathOptions) {
[…]

### 2020-08-14 r-package/R/r5_plan_single_trip.R

@@ -26,11 +26,14 @@ r5_plan_single_trip <- function(r5r_core, fromLat, fromLon, toLat, toLon, direct
   transit_modes <- paste0(transit_modes, collapse = ";")
 
   # Call to method inside R5RCore object
-  rJava::.jcall(r5r_core, returnSig = "V", method = "planSingleTrip",
-         fromLat, fromLon, toLat, toLon, direct_modes, transit_modes, trip_date, departure_time, max_street_time)
+  path_options <- r5r_core$planSingleTrip(fromLat, fromLon, toLat, toLon,
+                                             direct_modes, transit_modes,
+                                             trip_date, departure_time, max_street_time)
+    # rJava::.jcall(r5r_core, returnSig = "O", method = "planSingleTrip",
+    #               fromLat, fromLon, toLat, toLon, direct_modes, transit_modes, trip_date, departure_time, max_street_time)
 
   # Collects results from R5 and transforms them into simple features objects
-  path_options_df <- jdx::convertToR(r5r_core$getPathOptionsTable()) %>%
+  path_options_df <- jdx::convertToR(path_options) %>%
     mutate(geometry = st_as_sfc(geometry)) %>%
     st_sf(crs = 4326) # WGS 84
 

### 2020-08-14 r-package/tests_marcus/rjava_tests.Rmd
Included R5RCore class in R5 main repository. Now, a single .jar file can be generated. Updated r5_setup function to access a single .jar file.

@@
 ```{r include=FALSE}
+# Caminho do arquivo JAR
 r5_path <- "/Users/marcussaraiva/Repos/r5r/JAR/"
+
+#Caminho da pasta com os dados: .pbf e gtfs, ou network.dat
 data_path <- "/Users/marcussaraiva/Repos/data_r5r/"

### 2020-08-14 java-r5/src/main/java/com/conveyal/r5/R5RCore.java

@@ -1,0 +1,390 @@
+package com.conveyal.r5;
+
+import com.conveyal.r5.OneOriginResult;
+import com.conveyal.r5.analyst.FreeFormPointSet;
+import com.conveyal.r5.analyst.TravelTimeComputer;
+import com.conveyal.r5.analyst.cluster.AnalysisTask;
+import com.conveyal.r5.analyst.cluster.RegionalTask;
+import com.conveyal.r5.analyst.scenario.Scenario;
+import com.conveyal.r5.api.ProfileResponse;
+import com.conveyal.r5.api.util.*;
+import com.conveyal.r5.kryo.KryoNetworkSerializer;
+import com.conveyal.r5.point_to_point.builder.PointToPointQuery;
+import com.conveyal.r5.transit.TransportNetwork;
+import com.conveyal.r5.transit.TripPattern;
+import java.io.File;
+import java.io.IOException;
+import java.text.DateFormat;
+import java.text.ParseException;
+import java.text.SimpleDateFormat;
+import java.time.LocalDate;
+import java.util.*;
+import java.util.stream.Collectors;
+
+import static com.conveyal.r5.streets.VertexStore.FIXED_FACTOR;
+
+public class R5RCore {
+
+    private TransportNetwork transportNetwork;
+//    private LinkedHashMap<String, Object> pathOptionsTable;
+
+    public R5RCore(String dataFolder) {
+        File file = new File(dataFolder + "network.dat");
+        if (!file.isFile()) {
+            // network.dat file does not exist. create!
+            transportNetwork = TransportNetwork.fromDirectory(new File(dataFolder));
+            try {
+                KryoNetworkSerializer.write(transportNetwork, new File(dataFolder, "network.dat"));
+            } catch (IOException e) {
+                e.printStackTrace();
+            }
+        }
+
+        try {
+            transportNetwork = KryoNetworkSerializer.read(new File(dataFolder, "network.dat"));
+//            transportNetwork.readOSM(new File(dir, "osm.mapdb"));
+            transportNetwork.transitLayer.buildDistanceTables(null);
+        } catch (Exception e) {
+            e.printStackTrace();
+        }
+    }
+
+//    public void loadDestinationPointsFromCsv(String csvFile) {
+//        CSVInputStreamProvider csvInputStream;
+//        try {
+//            csvInputStream = new CSVInputStreamProvider(csvFile);
+//            destinationPointSet = FreeFormPointSet.fromCsv(csvInputStream, "lat", "lon", "id", "count");
[…]

### 2020-08-14 r-package/tests_marcus/rjava_tests.Rmd
Added @export tag to matrix calculation functions

@@
 
-origins_df <- read_csv(paste0(data_path, "poa_hexgrid.csv"))  %>% sample_n(20)
+origins_df <- read_csv(paste0(data_path, "poa_hexgrid.csv"))  %>% sample_n(100)
 destinations_df <- read_csv(paste0(data_path, "poa_hexgrid.csv")) %>% sample_n(100)
@@
 travel_times <- r5_travel_time_matrix(r5r_core, origins = origins_df, destinations = destinations_df, 
-                                      direct_modes = "WALK", transit_modes = "BUS", 
+                                      direct_modes = c("WALK"), transit_modes = "", 
                                       trip_date, departure_time, street_time, trip_time)

### 2020-08-14 java-r5/src/main/java/com/conveyal/r5/R5RCore.java

@@ -186,7 +186,7 @@ public class R5RCore {
                         segmentIndex++;
                         TripPattern tripPattern = transportNetwork.transitLayer.tripPatterns.get(pattern.patternIdx);
 
-                        String geometry = "NA";
+                        StringBuilder geometry = new StringBuilder("NA");
 
                         for (int stop = pattern.fromIndex; stop <= pattern.toIndex; stop++) {
                             int stopIdx = tripPattern.stops[stop];
@@ -194,20 +194,20 @@ public class R5RCore {
                             coord.x = coord.x / FIXED_FACTOR;
                             coord.y = coord.y / FIXED_FACTOR;
 
-                            if (geometry.equals("NA")) {
-                                geometry = "LINESTRING (" + coord.x + " " + coord.y;
+                            if (geometry.toString().equals("NA")) {
+                                geometry = new StringBuilder("LINESTRING (" + coord.x + " " + coord.y);
                             } else {
-                                geometry = geometry + ", " + coord.x + " " + coord.y;
+                                geometry.append(", ").append(coord.x).append(" ").append(coord.y);
                             }
                         }
-                        geometry = geometry + ")";
+                        geometry.append(")");
 
                         optionCol.add(optionIndex);
                         segmentCol.add(segmentIndex);
                         modeCol.add(transit.mode.toString());
                         durationCol.add(transit.rideStats.avg);
                         routeCol.add(tripPattern.routeId);
-                        geometryCol.add(geometry);
+                        geometryCol.add(geometry.toString());
                     }
 
                     if (transit.middle != null) {
@@ -302,18 +302,10 @@ public class R5RCore {
         request.computeTravelTimeBreakdown = false;
         request.recordTimes = true;
 
-        request.directModes = EnumSet.noneOf(LegMode.class);
-        String[] modes = directModes.split(";");
-        if (!directModes.equals("") & modes.length > 0) {
-            for (String mode : modes) {
-                request.directModes.add(LegMode.valueOf(mode));
-            }
-        }
-
         request.transitModes = EnumSet.noneOf(TransitModes.class);
         request.accessModes = EnumSet.noneOf(LegMode.class);
         request.egressModes = EnumSet.noneOf(LegMode.class);
-        modes = transitModes.split(";");
+        String[] modes = transitModes.split(";");
         if (!transitModes.equals("") & modes.length > 0) {
             request.accessModes.add(LegMode.WALK);
[…]

### 2020-08-14 r-package/R/r5_travel_times_from_origin.R

@@ -1,3 +1,21 @@
+#' Title
+#'
+#' @param r5r_core
+#' @param fromId
+#' @param fromLat
+#' @param fromLon
+#' @param destinations
+#' @param direct_modes
+#' @param transit_modes
+#' @param trip_date
+#' @param departure_time
+#' @param max_street_time
+#' @param max_trip_duration
+#'
+#' @return
+#' @export
+#'
+#' @examples
 r5_travel_times_from_origin <- function(r5r_core, fromId, fromLat, fromLon, destinations,
                                         direct_modes, transit_modes, trip_date, departure_time,
                                         max_street_time, max_trip_duration) {

### 2020-08-15 r-package/tests_marcus/rjava_tests.Rmd
Java is quite picky regarding data types, while R is more permissive. Every parameter passed that Java expects to be an integer value shall be forcefully cast to integer using as.integer() before passing to rJava. I have updated the detailed_itineraries() function to do that.

@@
 ```
+
+## Rafa Tests
+
+```{r}
+library(r5r)
+ library(data.table)
+ library(dplyr)
+ library(sf)
+
+ # build transport network
+ path <- system.file("extdata", package = "r5r")
+ r5_core <- setup_r5(data_path = path)
+
+ # load origin/destination points
+ points <- read.csv(system.file("extdata/poa_hexgrid.csv", package = "r5r"))
+
+ # input
+ fromLat <- points[1,]$lat
+ fromLon <- points[1,]$lon
+ toLat <- points[100,]$lat
+ toLon <- points[100,]$lon
+ trip_date <- "2019-03-17"
+ departure_time <- "14:00:00"
+ street_time = 15L
+ direct_modes <- c("WALK", "BICYCLE", "CAR")
+ transit_modes <-"BUS"
+ max_street_time = 30
+
+ trip <- detailed_itineraries( r5_core,
+                              fromLat = fromLat,
+                              fromLon = fromLon,
+                              toLat = toLat,
+                              toLon = toLon,
+                              trip_date = trip_date,
+                              departure_time = departure_time,
+                              direct_modes = direct_modes,
+                              transit_modes = transit_modes,
+                              max_street_time = max_street_time
+                             )
+```
+
+```{r}
+trip
+```

### 2020-08-15 r-package/R/detailed_itineraries.R

@@ -71,6 +71,9 @@ detailed_itineraries <- function(r5_core,
   direct_modes <- paste0(direct_modes, collapse = ";")
   transit_modes <- paste0(transit_modes, collapse = ";")
 
+  # Java expects street times to be integers. Casting the parameter to integer to make sure it works.
+  max_street_time = as.integer(max_street_time)
+
   # Call to method inside R5RCore object
   path_options <- r5_core$planSingleTrip(fromLat, fromLon, toLat, toLon,
                                              direct_modes, transit_modes,

### 2020-08-15 r-package/tests_marcus/rjava_tests.Rmd

@@
 ```{r}
-trip
+ library(r5r)
+ library(data.table)
+ library(dplyr)
+ library(sf)
+
+ # build transport network
+ path <- system.file("extdata", package = "r5r")
+ r5_core <- setup_r5(data_path = path)
+
+ # load origin/destination points
+ points <- read.csv(system.file("extdata/poa_hexgrid.csv", package = "r5r"))
+
+ # input
+ trip_date <- "2019-03-17"
+ departure_time <- "14:00:00"
+ street_time = 15L
+ direct_modes <- c("WALK", "BICYCLE", "CAR")
+ transit_modes <-"BUS"
+ max_street_time = 30
+
+trip_requests <- data.frame(id = 1:5,
+                            fromLat = points[1:5,]$lat,
+                            fromLon = points[1:5,]$lon,
+                            toLat = points[96:100,]$lat,
+                            toLon = points[96:100,]$lon )
+
+trips <- multiple_detailed_itineraries( r5_core,
+                                         trip_requests,
+                                         trip_date = trip_date,
+                                         departure_time = departure_time,
+                                         direct_modes = direct_modes,
+                                         transit_modes = transit_modes,
+                                         max_street_time = max_street_time
+                                         )
 ```
+
+```{r}
+ trips <- multiple_detailed_itineraries( r5_core,
+                                         trip_requests,
+                                         trip_date = trip_date,
+                                         departure_time = departure_time,
+                                         direct_modes = direct_modes,
+                                         transit_modes = transit_modes,
+                                         max_street_time = max_street_time
+                                         )
+```
+
+```{r}
+trips %>%
+  ggplot() +
+  geom_sf(aes(colour = mode)) +
+  facet_grid(request~option)
+```

### 2020-08-15 r-package/tests_rafa/test_rafa.R
minor fix to vignette

@@ -165,26 +165,6 @@ system("R CMD Rd2pdf --title=Package gtfs2gps --output=./gtfs2gps/manual.pdf")
 
 
 
-# Ignore these files/folders when building the package (but keep them on github)
-setwd("R:/Dropbox/git_projects/gtfs2gps")
-
-
-usethis::use_build_ignore("test")
-usethis::use_build_ignore("prep_data")
-usethis::use_build_ignore("manual.pdf")
-
-# script da base de dados e a propria base armazenada localmente, mas que eh muito grande para o CRAN
-usethis::use_build_ignore("brazil_2010.R")
-usethis::use_build_ignore("brazil_2010.RData")
-usethis::use_build_ignore("brazil_2010.Rd")
-
-# Vignette que ainda nao esta pronta
-usethis::use_build_ignore("  Georeferencing-gain.R")
-usethis::use_build_ignore("  Georeferencing-gain.Rmd")
-
-# temp files
-usethis::use_build_ignore("crosswalk_pre.R")
-
 
 
 ### CMD Check ----------------

### 2020-08-19 r-package/paper2020/paper2020.Rmd
kickstart of package paper

---
title: 'r5r: Rapid Realistic Routing with R5 in R'
author: "   "
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette: default
  github_document: default
abstract: " write abstract"
urlcolor: blue
vignette: |
  %\VignetteIndexEntry{r5r-intro} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

#> Examples of similar papers
#> - https://transportfindings.org/article/8416-accessibility-toolbox-for-r-and-arcgis
#> 
#> - https://transportfindings.org/article/6945-dodgr-an-r-package-for-network-flow-aggregation
#> 
#> - https://joss.theoj.org/papers/10.21105/joss.01926

# RESEARCH QUESTIONS AND HYPOTHESES

Transport routing is a fundamental step for the transport modeling required in 
accessibility analysis, fleet allocation and transport simulation and planning 
more broadly.

context that researcher and practitioners face practical challenges in transport routing
and accessibility analysis.
- licensed software
- time consuming compuations particularly for large and complex transport
mulimodal networks
- most options available only focus on fastest route, whach overlooks...

This article presents **r5r**, a new open-source `R` package for rapid 
realistic routing on multimodal transport networks.

# METHODS AND DATA

The **r5r** package provides a simple and convenient friendly interface to R<sup>5</sup>, [Rapid Realistic Routing on Real-world and Reimagined networks](https://github.com/conveyal/r5). The  developed by [Conveyal](https://www.conveyal.com/).

papers under R5 we should cite
- https://repository.asu.edu/items/54162
- https://www.jtlu.org/index.php/jtlu/article/view/1074

# FINDINGS

```{r, message = FALSE}
# Load packages
library(r5r)
library(sf)
library(data.table)
library(dplyr)
library(ggplot2)

```

To illustrate functionality, the package includes a small sample data for the 
city of Porto Alegre (Brazil). It includes three files:
- An Open Street Map network in `.pbf` format (*mandatory*)
- A public transport network in `GTFS.zip` format
- The spatial coordinates of the origin destination pairs in a `.csv` file.

```{r}
path <- system.file("extdata", package = "r5r")
list.files(path)

points <- fread(system.file("extdata/poa_hexgrid.csv", package = "r5r"))
points <- points[ c(sample(1:nrow(points), 10, replace=TRUE)), ]
head(points)

```

By default, R allocates only 512MB of memory for Java processes, which is not enough for large queries using r5r. To increase available memory to 4GB, for example, we need to set the java.parameters option at the beginning of the script, as follows:

```{r, message = FALSE}
options(java.parameters = "-Xmx4G")
```

### Build routable transport network

The first step is to build the multimodal transport used for routing in R5. This
is done with the `setup_r5` function. It does two things: (1) downloads a compiled JAR file
of R5 and store it locally in the r5r package directory; (2) combines the osm.pbf and gtfs.zip data sets to build a routable network object. The function

```{r, message = FALSE}
# Indicate the path where OSM and GTFS data are stored
r5r_core <- setup_r5(data_path = path)

```

**r5** has two core routing functions.

- `detailed_itineraries`
   * Returns a `data.frame sf LINESTRINGs` with one or multiple alternative routes
   between one or multiple origin destination pairs. The data output brings 
   detailed information on transport mode, travel time, walk distance etc for 
   each trip section

- `travel_time_matrix`
   * Fast function that returns a simple 'data.frame' with travel time 
   estimates between one or multiple origin destination pairs.

## Fast many to many travel time matrix

```{r, message = FALSE}
# calculate a travel time matrix
ttm <- travel_time_matrix( r5r_core = r5r_core,
                           origins = points,
                           destinations = points,
                           trip_date = "2019-05-20",
                           departure_time = "14:00:00",
                           mode = c("WALK", "TRANSIT"),
                           max_street_time = 7200,
                           max_trip_duration = 7200
                          )
head(ttm)
```

## Detailed itineraries

Allows one to get the fastest or multiple alternative routes between a set of
origins and destinations.

Create simple example here and plot the result below

### visualize results

```{r, message = FALSE}
# extract OSM network
street_net <- street_network_to_sf(r5r_core)

# plot
ggplot() +
        geom_sf(data = street_net$edges, color='gray85') +
        theme_minimal()

```

# Acknowledgments

# References

### 2020-08-19 r-package/tests_marcus/tests r5r.Rmd
Changed requestId in detailed itineraries to fromId and toId fields

---
title: "R5R"
output: html_document
---

```{r setup, include=FALSE}
options(java.parameters = "-Xmx8G")
knitr::opts_chunk$set(echo = TRUE)
library("r5r")
library("tidyverse")
library("sf")
library("tmap")
library("mapproj")
```

##

```{r}
r5r_core <- setup_r5(system.file("extdata", package = "r5r"), quiet = TRUE)
```

```{r}
r5r_core$silentMode()
r5r_core$verboseMode()
r5r_core$setLogMode("ERROR")
r5r_core$setLogMode("OFF")
```

```{r}
fromLat <- -29.997611
fromLon <- -51.197720
toLat <- -30.048951
toLon <- -51.229533

origin <- tibble(id="1", lat = fromLat, lon = fromLon)
destination <- tibble(id="2", lat=toLat, lon = toLon)

trip_date <- "2019-03-20"
departure_time <- "14:00:00"
street_time = 120L

paths_bus_df <- detailed_itineraries(r5r_core, 
                                     origins = origin, destinations = destination,
                                     mode = c("WALK", "BUS"), 
                                     trip_date = trip_date, departure_time = departure_time,
                                     max_street_time = street_time, shortest_path = TRUE)

```

```{r}
street_net <- street_network_to_sf(r5r_core)
```

```{r}
best_options <- paths_bus_df %>%
  st_set_geometry(NULL) %>%
  mutate(route = if_else(route=="", mode, route)) %>%
  group_by(option) %>%
  summarise(pattern = paste(route, collapse = " "), duration = sum(duration), .groups = "drop") %>%
  group_by(pattern) %>%
  arrange(duration) %>%
  slice(1)
```

```{r}
paths_bus_df %>% filter(option==2) %>% View()
```

```{r}
paths_bus_df %>% filter(option %in% best_options$option) %>%
  ggplot() +
  geom_sf(data=street_net$edges, colour="grey85", aes(geometry=geometry)) +
  geom_sf(aes(geometry=geometry, colour=mode)) +
  facet_wrap(~option)
```

```{r}
points <- read_csv(system.file("extdata/poa_hexgrid.csv", package = "r5r"))
```

```{r}

```

```{r}
ttm <- travel_time_matrix(r5r_core, points, points, trip_date, departure_time, mode=c("WALK", "BUS"))
```

```{r}
ttm
```

```{r}
orig <- (points %>% sample_n(1))$id

ttm %>% filter(fromId==orig) %>%
  ggplot() +
  geom_point(aes(x=toLon, y=toLat, colour=travel_time)) +
  scale_color_distiller(palette = "Spectral") +
  coord_map()
```

```{r}
# origins
"89a9012804fffff"

```

### 2020-08-21 r-package/tests_marcus/tests r5r.Rmd
Updated documentation and tests.Rmd

@@
 knitr::opts_chunk$set(echo = TRUE)
-library("r5r")
+# library("r5r")
 library("tidyverse")
@@
 library("mapproj")
+library("lubridate")
 ```
@@
 ```{r}
-r5r_core <- setup_r5(system.file("extdata", package = "r5r"), quiet = TRUE)
+r5r_core <- setup_r5(system.file("extdata", package = "r5r"), verbose = FALSE)
 ```
@@
 departure_time <- "14:00:00"
+trip_date_time <- lubridate::as_datetime("2019-03-20 14:00:00")
+
 street_time = 120L
 
-paths_bus_df <- detailed_itineraries(r5r_core, 
-                                     origins = origin, destinations = destination,
-                                     mode = c("WALK", "BUS"), 
-                                     trip_date = trip_date, departure_time = departure_time,
-                                     max_street_time = street_time, shortest_path = TRUE)
+paths_df <- detailed_itineraries(r5r_core = r5r_core, origins = origin, destinations = destination, 
+                     departure_datetime = trip_date_time, max_walk_dist = 800, mode = c("WALK", "BUS"),
+                     shortest_path = FALSE, verbose = FALSE)
 
@@
 ```{r}
-best_options <- paths_bus_df %>%
-  st_set_geometry(NULL) %>%
-  mutate(route = if_else(route=="", mode, route)) %>%
-  group_by(option) %>%
-  summarise(pattern = paste(route, collapse = " "), duration = sum(duration), .groups = "drop") %>%
-  group_by(pattern) %>%
-  arrange(duration) %>%
-  slice(1)
-```
-
-```{r}
-paths_bus_df %>% filter(option==2) %>% View()
-```
-
-```{r}
-paths_bus_df %>% filter(option %in% best_options$option) %>%
+paths_df %>%
   ggplot() +
@@
 ```{r}
-ttm <- travel_time_matrix(r5r_core, points, points, trip_date, departure_time, mode=c("WALK", "BUS"))
+ttm <- travel_time_matrix(r5r_core, points, points, departure_datetime = trip_date_time, 
+                          mode=c("WALK", "BUS"), 
+                          max_walk_dist = 500, max_trip_duration = 120L)
 ```
@@
 ```{r}
-orig <- (points %>% sample_n(1))$id
+orig <- (points %>% sample_n(4))$id
 
-ttm %>% filter(fromId==orig) %>%
+ttm %>% filter(fromId %in% orig) %>%
   ggplot() +
+  geom_point(data=points, aes(x=lon, y=lat), colour="grey85") +
   geom_point(aes(x=toLon, y=toLat, colour=travel_time)) +
   scale_color_distiller(palette = "Spectral") +
-  coord_map()
+  coord_map() +
+  facet_wrap(~fromId)
 ```

### 2020-08-21 r-package/tests_marcus/tests r5r.Rmd
Fixed travel distance in detailed_itineraries. Converted units from seconds and millimetres to minutes and meters.

@@
 library("lubridate")
+library("purrr")
 ```
@@
 
-```{r}
+```{r include=FALSE}
 r5r_core <- setup_r5(system.file("extdata", package = "r5r"), verbose = FALSE)
@@
 ```
+```{r}
+
+```
+
+```{r}
+paths_df %>%
+  mutate(dist_m = distance, dist_geo = st_length(geometry)) %>% View()
+
+  mutate(dist_sum = map_dbl(route, function(txt) {
+           txt_list <- str_split(txt, pattern = " ", simplify = TRUE)
+           num_list = as.numeric(txt_list)
+           return(sum(num_list))
+         }))
+```
+
+```{r}
+paths_df %>% st_write("/Users/marcussaraiva/paths3.shp")
+street_net$edges %>% st_write("/Users/marcussaraiva/streets.shp")
+```
 
@@
 points <- read_csv(system.file("extdata/poa_hexgrid.csv", package = "r5r"))
+poi <- read_csv(system.file("extdata/poa_points_of_interest.csv", package = "r5r"))
 ```

### 2020-08-21 java-r5/src/main/java/com/conveyal/r5/R5RCore.java

@@ -292,8 +292,20 @@ public class R5RCore {
                         optionCol.add(optionIndex);
                         segmentCol.add(1);
                         modeCol.add(segment.mode.toString());
-                        durationCol.add(segment.duration);
-                        distanceCol.add(segment.distance);
+                        durationCol.add(segment.duration / 60); // Converting from seconds to minutes
+
+//                        distanceCol.add(segment.distance);
+                        // try getting distances from street edges
+                        int dist = 0;
+                        StringBuilder distances = new StringBuilder();
+                        for (int i = 0; i < segment.streetEdges.size(); i++) {
+                            dist += segment.streetEdges.get(i).distance;
+                            distances.append(segment.streetEdges.get(i).distance).append(" ");
+                        }
+                        distanceCol.add(dist / 1000); // Converting from millimeters to meters
+//                        routeCol.add(distances.toString());
+
+
                         routeCol.add("");
                         waitCol.add(0);
                         geometryCol.add(segment.geometry.toString());
@@ -317,9 +329,22 @@ public class R5RCore {
                         segmentIndex++;
                         segmentCol.add(segmentIndex);
                         modeCol.add(segment.mode.toString());
-                        durationCol.add(segment.duration);
-                        distanceCol.add(segment.distance);
+                        durationCol.add(segment.duration / 60);
+
+//                        distanceCol.add(segment.distance);
+                        // try getting distances from street edges
+                        int dist = 0;
+                        StringBuilder distances = new StringBuilder();
+                        for (int i = 0; i < segment.streetEdges.size(); i++) {
+                            dist += segment.streetEdges.get(i).distance;
+                            distances.append(segment.streetEdges.get(i).distance).append(" ");
+                        }
+                        distanceCol.add(dist / 1000);
+//                        routeCol.add(distances.toString());
+
+
                         routeCol.add("");
+
                         waitCol.add(0);
                         geometryCol.add(segment.geometry.toString());
                     }
@@ -356,7 +381,7 @@ public class R5RCore {
                         optionCol.add(optionIndex);
                         segmentCol.add(segmentIndex);
                         modeCol.add(transit.mode.toString());
-                        durationCol.add(transit.rideStats.avg);
+                        durationCol.add(transit.rideStats.avg/60);
                         distanceCol.add(-1);
[…]

### 2020-08-21 r-package/tests_marcus/tests r5r.Rmd
Updated documentation and test.rmd

@@
 paths_df <- detailed_itineraries(r5r_core = r5r_core, origins = origin, destinations = destination, 
-                     departure_datetime = trip_date_time, max_walk_dist = 800, mode = c("WALK", "BUS"),
+                     departure_datetime = trip_date_time, max_walk_dist = 0.8, 
+                     mode = c("WALK", "BICYCLE", "BUS"),
                      shortest_path = FALSE, verbose = FALSE)
@@
   ggplot() +
-  geom_sf(data=street_net$edges, colour="grey85", aes(geometry=geometry)) +
+  # geom_sf(data=street_net$edges, colour="grey85", aes(geometry=geometry)) +
   geom_sf(aes(geometry=geometry, colour=mode)) +
@@
                           mode=c("WALK", "BUS"), 
-                          max_walk_dist = 500, max_trip_duration = 120L)
+                          max_walk_dist = 500, max_trip_duration = 120L, verbose = FALSE)
 ```

### 2020-08-22 r-package/tests_marcus/tests r5r.Rmd
Testing cleaner version of travel time matrix.

@@
 ttm %>% filter(fromId %in% orig) %>%
+  left_join(points, by=c("toId" = "id")) %>%
   ggplot() +
   geom_point(data=points, aes(x=lon, y=lat), colour="grey85") +
-  geom_point(aes(x=toLon, y=toLat, colour=travel_time)) +
+  geom_point(aes(x=lon, y=lat, colour=travel_time)) +
   scale_color_distiller(palette = "Spectral") +

### 2020-08-22 r-package/tests_marcus/tests r5r.Rmd
Fixed detailed_itineraries function return when drop_geometry = TRUE

@@
                      mode = c("WALK", "BICYCLE", "BUS"),
-                     shortest_path = FALSE, verbose = FALSE)
+                     shortest_path = FALSE, verbose = FALSE, drop_geometry = FALSE)
 
@@
 street_net <- street_network_to_sf(r5r_core)
+```
+
+```{r}
+paths_df
 ```

### 2020-08-22 r-package/R/detailed_itineraries.R

@@ -254,11 +254,14 @@ detailed_itineraries <- function(r5r_core,
 
   path_options[, grep("temp_", names(path_options), value = TRUE) := NULL]
 
-  # convert path_options from data.frame to data.table with sfc column
-  data.table::setDT(path_options)[, geometry := sf::st_as_sfc(geometry)]
+  # if results includes the geometry, convert path_options from data.frame to
+  # data.table with sfc column
+  if (!drop_geometry) {
+    data.table::setDT(path_options)[, geometry := sf::st_as_sfc(geometry)]
 
-  # convert path_options from data.table to sf with CRS WGS 84 (EPSG 4326)
-  path_options <- sf::st_sf(path_options, crs = 4326)
+    # convert path_options from data.table to sf with CRS WGS 84 (EPSG 4326)
+    path_options <- sf::st_sf(path_options, crs = 4326)
+  }
 
   return(path_options)
 

### 2020-08-23 r-package/paper2020/paper2020.Rmd
updates do paper2020.Rmd draft intro

@@
 
-Transport routing is a fundamental step for the transport modeling required in 
+Transport routing is the process of finding the alternative routes and costs 
+that connect places on a given transport network. Routing often involves complex
+mathematical algorithms and it is a fundamental step required for transport
 accessibility analysis, fleet allocation and transport simulation and planning
@@
 
-context that researcher and practitioners face practical challenges in transport routing
-and accessibility analysis.
-- licensed software
-- time consuming compuations particularly for large and complex transport
-mulimodal networks
-- most options available only focus on fastest route, whach overlooks...
+Despite the long standing importance of routing analysis to the field, transport 
+researchers and practitioners still frequently face practical challenges with 
+this task. These challenges involve for example the monetary costs of licensed 
+software, data availability issues for data hungry models, and the long 
+computation times to run multiple routing scenarios, particularly in large and 
+complex multimodal transport networks. While there is a growing number of open source routing models (ref: stplnr, otp, dodgr), most options available only return the fastest route, what overlooks....
 

### 2020-08-24 r-package/tests/testthat/test-street_network_to_sf.R
tests with skip_on_cran and skip_on_travis

@@ -1,5 +1,9 @@
 context("street_network_to_sf")
 
+testthat::skip_on_cran()
+testthat::skip_on_travis()
+
+
 # setup_r5
 path <- system.file("extdata", package = "r5r")
 r5_core <- setup_r5(data_path = path, verbose=FALSE)

### 2020-08-24 r-package/R/zzz.R
add  eval = FALSE to vignette

@@ -19,7 +19,7 @@ NULL
 ## quiets concerns of R CMD check re: the .'s that appear in pipelines
 if(getRversion() >= "2.15.1") utils::globalVariables(
   c('is', 'duration', 'fromId', 'toId', 'option', 'option', '.SD', 'geometry',
-    'route', 'temp_duration', 'temp_route', 'route', 'temp_sign'))
+    'route', 'temp_duration', 'temp_route', 'route', 'temp_sign', '.I'))
 
 
 

### 2020-08-24 r-package/tests_rafa/test_rafa.R
add 1st plot to vignette

@@ -411,5 +411,5 @@ system("R CMD check --as-cran gtfs2gps_1.0-0.tar.gz")
 
 
 
-
+ggsave('ggplot.png', width = 16, height = 8, units = 'cm', dpi=200)
 

### 2020-08-24 r-package/doc/intro_to_r5r.Rmd

---
title: 'r5r: Rapid Realistic Routing with R5 in R'
author: "Rafael H. M. Pereira, Marcus Saraiva, Daniel Herszenhut, Carlos Kaue Braga"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette: default
  github_document: default
abstract: "`r5r` is an R package for rapid realistic routing on multimodal transport networks (walk, bike, public transport and car) using R5. The package allows users to generate detailed routing analysis or calculate travel time matrices using seamless parallel computing on top of the R5 Java machine <https://github.com/conveyal/r5>"
urlcolor: blue
vignette: >
  %\VignetteIndexEntry{r5r-intro} 
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown} 
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

**r5r** is an [R package for rapid realistic routing on multimodal transport networks](https://github.com/ipeaGIT/r5r) (walk, bike, public transport and car).
It provides a simple and friendly interface to R5, a really fast and open source 
routing engine based in `Java` developed by [Conveyal](https://www.conveyal.com/).
R5 stands for [Rapid Realistic Routing on Real-world and Reimagined networks](https://github.com/conveyal/r5).

# Installation
To use `r5r`, you need to have [Java SE Development Kit 11.0.8](https://www.oracle.com/java/technologies/javase-jdk11-downloads.html) 
installed on your computer. No worries, it's free.

Install the development version of `r5r` from GitHub (soon on CRAN) and load packages.

```{r, eval = FALSE}
devtools::install_github("ipeaGIT/r5r", subdir = "r-package")

```

```{r, message = FALSE}
library(r5r)
library(sf)
library(data.table)
library(ggplot2)
library(mapview)

```

The package has 3 fundamental functions.

1. `setup_r5` to build a routable transport network given a street and public
transport networks

2. `travel_time_matrix` for fast computation of travel time estimates between 
origin destination pairs

3. `detailed_itineraries` to get detailed information on one or multiple 
alternative routes between origin destination pairs

Let's have a quick look at how these functions work using a sample data set.

# Demonstration on sample data

## Data
To illustrate functionality, the package includes a small sample data for the 
city of Porto Alegre (Brazil). It includes three files:

* An Open Street Map network in `.pbf` format (*mandatory*)
* A public transport network in `GTFS.zip` format
* The spatial coordinates of the origin destination pairs in a `.csv` file.

```{r}
data_path <- system.file("extdata", package = "r5r")
list.files(data_path)

```
The data with origin destination pairs looks like this below. In this example, 
we will be looking at the transport routes between ten random points in this 
data set.
```{r}
points <- fread(system.file("extdata/poa_hexgrid.csv", package = "r5r"))
points <- points[ c(sample(1:nrow(points), 10, replace=TRUE)), ]
head(points)

```
#### Set memory

Before we start, we need to increase the memory available to Java. This is
because by default `R` allocates only 512MB of memory for Java processes, which is
would not b enough for large queries using `r5r`. To increase available memory 
to 2GB, for example, we need to set the java.parameters option at the beginning 
of the script, as follows:

```{r, message = FALSE, eval = FALSE}
options(java.parameters = "-Xmx2G")
```

## Build routable transport network with `setup_r5()`

The first step is to build the multimodal transport network used for routing in 
R5. This is done with the `setup_r5` function. This function does two things: (1)
downloads/updates a compiled JAR file of R5 and stores it locally in the `r5r` package 
directory for future use; and (2) combines the osm.pbf and gtfs.zip data sets to
build a routable network object.

```{r, message = FALSE, eval = FALSE}
# Indicate the path where OSM and GTFS data are stored
r5r_core <- setup_r5(data_path = data_path, verbose = FALSE)

```

## Routing analysis

For fast routing analysis, **r5r** currently has two core functions: `travel_time_matrix` 
and `detailed_itineraries`.

### Fast many to many travel time matrix

The `travel_time_matrix` function is a really simple and fast function to compute
travel time estimates between one or multiple origin destination pairs. The 
origin/destination input can be either a spatial `sf POINT` object, or a 
`data.frame` containing the columns `id, lon, lat`.

```{r, message = FALSE, eval = FALSE}
# calculate a travel time matrix
ttm <- travel_time_matrix( r5r_core = r5r_core,
                           origins = points,
                           destinations = points,
                           departure_datetime = lubridate::as_datetime("2019-03-20 14:00:00"),
                           mode = c("WALK", "TRANSIT"),
                           max_walk_dist = 5,
                           max_trip_duration = 7200,
                           verbose = FALSE
                          )
head(ttm)
```

```{r ttm head, echo = FALSE, message = FALSE}
knitr::include_graphics(system.file("img", "vig_output_ttm.png", package="r5r"))
```

### Detailed itineraries

Most routing packages only return the fastest route. A key advantage of the 
`detailed_itineraries` function is that is allows for fast routing analysis while
providing multiple alternative routes between origin destination pairs. The output
also brings detailed information for each route alternative at the trip segment
level, including the transport mode, waiting times, travel time and distance of
each trip segment.

In this example below, we want to know some alternative routes between one origin
destination pair only.

```{r, message = FALSE, eval = FALSE}
# inputs
points <- read.csv(file.path(data_path, "poa_points_of_interest.csv"))
origins <- points[10,]
destinations <- points[12,]
mode = c("WALK", "TRANSIT")
max_walk_dist <- 10000
departure_datetime <- lubridate::as_datetime("2019-03-20 14:00:00")

df <- detailed_itineraries(r5r_core = r5r_core,
                         origins,
                         destinations,
                         mode,
                         departure_datetime,
                         max_walk_dist,
                         shortest_path = FALSE,
                         verbose = FALSE)

head(df)
```
```{r detailed head, echo = FALSE, message = FALSE}
knitr::include_graphics(system.file("img", "vig_output_detailed.png", package="r5r"))
```

The output is a `data.frame sf` object, so we can easily visualize the results.

#### Visualize results

**Static visualization** with `ggplot2` package: To facilitate the 
visualization of the results in `ggplot2`, you can also use the 
`street_network_to_sf` function to  extract the OSM street network used in the
routing.

```{r, message = FALSE, eval = FALSE}
# extract OSM network
street_net <- street_network_to_sf(r5r_core)

# plot
ggplot() +
  geom_sf(data = street_net$edges, color='gray85') +
  geom_sf(data = df, aes(color=mode)) +
  facet_wrap(.~option) + 
  theme_void()

```

```{r ggplot2 output, echo = FALSE, message = FALSE}
knitr::include_graphics(system.file("img", "vig_detailed_ggplot.png", package="r5r"))
```

**Interactive visualization** with `mapview`:

```{r, message = FALSE, eval = FALSE}
mapviewOptions(platform = 'leafgl')
mapview(df, zcol = 'option')

```

```{r mapview output, echo = FALSE, message = FALSE}
knitr::include_graphics(system.file("img", "vig_detailed_mapview.png", package="r5r"))
```

If you have any suggestions or want to report an error, please visit the [the package GitHub page](https://github.com/ipeaGIT/r5r).

### 2020-08-25 r-package/paper2020/paper2020.Rmd
tests with skip_on_cran and skip_on_travis

@@
 
-This article presents **r5r**, a new open-source `R` package for rapid 
-realistic routing on multimodal transport networks.
+...
 
@@
 
+Section to demonstrate how to use the core functions of the package. We can use 
+our intro vignette as a starting point and build from there with more the analysis
+of larger data sets, include couple more data visualization of results and emphasize
+the results of multiple route alternatives.
+
+To illustrate functionality, the package includes a small sample data for the 
+city of Porto Alegre (Brazil). It includes three files:
+- An Open Street Map network in `.pbf` format (*mandatory*)
+- A public transport network in `GTFS.zip` format
+- The spatial coordinates of the origin destination pairs in a `.csv` file.
+
 ```{r, message = FALSE}
@@
 ```
-
-To illustrate functionality, the package includes a small sample data for the 
-city of Porto Alegre (Brazil). It includes three files:
-- An Open Street Map network in `.pbf` format (*mandatory*)
-- A public transport network in `GTFS.zip` format
-- The spatial coordinates of the origin destination pairs in a `.csv` file.
 
@@
 
+Conveyal for R5
+
 # References
+
+#> Rafa's notes
+#> jean: #>-  mass of assumptions regarding decision-making and routing (Miller, 2018). 
+#>Haugen, 2011; Schwanen, 2008
+#>
+#>Miller, E.J., 2018. Accessibility: measurement and application in transportation planning. Transp. Rev. 38, #>551–555. https://doi.org/10.1080/01441647.2018.1492778
+#>
+#>This article presents **r5r**, a new open-source `R` package for rapid 
+#>realistic routing on multimodal transport networks.

### 2020-09-01 r-package/R/detailed_itineraries.R
specify timezone parameters to datetime functions

@@ -64,7 +64,8 @@
 #' max_walk_dist <- 1000
 #' max_trip_duration <- 120L
 #' departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
-#'                                  format = "%d-%m-%Y %H:%M:%S")
+#'                                  format = "%d-%m-%Y %H:%M:%S",
+#'                                  tz = "America/Sao_Paulo")
 #'
 #' df <- detailed_itineraries(r5r_core,
 #'                            origins,

### 2020-09-02 r-package/R/detailed_itineraries.R
remove lubridate suggestion dependency

@@ -67,12 +67,12 @@
 #'                                  format = "%d-%m-%Y %H:%M:%S",
 #'                                  tz = "America/Sao_Paulo")
 #'
-#' df <- detailed_itineraries(r5r_core,
-#'                            origins,
-#'                            destinations,
-#'                            mode,
-#'                            departure_datetime,
-#'                            max_walk_dist)
+#' dit <- detailed_itineraries(r5r_core,
+#'                             origins,
+#'                             destinations,
+#'                             mode,
+#'                             departure_datetime,
+#'                             max_walk_dist)
 #'
 #' stop_r5(r5r_core)
 #' rJava::.jgc(R.gc = TRUE)

### 2020-09-02 r-package/R/travel_time_matrix.R

@@ -51,20 +51,20 @@
 #' points <- read.csv(system.file("extdata/poa_hexgrid.csv", package = "r5r"))[1:5,]
 #'
 #' mode <- c("WALK", "TRANSIT")
-#' max_walk_dist = Inf
+#' max_walk_dist <- Inf
 #' max_trip_duration <- 120L
 #' departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
 #'                                  format = "%d-%m-%Y %H:%M:%S",
 #'                                  tz = "America/Sao_Paulo")
 #'
 #' # estimate travel time matrix
-#' df <- travel_time_matrix(r5r_obj,
-#'                          origins = points,
-#'                          destinations = points,
-#'                          mode,
-#'                          departure_datetime,
-#'                          max_walk_dist,
-#'                          max_trip_duration)
+#' ttm <- travel_time_matrix(r5r_obj,
+#'                           origins = points,
+#'                           destinations = points,
+#'                           mode,
+#'                           departure_datetime,
+#'                           max_walk_dist,
+#'                           max_trip_duration)
 #'
 #' stop_r5(r5r_obj)
 #' rJava::.jgc(R.gc = TRUE)

### 2020-09-15 r-package/paper2020/paper2020.Rmd
paper: 1st paragraph done

@@
 accessibility analysis, fleet allocation and transport simulation and planning
-more broadly.
-
-Despite the long standing importance of routing analysis to the field, transport 
-researchers and practitioners still frequently face practical challenges with 
-this task. These challenges involve for example the monetary costs of licensed 
-software, data availability issues for data hungry models, and the long 
-computation times to run multiple routing scenarios, particularly in large and 
-complex multimodal transport networks. While there is a growing number of open source routing models (ref: stplnr, otp, dodgr), most options available only return the fastest route, what overlooks....
+more broadly. Despite the long-standing importance of routing analysis to the 
+field, transport researchers and practitioners still frequently face practical 
+challenges with this task. These challenges involve for example the monetary 
+costs of licensed software, data availability issues for data hungry models, and
+the long computation times to run multiple routing scenarios, particularly in 
+large and complex multimodal transport networks. While there is a growing number
+of open source routing models (ref: stplnr, otp, dodgr), most options available 
+only return the fastest route option between a set of origins and destinations. 
+This simple approach overlooks how often people's actual route choices deviate 
+from shortest travel time paths due to imperfect information, monetary costs and
+safety issues among other reasons (refs, 1 2 3).
+Refs:
+1 - https://tram.mcgill.ca/Research/Publications/Who_will_use_it.pdf
+2 - https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0134322
+3 - https://www.tandfonline.com/doi/abs/10.1080/13658816.2019.1605075
 

### 2020-09-15 r-package/paper2020/paper2020.Rmd
paper paragraph 2 done

@@
 
-...
+This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open source 
+R package for rapid realistic routing on multimodal transport networks based on
+the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5). R<sup>5</sup> is a powerful 
+next-generation routing engine written in Java and developed at Conveyal (ref 1-2)
+to replace its predecessor OpenTripPlanner. The r5r package provides a simple 
+and friendly interface to run R<sup>5</sup> locally from within R, what allows 
+users to efficiently calculate travel time matrices or generate multiple route 
+alternatives between origins and destinations using seamless parallel computing.
+1 - https://repository.asu.edu/items/54162
+2- https://www.jtlu.org/index.php/jtlu/article/view/1074
 
@@
 
-The **r5r** package provides a simple and convenient friendly interface to R<sup>5</sup>, [Rapid Realistic Routing on Real-world and Reimagined networks](https://github.com/conveyal/r5).
-The  developed by [Conveyal](https://www.conveyal.com/).
-
-papers under R5 we should cite
-- https://repository.asu.edu/items/54162
-- https://www.jtlu.org/index.php/jtlu/article/view/1074
+p
 

### 2020-09-15 r-package/paper2020/paper2020.Rmd
paper2020 - METHODS AND DATA section

@@
 
-p
+r5r has low data requirements, enabling routing analysis on 
+different contexts wherever there is open data on street networks from 
+OpenStreetMap (OSM). The package is also easily scalable allowing for fast 
+computations at either the city or country level analysis. For multimodal 
+transport routing, r5r will automatically combine OpenStreetMap with one or 
+multiple public transport data sets in the stantard GTFS format to create a routable
+transport network.
+
+The r5r package has 3 fundamental functions. The `setup_r5()` function builds a 
+multimodal transport network used for routing in R5. This function (1) 
+downloads/updates a compiled JAR file of R5 and stores it locally in the r5r
+package directory for future use; and (2) combines the OSM and GTFS data sets to
+build a routable network object. For fast routing analysis, r5r currently has 
+two functions:
+
+- `travel_time_matrix`: a simple and fast function to efficiently compute travel
+time estimates between one or multiple origin/destination pairs. The origin/destination
+input can be either a spatial `sf POINT` object, or a `data.frame` containing the
+columns `id`, `lon`, `lat.`
+
+- `detailed_itineraries`:  function that provides the shortest or 
+multiple alternative routes between origin/destination pairs. The output also 
+brings detailed information for each route alternative at the trip segment level,
+including the transport mode, waiting times, travel time and distance of each 
+trip segment.
+
+Both routing functions are versatile so users can easily can set customized inputs
+such as transport mode, departure date and times, walking  an cycling speeds, 
+maximum trip duration, walking distances or number of public transport transfers
+etc.
 

### 2020-09-17 r-package/paper2020/paper2020.Rmd
updates paper's travel time matrix example

@@
 ```{r, message = FALSE}
-# calculate a travel time matrix
-ttm <- travel_time_matrix( r5r_core = r5r_core,
+mode <- c("WALK", "TRANSIT")
+max_walk_dist <- Inf
+max_trip_duration <- 120L
+departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
+                                 format = "%d-%m-%Y %H:%M:%S",
+                                 tz = "America/Sao_Paulo")
+
+# estimate travel time matrix
+ttm <- travel_time_matrix(r5r_core,
                           origins = points,
                           destinations = points,
-                           trip_date = "2019-05-20",
-                           departure_time = "14:00:00",
-                           mode = c("WALK", "TRANSIT"),
-                           max_street_time = 7200,
-                           max_trip_duration = 7200
-                          )
+                          mode,
+                          departure_datetime,
+                          max_walk_dist,
+                          max_trip_duration)
+
 head(ttm)

### 2020-09-17 r-package/paper2020/paper2020.Rmd
removes tz argument from datetimes in examples

@@
   %\VignetteIndexEntry{r5r-intro} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
+bibliography: references.json
 ---
@@
 3 - https://www.tandfonline.com/doi/abs/10.1080/13658816.2019.1605075
+
+Test [@http://zotero.org/users/local/FtLgDWw9/items/8CHCA2S3]
 
@@
 departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
-                                 format = "%d-%m-%Y %H:%M:%S",
-                                 tz = "America/Sao_Paulo")
+                                 format = "%d-%m-%Y %H:%M:%S")
 

### 2020-09-17 r-package/R/detailed_itineraries.R

@@ -64,8 +64,7 @@
 #' max_walk_dist <- 1000
 #' max_trip_duration <- 120L
 #' departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
-#'                                  format = "%d-%m-%Y %H:%M:%S",
-#'                                  tz = "America/Sao_Paulo")
+#'                                  format = "%d-%m-%Y %H:%M:%S")
 #'
 #' dit <- detailed_itineraries(r5r_core,
 #'                             origins,

### 2020-09-17 r-package/R/travel_time_matrix.R

@@ -45,7 +45,7 @@
 #'
 #' # build transport network
 #' data_path <- system.file("extdata", package = "r5r")
-#' r5r_obj <- setup_r5(data_path = data_path)
+#' r5r_core <- setup_r5(data_path = data_path)
 #'
 #' # load origin/destination points
 #' points <- read.csv(system.file("extdata/poa_hexgrid.csv", package = "r5r"))[1:5,]
@@ -54,11 +54,10 @@
 #' max_walk_dist <- Inf
 #' max_trip_duration <- 120L
 #' departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
-#'                                  format = "%d-%m-%Y %H:%M:%S",
-#'                                  tz = "America/Sao_Paulo")
+#'                                  format = "%d-%m-%Y %H:%M:%S")
 #'
 #' # estimate travel time matrix
-#' ttm <- travel_time_matrix(r5r_obj,
+#' ttm <- travel_time_matrix(r5r_core,
 #'                           origins = points,
 #'                           destinations = points,
 #'                           mode,
@@ -66,7 +65,7 @@
 #'                           max_walk_dist,
 #'                           max_trip_duration)
 #'
-#' stop_r5(r5r_obj)
+#' stop_r5(r5r_core)
 #' rJava::.jgc(R.gc = TRUE)
 #'
 #' }

### 2020-09-17 r-package/paper2020/paper2020.Rmd
updates paper references

@@
   github_document: default
-abstract: " write abstract"
+abstract: "write abstract"
 urlcolor: blue
@@
 large and complex multimodal transport networks. While there is a growing number
-of open source routing models (ref: stplnr, otp, dodgr), most options available 
-only return the fastest route option between a set of origins and destinations. 
-This simple approach overlooks how often people's actual route choices deviate 
-from shortest travel time paths due to imperfect information, monetary costs and
-safety issues among other reasons (refs, 1 2 3).
-Refs:
-1 - https://tram.mcgill.ca/Research/Publications/Who_will_use_it.pdf
-2 - https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0134322
-3 - https://www.tandfonline.com/doi/abs/10.1080/13658816.2019.1605075
-
-Test [@http://zotero.org/users/local/FtLgDWw9/items/8CHCA2S3]
+of open source routing models [@lovelace2019stplanr; @padgham2019dodgr], most
+options available only return the fastest route option between a set of origins 
+and destinations. This simple approach overlooks how often people's actual route 
+choices deviate from shortest travel time paths due to imperfect information, 
+monetary costs and safety issues among other reasons [@larsen2011travel;
+@zhu2015people; @conway2019getting].
 
@@
 R package for rapid realistic routing on multimodal transport networks based on
-the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5). R<sup>5</sup> is a powerful 
-next-generation routing engine written in Java and developed at Conveyal (ref 1-2)
-to replace its predecessor OpenTripPlanner. The r5r package provides a simple 
-and friendly interface to run R<sup>5</sup> locally from within R, what allows 
-users to efficiently calculate travel time matrices or generate multiple route 
+the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5). 
+R<sup>5</sup> is a powerful next-generation routing engine written in Java and
+developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to
+replace its predecessor OpenTripPlanner. The r5r package provides a simple and
+friendly interface to run R<sup>5</sup> locally from within R, what allows users
+to efficiently calculate travel time matrices or generate multiple route 
 alternatives between origins and destinations using seamless parallel computing.
-1 - https://repository.asu.edu/items/54162
-2- https://www.jtlu.org/index.php/jtlu/article/view/1074
 

### 2020-09-17 r-package/paper2020/paper2020.Rmd
fixes comments in paper

@@
 
-#> Examples of similar papers
-#> - https://transportfindings.org/article/8416-accessibility-toolbox-for-r-and-arcgis
-#> 
-#> - https://transportfindings.org/article/6945-dodgr-an-r-package-for-network-flow-aggregation
-#> 
-#> - https://joss.theoj.org/papers/10.21105/joss.01926
+<!-- Examples of similar papers -->
+<!--  - https://transportfindings.org/article/8416-accessibility-toolbox-for-r-and-arcgis -->
+
+<!--  - https://transportfindings.org/article/6945-dodgr-an-r-package-for-network-flow-aggregation -->
+
+<!--  - https://joss.theoj.org/papers/10.21105/joss.01926 -->
 
@@
 
-#> Rafa's notes
-#> jean: #>-  mass of assumptions regarding decision-making and routing (Miller, 2018). 
-#>Haugen, 2011; Schwanen, 2008
-#>
-#>Miller, E.J., 2018. Accessibility: measurement and application in transportation planning. Transp. Rev. 38, #>551–555. https://doi.org/10.1080/01441647.2018.1492778
-#>
-#>This article presents **r5r**, a new open-source `R` package for rapid 
-#>realistic routing on multimodal transport networks.
+<!-- Rafa's notes -->
+<!-- jean: #>-  mass of assumptions regarding decision-making and routing (Miller, 2018).  -->
+<!-- Haugen, 2011; Schwanen, 2008 -->
+
+<!-- Miller, E.J., 2018. Accessibility: measurement and application in transportation planning. Transp. Rev. 38, #>551–555. https://doi.org/10.1080/01441647.2018.1492778 -->
+
+<!-- This article presents **r5r**, a new open-source `R` package for rapid  -->
+<!-- realistic routing on multimodal transport networks. -->

### 2020-09-17 r-package/paper2020/paper2020.Rmd
updates paper

@@
 title: 'r5r: Rapid Realistic Routing with R5 in R'
-author: "   "
+author: "Rafael H. M. Pereira, Marcus Saraiva, Daniel Herszenhut, Carlos Kaue Braga"
 date: "`r Sys.Date()`"
@@
 )
+
+set.seed(0)
 ```
@@
 <!--  - https://joss.theoj.org/papers/10.21105/joss.01926 -->
+
+dasdasdds
 
@@
 large and complex multimodal transport networks. While there is a growing number
-of open source routing models [@lovelace2019stplanr; @padgham2019dodgr], most
-options available only return the fastest route option between a set of origins 
-and destinations. This simple approach overlooks how often people's actual route 
-choices deviate from shortest travel time paths due to imperfect information, 
-monetary costs and safety issues among other reasons [@larsen2011travel;
-@zhu2015people; @conway2019getting].
+of open source routing models [@opentripplanneropentripplanner; 
+@lovelace2019stplanr; @padgham2019dodgr], most options available only return the
+fastest route option between a set of origins and destinations. This simple
+approach overlooks how often people's actual route choices deviate from shortest
+travel time paths due to imperfect information, monetary costs and safety issues
+among other reasons [@larsen2011travel; @zhu2015people; @conway2019getting].
 
@@
 
-r5r has low data requirements, enabling routing analysis on 
-different contexts wherever there is open data on street networks from 
-OpenStreetMap (OSM). The package is also easily scalable allowing for fast 
-computations at either the city or country level analysis. For multimodal 
-transport routing, r5r will automatically combine OpenStreetMap with one or 
-multiple public transport data sets in the stantard GTFS format to create a routable
-transport network.
+r5r has low data requirements, enabling routing analysis on different contexts 
+wherever there is open data on street networks from OpenStreetMap (OSM). The 
+package is also easily scalable allowing for fast computations at either the 
+city or country level analysis. For multimodal transport routing, r5r will 
+automatically combine OSM data with one or multiple public transport data sets 
+in the standard GTFS format to create a routable transport network.
 
 The r5r package has 3 fundamental functions. The `setup_r5()` function builds a 
-multimodal transport network used for routing in R5. This function (1) 
-downloads/updates a compiled JAR file of R5 and stores it locally in the r5r
-package directory for future use; and (2) combines the OSM and GTFS data sets to
-build a routable network object. For fast routing analysis, r5r currently has 
-two functions:
+multimodal transport network used for routing in R<sup>5</sup>. This function 
+(1) downloads/updates a compiled R<sup>5</sup> JAR file, tailored for the 
+purposes of the package, and stores it locally in the r5r package directory for 
+future use; and (2) combines the OSM and GTFS data sets to build a routable 
+network object. For fast routing analysis, r5r currently has two functions:
 
-- `travel_time_matrix`: a simple and fast function to efficiently compute travel
-time estimates between one or multiple origin/destination pairs. The origin/destination
-input can be either a spatial `sf POINT` object, or a `data.frame` containing the
-columns `id`, `lon`, `lat.`
+- `travel_time_matrix()`: used to compute travel time estimates between one or 
+multiple origin/destination pairs. The origin/destination input can be either a 
+spatial `sf POINT` object, or a `data.frame` containing the columns `id`, `lon`, 
+`lat.`
 
-- `detailed_itineraries`:  function that provides the shortest or 
-multiple alternative routes between origin/destination pairs. The output also 
-brings detailed information for each route alternative at the trip segment level,
-including the transport mode, waiting times, travel time and distance of each 
-trip segment.
+- `detailed_itineraries()`:  used to compute detailed information at the trip 
+segment level on either the shortest or multiple alternative routes between one
+or multiple origin/destination pairs. Includes information such as the transport 
+mode, waiting time, travel time and distance.
 
-Both routing functions are versatile so users can easily can set customized inputs
-such as transport mode, departure date and times, walking  an cycling speeds, 
-maximum trip duration, walking distances or number of public transport transfers
-etc.
+Both routing functions are versatile so users can easily set customized inputs
+such as transport mode, departure date and times, walking and cycling speeds, 
+maximum trip duration, walking distances and number of public transport 
+transfers, etc.
 
@@
 
-Section to demonstrate how to use the core functions of the package. We can use 
-our intro vignette as a starting point and build from there with more the analysis
-of larger data sets, include couple more data visualization of results and emphasize
-the results of multiple route alternatives.
+The package can be installed with the following command:
 
-To illustrate functionality, the package includes a small sample data for the 
-city of Porto Alegre (Brazil). It includes three files:
-- An Open Street Map network in `.pbf` format (*mandatory*)
-- A public transport network in `GTFS.zip` format
-- The spatial coordinates of the origin destination pairs in a `.csv` file.
+```{r, message = FALSE, eval = FALSE}
+devtools::install_github("ipeaGIT/r5r", subdir = "r-package")
+```
+
+The first step when using r5r is to set the amount of memory available to Java. 
+By default `R` allocates only 512 MB of memory for Java processes, which is not 
+enough for large queries using r5r. To increase available memory to 2GB, for 
+example, we need to set the `java.parameters` option at the beginning of the 
+script as follows:
 
 ```{r, message = FALSE}
-# Load packages
+options(java.parameters = "-Xmx2G")
+```
+
+Please note that `R` only allocates memory to Java once during a running 
+session. If a Java process has been initiated before, either by r5r or any other 
+Java-dependent package or function, it might be required to restart the current 
+`R` session in order to effectively change the amount of memory allocated.
+
+The package can be attached (with some other packages used in this article) with 
+the following command:
+
+```{r, message = FALSE}
 library(r5r)
@@
 library(data.table)
-library(dplyr)
 library(ggplot2)
-
 ```
 
+To illustrate functionality, the package includes a small sample data set for 
+the city of Porto Alegre (Brazil). It includes three files:
+
+- An OSM network in `.pbf` format (*mandatory*).
+- A public transport network in `GTFS.zip` format.
+- The spatial coordinates of points covering the area in `.csv` format.
+
+The `.pbf` and `GTFS.zip` files are used to build the transport network. While 
+the first is mandatory, since it is required to create a street network, the 
+latter is only used to calculate public transport trips, thus may not be 
+necessary if that's not intended.
+
 ```{r}
-path <- system.file("extdata", package = "r5r")
-list.files(path)
+data_path <- system.file("extdata", package = "r5r")
+list.files(data_path)
 
 points <- fread(system.file("extdata/poa_hexgrid.csv", package = "r5r"))
-points <- points[ c(sample(1:nrow(points), 10, replace=TRUE)), ]
+points <- points[c(sample(1:nrow(points), 100))]
 head(points)
-
-```
-
-By default, R allocates only 512MB of memory for Java processes, which is not enough for large queries using r5r.
-To increase available memory to 4GB, for example, we need to set the java.parameters option at the beginning of the script, as follows:
-
-```{r, message = FALSE}
-options(java.parameters = "-Xmx4G")
 ```
@@
 # Indicate the path where OSM and GTFS data are stored
-r5r_core <- setup_r5(data_path = path)
-
+r5r_core <- setup_r5(data_path)
 ```

### 2020-09-17 r-package/paper2020/paper2020.Rmd
updates travel time matrix paper section

@@
 
-- An OSM network in `.pbf` format (*mandatory*).
+- An OSM network in `.pbf` format.
 - A public transport network in `GTFS.zip` format.
 - The spatial coordinates of points covering the area in `.csv` format.
-
-The `.pbf` and `GTFS.zip` files are used to build the transport network. While 
-the first is mandatory, since it is required to create a street network, the 
-latter is only used to calculate public transport trips, thus may not be 
-necessary if that's not intended.
 
@@
 
-### Build routable transport network
+### Build a routable transport network
 
-The first step is to build the multimodal transport used for routing in R5. This
-is done with the `setup_r5` function. It does two things: (1) downloads a compiled JAR file
-of R5 and store it locally in the r5r package directory; (2) combines the osm.pbf and gtfs.zip data sets to build a routable network object. The function
+Building a transport network with r5r is as simple as a one command line:
 
 ```{r, message = FALSE}
-# Indicate the path where OSM and GTFS data are stored
 r5r_core <- setup_r5(data_path)
@@
 
-**r5** has two core routing functions.
+The function uses the `.pbf` and the `GTFS.zip` files in the directory pointed 
+by `data_path` to create a multimodal transport network used for routing by 
+R<sup>5</sup>. The resulting `network.dat`, as well as some other files used by 
+R<sup>5</sup>, is saved inside the supplied directory.
 
-- `detailed_itineraries`
-   * Returns a `data.frame sf LINESTRINGs` with one or multiple alternative routes
-   between one or multiple origin destination pairs. The data output brings 
-   detailed information on transport mode, travel time, walk distance etc for 
-   each trip section
+While the presence of a `.pbf` file is mandatory, since it is required to create 
+a street network, the `GTFS.zip` file is only used to calculate public transport 
+trips, thus may not be necessary if that's not intended. Multiple GTFS can be 
+used simultaneously, in which case all files are merged into a single one 
+automatically.
 
-- `travel_time_matrix`
-   * Fast function that returns a simple 'data.frame' with travel time 
-   estimates between one or multiple origin destination pairs.
+### Calculate a travel time matrix
 
-## Fast many to many travel time matrix
+Calculating a travel time matrix with r5r can be done very efficiently with the
+`travel_time_matrix()` function. The function takes travel restrictions and
+origins/destinations spatial location as inputs and outputs a `data.table` with 
+the travel time between all combinations of origins and destinations that can be 
+reached with the given restrictions:
 
 ```{r, message = FALSE}
+# set restrictions
 mode <- c("WALK", "TRANSIT")
-max_walk_dist <- Inf
-max_trip_duration <- 120L
+max_walk_dist <- 1000
+max_trip_duration <- 120
 departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
@@
 
-# estimate travel time matrix
+# calculate travel time matrix
 ttm <- travel_time_matrix(r5r_core,
@@
 
-head(ttm)
+ttm
 ```
 
-## Detailed itineraries
+Note that the resulting `ttm` object only holds the travel time of 1043 
+origin/destination pairs, even though the `points` data set passed both to 
+`origins` and `destinations` argument contains 100 points. This is because out 
+of the total 10000 possible trips between the supplied points, only 1043 could 
+be completed within the specified restrictions. Removing any walking distance 
+limits (by setting it to infinite) and increasing maximum trip duration 
+accordingly results in a "complete" output?
+
+```{r, message = FALSE}
+# set softer restrictions
+max_walk_dist <- Inf
+max_trip_duration <- 360
+
+# calculate travel time matrix
+ttm <- travel_time_matrix(r5r_core,
+                          origins = points,
+                          destinations = points,
+                          mode,
+                          departure_datetime,
+                          max_walk_dist,
+                          max_trip_duration)
+
+ttm
+```
+
+### Detailed itineraries
 

### 2020-09-18 r-package/paper2020/paper2020.Rmd
updates package download command in paper

@@
 
-The package can be installed with the following command:
+The package can be installed with the following commands:
 
 ```{r, message = FALSE, eval = FALSE}
+# stable version
+install.packages("r5r")
+
+# development version
 devtools::install_github("ipeaGIT/r5r", subdir = "r-package")
@@
 limits (by setting it to infinite) and increasing maximum trip duration 
-accordingly results in a "complete" output?
+accordingly results in a "complete" output:
 

### 2020-09-18 r-package/paper2020/paper2020.Rmd
updates detailed_itineraries paper section

@@
 
-### Detailed itineraries
+### Calculate detailed itineraries
 
-Allows one to get the fastest or multiple alternative routes between a set of
-origins and destinations.
+Often enough one wants more information on the itineraries between 
+origin/destination pairs than simply the travel time. For such cases, the 
+package includes the `detailed_itineraries()` function, which also takes travel 
+restrictions as inputs, but outputs a much more detailed data set at the trip 
+segment level. Another key feature of this function is that it (optionally) 
+provides multiple alternative routes between origin/destination pairs:
 
-Create simple example here and plot the result below
+```{r, message = FALSE}
+# load points of interests and set origin and destination
+points <- read.csv(file.path(data_path, "poa_points_of_interest.csv"))
+origins <- points[c(10, 14),]
+destinations <- points[c(12, 15),]
+
+# set restrictions
+mode <- c("WALK", "TRANSIT")
+max_walk_dist <- Inf
+max_trip_duration <- 120
+departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
+                                 format = "%d-%m-%Y %H:%M:%S")
+
+dit <- detailed_itineraries(r5r_core,
+                            origins,
+                            destinations,
+                            mode,
+                            departure_datetime,
+                            max_walk_dist,
+                            shortest_path = FALSE)
+
+dit
+```
+
+Since the `shortest_path` argument has been set to `FALSE`, `dit` holds 
+information for multiple itineraries between the same origin/destination pairs:
+
+```{r, message = FALSE}
+setDT(copy(dit))[, .(n_itineraries = uniqueN(option)), by = .(fromId, toId)]
+```
+
+Had `shortest_path` been set to `TRUE`, only the shortest path would have been 
+returned:
+
+```{r, message = FALSE}
+dit_short <- detailed_itineraries(r5r_core,
+                                  origins,
+                                  destinations,
+                                  mode,
+                                  departure_datetime,
+                                  max_walk_dist,
+                                  shortest_path = TRUE)
+
+setDT(copy(dit_short))[, .(n_itineraries = uniqueN(option)), by = .(fromId, toId)]
+```
 
 ### visualize results
+
+`detailed_itineraries()` output is a `sf` object, which means that each trip segment is associated to a geometry
 

### 2020-09-18 r-package/paper2020/paper2020.Rmd
updates visualization paper section

@@
 <!--  - https://joss.theoj.org/papers/10.21105/joss.01926 -->
-
-dasdasdds
 
@@
 
-`detailed_itineraries()` output is a `sf` object, which means that each trip segment is associated to a geometry
+`detailed_itineraries()` output is a `sf` object, which means that each trip 
+segment is associated to a spatial geometry that can be plotted with common `R` 
+packages, such as `ggplot2`. The package also offers a handy 
+`street_network_to_sf()` function that extracts the OSM street network used in 
+the routing as a `sf` object and can be used to provide geographic context for 
+the visualization.
 
-```{r, message = FALSE}
+```{r, message = FALSE, fig.width = 7, fig.height = 5}
 # extract OSM network
@@
 
-# plot
+dit$facet_labels <- paste0(dit$fromId, " to\n", dit$toId)
+
 ggplot() +
-   geom_sf(data = street_net$edges, color='gray85') +
-   theme_minimal()
-
+   geom_sf(data = street_net$edges, color = "gray85") +
+   geom_sf(data = dit, aes(color = mode), size = 1) +
+   facet_wrap(~ facet_labels, nrow = 1) +
+   theme_minimal() + 
+   theme(legend.position = "bottom", axis.text.x = element_text(angle = 30, hjust = 1))
 ```

### 2020-09-18 r-package/paper2020/paper2020.Rmd
updates visualization paper section

@@
 
-### visualize results
+### Visualize results
 
@@
 packages, such as `ggplot2`. The package also offers a handy 
-`street_network_to_sf()` function that extracts the OSM street network used in 
-the routing as a `sf` object and can be used to provide geographic context for 
-the visualization.
+`street_network_to_sf()` function which extracts the OSM street network used by 
+R<sup>5</sup> as a `sf` object and that can be used to provide geographic 
+context for the visualization.
 
@@
 
+The picture above can be used to compare the walking-only to the walking and 
+transit itineraries between each origin/destination pair. Given that the sample 
+`GTFS.zip` file includes only very few public transport routes, the resulting 
+plot picture some odd scenarios, such as a fairly lengthy circular-like bus 
+route, in the leftmost facet, and a very long walking access trip to the bus, 
+in the rightmost facet.
+
 # Acknowledgments

### 2020-09-26 r-package/paper2020/paper2020.Rmd
updates paper acknowledgments
additions to paper intro and methods

@@
 ---
-title: 'r5r: Rapid Realistic Routing with R5 in R'
+title: 'r5r: rapid realistic routing on multimodal transport networks with R5 in R'
 author: "Rafael H. M. Pereira, Marcus Saraiva, Daniel Herszenhut, Carlos Kaue Braga"
@@
 more broadly. Despite the long-standing importance of routing analysis to the 
-field, transport researchers and practitioners still frequently face practical 
-challenges with this task. These challenges involve for example the monetary 
-costs of licensed software, data availability issues for data hungry models, and
-the long computation times to run multiple routing scenarios, particularly in 
-large and complex multimodal transport networks. While there is a growing number
-of open source routing models [@opentripplanneropentripplanner; 
-@lovelace2019stplanr; @padgham2019dodgr], most options available only return the
-fastest route option between a set of origins and destinations. This simple
-approach overlooks how often people's actual route choices deviate from shortest
-travel time paths due to imperfect information, monetary costs and safety issues
-among other reasons [@larsen2011travel; @zhu2015people; @conway2019getting].
+field of transport and mobility studies, researchers and practitioners still 
+frequently face practical challenges with this task. These challenges involve for
+example the monetary costs of licensed software, data availability issues for data
+hungry models, and the long computation times to run multiple routing scenarios, 
+particularly in  large and complex multimodal transport networks.
+
+While there is a growing number of open source routing models [@opentripplanneropentripplanner; 
+@lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently 
+process large transport networks. More importantly, these softwares generate
+deterministic estimates of travel times without accounting for issues of service
+reliability (REF) and they only return the fastest route option  between a set of
+origins and destinations. This simple approach overlooks how often people's 
+actual route choices deviate from shortest travel time paths, what
+can occur due to multiple factors such as imperfect information, monetary costs 
+and safety issues [@larsen2011travel; @zhu2015people; @conway2019getting].
 
@@
 automatically combine OSM data with one or multiple public transport data sets 
-in the standard GTFS format to create a routable transport network.
+provided by the user in  the standard GTFS format to create a routable transport
+network.
 
@@
 - `travel_time_matrix()`: used to compute travel time estimates between one or 
-multiple origin/destination pairs. The origin/destination input can be either a 
+multiple origin/destination pairs for a single departure time or for multiple
+departure times over a time window. The origin/destination input can be either a 
 spatial `sf POINT` object, or a `data.frame` containing the columns `id`, `lon`, 
-`lat.`
+`lat.` The output is a `data.frame` with the travel time estimate between each 
+origin/destination pair. If the user passes a `time_window` parameter to the,
+function, r5r will calculate multiple travel time matrices departing each minute
+and return additional information on the percentile distribution of travel times
+that allows one to grasp service reliability issues.
 
-- `detailed_itineraries()`:  used to compute detailed information at the trip 
-segment level on either the shortest or multiple alternative routes between one
-or multiple origin/destination pairs. Includes information such as the transport 
-mode, waiting time, travel time and distance.
+- `detailed_itineraries()`: used to compute detailed information on either the
+shortest or multiple alternative routes between one or multiple origin/destination
+pairs for a single departure time. The output includes detailed information on
+route alternatives such as the transport mode, waiting time, travel time and 
+distance of each segment of the trip.
 
@@
 
-Conveyal for R5
+The [R<sup>5</sup> routing engine](https://github.com/conveyal/r5) is developed 
+at [Conveyal](https://www.conveyal.com/) with contributions from several people. 
+This work was supported by the Brazilian Institute for Applied Economic Research 
+(Ipea).
 

### 2020-09-26 r-package/paper2020/paper2020.Rmd
comment to paper

@@
 
-The package can be installed with the following commands:
+The package can be installed with the following command:
 
@@
 
-# development version
-devtools::install_github("ipeaGIT/r5r", subdir = "r-package")
 ```
@@
 
+#> I think this is perhaps unecessary detail for the paper publication.
 Please note that `R` only allocates memory to Java once during a running 
@@
 
-Building a transport network with r5r is as simple as a one command line:
+To build a routable transport network with r5r, the user only needs one command
+line pointing to `setup_r5()` where the input data is stored:
 
@@
 by `data_path` to create a multimodal transport network used for routing by 
-R<sup>5</sup>. The resulting `network.dat`, as well as some other files used by 
-R<sup>5</sup>, is saved inside the supplied directory.
+R<sup>5</sup>. The resulting `network.dat` as well as some other files used by 
+R<sup>5</sup> are saved inside the supplied directory.
 
-While the presence of a `.pbf` file is mandatory, since it is required to create 
-a street network, the `GTFS.zip` file is only used to calculate public transport 
-trips, thus may not be necessary if that's not intended. Multiple GTFS can be 
-used simultaneously, in which case all files are merged into a single one 
-automatically.
+The presence of a `.pbf` file with OpenStreetMap data is mandatory because it is
+required to create a street network. Meanwhile, a `GTFS.zip` file is only used to
+calculate public transport trips, thus may not be necessary if that's not intended.
+Multiple GTFS can be used simultaneously, in which case all files are merged into
+a single one automatically by R<sup>5</sup>.
 
@@
 Calculating a travel time matrix with r5r can be done very efficiently with the
-`travel_time_matrix()` function. The function takes travel restrictions and
+`travel_time_matrix()` function. The function takes travel parameters and
 origins/destinations spatial location as inputs and outputs a `data.table` with 
 the travel time between all combinations of origins and destinations that can be 
-reached with the given restrictions:
+reached with the given parameters:
 
 ```{r, message = FALSE}
-# set restrictions
-mode <- c("WALK", "TRANSIT")
-max_walk_dist <- 1000
-max_trip_duration <- 120
 departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
@@
                           destinations = points,
-                          mode,
+                          mode = c("WALK", "TRANSIT"),
                           departure_datetime,
-                          max_walk_dist,
-                          max_trip_duration)
+                          max_walk_dist = 1000,
+                          max_trip_duration = 120)
 
-ttm
+head(ttm)
 ```

### 2020-10-03 r-package/R/setup_r5.R
moving poa sample data and updating code. Closes #116

@@ -23,7 +23,7 @@
 #' library(r5r)
 #'
 #' # directory with street network and gtfs files
-#' path <- system.file("extdata", package = "r5r")
+#' path <- system.file("extdata/poa", package = "r5r")
 #'
 #' r5r_core <- setup_r5(data_path = path, temp_dir = TRUE)
 #' }

### 2020-10-03 r-package/R/stop_r5.R

@@ -8,7 +8,7 @@
 #' @examples
 #' \donttest{library(r5r)
 #'
-#' path <- system.file("extdata", package = "r5r")
+#' path <- system.file("extdata/poa", package = "r5r")
 #'
 #' r5r_core <- setup_r5(data_path = path)
 #'

### 2020-10-03 r-package/R/street_network_to_sf.R

@@ -13,7 +13,7 @@
 #' library(r5r)
 #'
 #' # build transport network
-#' path <- system.file("extdata", package = "r5r")
+#' path <- system.file("extdata/poa", package = "r5r")
 #' r5r_core <- setup_r5(data_path = path)
 #'
 #' # load origin/destination points

### 2020-10-03 r-package/R/travel_time_matrix.R

@@ -57,11 +57,11 @@
 #' library(r5r)
 #'
 #' # build transport network
-#' data_path <- system.file("extdata", package = "r5r")
+#' data_path <- system.file("extdata/spo", package = "r5r")
 #' r5r_core <- setup_r5(data_path = data_path)
 #'
 #' # load origin/destination points
-#' points <- read.csv(system.file("extdata/poa_hexgrid.csv", package = "r5r"))[1:5,]
+#' points <- read.csv(file.path(data_path, "spo_hexgrid.csv"))[1:5,]
 #'
 #' departure_datetime <- as.POSIXct("13-03-2019 14:00:00", format = "%d-%m-%Y %H:%M:%S")
 #'

### 2020-10-03 r-package/tests/testthat/test-detailed_itineraries.R

@@ -5,7 +5,7 @@ testthat::skip_on_cran()
 
 # load required data and setup r5r_core
 
-data_path <- system.file("extdata", package = "r5r")
+data_path <- system.file("extdata/poa", package = "r5r")
 r5r_core <- setup_r5(data_path, verbose = FALSE)
 points <- read.csv(file.path(data_path, "poa_points_of_interest.csv"))
 

### 2020-10-03 r-package/tests/testthat/test-stop_r5.R

@@ -3,7 +3,7 @@ context("Stop r5r core")
 testthat::skip_on_cran()
 
 test_that("stop_r5 raises warnings and errors when supplied wrong objects", {
-  path <- system.file("extdata", package = "r5r")
+  path <- system.file("extdata/poa", package = "r5r")
   old_env <- environment()
   r5r_core <- setup_r5(data_path = path, verbose=FALSE)
 
@@ -19,7 +19,7 @@ test_that("stop_r5 raises warnings and errors when supplied wrong objects", {
 })
 
 test_that("stop_r5 successfully stops multiple running r5r cores", {
-  path <- system.file("extdata", package = "r5r")
+  path <- system.file("extdata/poa", package = "r5r")
   old_env <- environment()
 
   # stops all running r5r cores

### 2020-10-03 r-package/tests/testthat/test-travel_time_matrix.R

@@ -5,9 +5,9 @@ testthat::skip_on_cran()
 
 # load required data and setup r5r_core
 
-data_path <- system.file("extdata", package = "r5r")
+data_path <- system.file("extdata/spo", package = "r5r")
 r5r_core <- setup_r5(data_path, verbose = FALSE)
-points <- read.csv(file.path(data_path, "poa_hexgrid.csv"))
+points <- read.csv(file.path(data_path, "spo_hexgrid.csv"))
 
 # create testing function
 

### 2020-10-03 r-package/tests/testthat/test-utils.R

@@ -3,7 +3,7 @@ context("Support functions")
 testthat::skip_on_cran()
 
 # load required data and setup r5r_obj
-data_path <- system.file("extdata", package = "r5r")
+data_path <- system.file("extdata/poa", package = "r5r")
 r5r_obj <- setup_r5(data_path, verbose = FALSE)
 points <- read.csv(file.path(data_path, "poa_points_of_interest.csv"))
 

### 2020-10-03 r-package/tests_rafa/test_rafa.R

@@ -13,17 +13,13 @@ library(mapview)
 library(covr)
 library(testthat)
 library(ggplot2)
-library(mapview)
 library(checkmate)
 library(geobr)
-
-
-devtools::build_vignettes()
-
+library(gtfs2gps)
+library(mapview)
 mapviewOptions(platform = 'leafgl')
 
-# Update documentation
-devtools::document(pkg = ".")
+
 
 
 ##### INPUT  ------------------------
@@ -32,13 +28,14 @@ devtools::document(pkg = ".")
 
 
 # build transport network
-path <- system.file("extdata", package = "r5r")
+poa <- system.file("extdata/poa", package = "r5r")
+spo <- system.file("extdata/spo", package = "r5r")
 
 # r5r::download_r5(force_update = T)
 r5r_core <- setup_r5(data_path = path, verbose = F)
 
 # load origin/destination points
-points <- read.csv(system.file("extdata/poa_hexgrid.csv", package = "r5r"))
+points <- read.csv(system.file("extdata/poa/poa_hexgrid.csv", package = "r5r"))
 points_sf <- sfheaders::sf_multipoint(points, x='lon', y='lat', multipoint_id = 'id')
 
 # remove files
@@ -72,24 +69,29 @@ get_all_od_combinations <- function(origins, destinations){
         }
 
 # example
-origins <- read.csv(system.file("extdata/poa_hexgrid.csv", package = "r5r"))[1:800,]
-destinations <- read.csv(system.file("extdata/poa_hexgrid.csv", package = "r5r"))[400:1200,]
+origins <- read.csv(system.file("extdata/poa/poa_hexgrid.csv", package = "r5r"))[1:800,]
+destinations <- read.csv(system.file("extdata/poa/poa_hexgrid.csv", package = "r5r"))[400:1200,]
 
 df <- get_all_od_combinations(origins, destinations)
 
 ##### TESTS street_network_to_sf ------------------------
 
+gtfs_shapes <- gtfs2gps::read_gtfs( system.file("extdata/spo/spo.zip", package = "r5r") ) %>%
+        gtfs2gps::gtfs_shapes_as_sf()
+
+
+
+spo <- system.file("extdata/spo", package = "r5r")
+
+r5r_core <- setup_r5(data_path = spo, verbose = F)
+
 street_net <- street_network_to_sf(r5r_core)
 
 
-mapview(street_net$edges ) + street_net$vertices + a
[…]

### 2020-10-21 r-package/paper2020/paper2020.Rmd
adds travel time matrix example using time window to paper

@@
 multiple origin/destination pairs for a single departure time or for multiple
-departure times over a time window. The origin/destination input can be either a 
-spatial `sf POINT` object, or a `data.frame` containing the columns `id`, `lon`, 
-`lat.` The output is a `data.frame` with the travel time estimate between each 
-origin/destination pair. If the user passes a `time_window` parameter to the,
-function, r5r will calculate multiple travel time matrices departing each minute
-and return additional information on the percentile distribution of travel times
+departure times over a time window. The output comprises the travel time between
+each origin/destination pair. If the user passes a `time_window` parameter
+to the function, r5r will calculate multiple travel time matrices departing each 
+minute and return additional information on the percentile distribution of travel times
 that allows one to grasp service reliability issues.
@@
 install.packages("r5r")
-
 ```
@@
 
-#> I think this is perhaps unecessary detail for the paper publication.
-Please note that `R` only allocates memory to Java once during a running 
-session. If a Java process has been initiated before, either by r5r or any other 
-Java-dependent package or function, it might be required to restart the current 
-`R` session in order to effectively change the amount of memory allocated.
-
 The package can be attached (with some other packages used in this article) with 
@@
 To illustrate functionality, the package includes a small sample data set for 
-the city of Porto Alegre (Brazil). It includes three files:
+the city of Porto Alegre (Brazil). It includes four files:
 
@@
 - The spatial coordinates of points covering the area in `.csv` format.
+- The spatial coordinates of some of the city's points of interest.
 
 ```{r}
-data_path <- system.file("extdata", package = "r5r")
+data_path <- system.file("extdata/poa", package = "r5r")
 list.files(data_path)
 
-points <- fread(system.file("extdata/poa_hexgrid.csv", package = "r5r"))
+points <- fread(system.file("extdata/poa/poa_hexgrid.csv", package = "r5r"))
 points <- points[c(sample(1:nrow(points), 100))]
@@
 Calculating a travel time matrix with r5r can be done very efficiently with the
-`travel_time_matrix()` function. The function takes travel parameters and
-origins/destinations spatial location as inputs and outputs a `data.table` with 
-the travel time between all combinations of origins and destinations that can be 
-reached with the given parameters:
+`travel_time_matrix()` function. The function takes origins/destinations spatial
+location (either as a spatial `sf POINT` object, or as a `data.frame` containing
+the columns `id`, `lon` and `lat`) and travel parameters as inputs and outputs a
+`data.table` with the travel time estimates between all combinations of origins 
+and destinations that can be reached with the given parameters.
+
+The `time_window` parameter is set to 1 by default, which means that travel time 
+estimates will be calculated using a single departure time (specified by 
+`departure_datetime`):
 
 ```{r, message = FALSE}
+mode <- c("WALK", "TRANSIT")
+max_walk_dist <- 1000
+max_trip_duration <- 120
 departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
@@
                           destinations = points,
-                          mode = c("WALK", "TRANSIT"),
-                          departure_datetime,
-                          max_walk_dist = 1000,
-                          max_trip_duration = 120)
+                          mode = mode,
+                          departure_datetime = departure_datetime,
+                          max_walk_dist = max_walk_dist,
+                          max_trip_duration = max_trip_duration)
 
@@
 
-Note that the resulting `ttm` object only holds the travel time of 1043 
-origin/destination pairs, even though the `points` data set passed both to 
-`origins` and `destinations` argument contains 100 points. This is because out 
-of the total 10000 possible trips between the supplied points, only 1043 could 
-be completed within the specified restrictions. Removing any walking distance 
-limits (by setting it to infinite) and increasing maximum trip duration 
-accordingly results in a "complete" output:
+Alternatively, a bigger time window can be specified. In such cases several 
+travel time matrices are calculated using every minute from `departure_datetime` 
+to the end of the given window as a distinct departure time. By default the 
+function outputs the 50th percentile (i.e. the median) travel time between each 
+origin and destination pair, which can be overruled by the `percentiles` 
+parameter:
 
 ```{r, message = FALSE}
-# set softer restrictions
-max_walk_dist <- Inf
-max_trip_duration <- 360
+time_window <- 60
+percentiles <- c(25, 50, 75)
 
@@
                           destinations = points,
-                          mode,
-                          departure_datetime,
-                          max_walk_dist,
-                          max_trip_duration)
+                          mode = mode,
+                          departure_datetime = departure_datetime,
+                          max_walk_dist = max_walk_dist,
+                          max_trip_duration = max_trip_duration,
+                          time_window = time_window,
+                          percentiles = percentiles)
 
@@
 ```
+
+Since the sample data holds information for few transit routes, many trips rely 
+only on walking to be completed. Consequently, the travel time estimates of such 
+trips are not affected by transit reliability issues. This is the case of the 
+first five rows of `ttm`, which present the same estimate regardless of the 
+percentile they represent. The effect of the `time_window` and `percentiles` 
+parameters can be observed above in the last rows of `ttm`, in which the 
+estimates grow along the increase of the percentile they depict.
 

### 2020-10-22 r-package/paper2020/paper2020.Rmd
paper updates and edits

@@
 r5r has low data requirements, enabling routing analysis on different contexts 
-wherever there is open data on street networks from OpenStreetMap (OSM). The 
+wherever data on street networks from OpenStreetMap (OSM) is available. The 
 package is also easily scalable allowing for fast computations at either the 
@@
 
-- `travel_time_matrix()`: used to compute travel time estimates between one or 
+- `travel_time_matrix()`: computes travel time estimates between one or 
 multiple origin/destination pairs for a single departure time or for multiple
-departure times over a time window. The output comprises the travel time between
-each origin/destination pair. If the user passes a `time_window` parameter
-to the function, r5r will calculate multiple travel time matrices departing each 
-minute and return additional information on the percentile distribution of travel times
-that allows one to grasp service reliability issues.
+departure times over a time window.
+The travel_time_matrix function uses an R5-specific extension to the RAPTOR routing algorithm developed by Conway et al.
+[@conway2017evidencebased], which provides an efficient and systematic sample of
+multiple departures per minute over a `time_window` set by the user. If the user
+passes a `time_window` parameter to the function, r5r will calculate multiple travel 
+time matrices departing each minute and return additional information on the percentile distribution of travel times that allows one to grasp service reliability issues.
 
-- `detailed_itineraries()`: used to compute detailed information on either the
+- `detailed_itineraries()`: computes detailed information on either the
 shortest or multiple alternative routes between one or multiple origin/destination
@@
 route alternatives such as the transport mode, waiting time, travel time and 
-distance of each segment of the trip.
+distance of each segment of the trip. This function uses an R5-specific extension
+to the McRAPTOR routing algorithm to find paths that are optimal or less than
+optimal. The specific extension to McRAPTOR to do suboptimal path routing are not documented yet, but a detailed description of base McRAPTOR can be found in Delling et al (2015).
+
+#> incluir essa ref
+#> - Delling, D., Pajor, T., & Werneck, R. F. (2015). Round-based public transit routing. Transportation Science, 49(3), 591-604.
 
 Both routing functions are versatile so users can easily set customized inputs
-such as transport mode, departure date and times, walking and cycling speeds, 
+such as transport modes, departure date and times, walking and cycling speeds, 
 maximum trip duration, walking distances and number of public transport 
@@
 The first step when using r5r is to set the amount of memory available to Java. 
-By default `R` allocates only 512 MB of memory for Java processes, which is not 
-enough for large queries using r5r. To increase available memory to 2GB, for 
+By default `R` allocates only 512 MB of memory for Java processes~~, which is not 
+enough for large queries using r5r~~. To increase available memory to 2GB, for 
 example, we need to set the `java.parameters` option at the beginning of the 
@@
 
+```{r, include = FALSE}
+
+# point to directory with data
+data_path <- system.file("extdata/spo", package = "r5r")
+
+```
+
 ```{r}
-data_path <- system.file("extdata/poa", package = "r5r")
+# list files in directory
 list.files(data_path)
 
-points <- fread(system.file("extdata/poa/poa_hexgrid.csv", package = "r5r"))
+# read points of origin and destination
+points <- fread(file.path(data_path, "spo_hexgrid.csv"))
 points <- points[c(sample(1:nrow(points), 100))]
@@
 To build a routable transport network with r5r, the user only needs one command
-line pointing to `setup_r5()` where the input data is stored:
+line pointing to where the input data is stored:
 
@@
 calculate public transport trips, thus may not be necessary if that's not intended.
-Multiple GTFS can be used simultaneously, in which case all files are merged into
-a single one automatically by R<sup>5</sup>.
+Multiple GTFS feeds can be used simultaneously, in which case R<sup>5</sup> will
+merge all files into a single network automatically.
 
@@
 
-Had `shortest_path` been set to `TRUE`, only the shortest path would have been 
-returned:
-
-```{r, message = FALSE}
-dit_short <- detailed_itineraries(r5r_core,
-                                  origins,
-                                  destinations,
-                                  mode,
-                                  departure_datetime,
-                                  max_walk_dist,
-                                  shortest_path = TRUE)
-
-setDT(copy(dit_short))[, .(n_itineraries = uniqueN(option)), by = .(fromId, toId)]
-```
-
 ### Visualize results
@@
 The [R<sup>5</sup> routing engine](https://github.com/conveyal/r5) is developed 
-at [Conveyal](https://www.conveyal.com/) with contributions from several people. 
+at [Conveyal](https://www.conveyal.com/) with contributions from several developers. 
 This work was supported by the Brazilian Institute for Applied Economic Research 

### 2020-10-22 r-package/paper2020/paper2020.Rmd
adds paper abstract

@@
   github_document: default
-abstract: "write abstract"
+abstract: "Routing is a key step in transport planning.
+Nonetheless, researchers and practitioners still face challenges regarding this task, such as the cost of licensed software and the long computation times to run complex scenarios.
+Conveyal's R<sup>5</sup> is a multimodal transport network router that offers meaningful planning features, such as accounting for service reliability issues and returning multiple travel time estimates and itineraries for a single origin/destination pair.
+This paper describes r5r, an open source R package built on top of R<sup>5</sup> that allow users to efficiently compute travel time matrices and generate detailed itineraries between a set of origins and destinations at no expense.
+The development of this package contributes to the transport field as it offers an easy-to-use powerful tool for robust planning and research."
 urlcolor: blue
@@
 
-```{r, include = FALSE}
-
-# point to directory with data
+```{r}
+# points to directory with data
 data_path <- system.file("extdata/spo", package = "r5r")
 
-```
-
-```{r}
-# list files in directory
 list.files(data_path)

### 2020-10-22 r-package/paper2020/paper2020.Rmd
updates paper references

@@
 
-The r5r package has 3 fundamental functions. The `setup_r5()` function builds a 
-multimodal transport network used for routing in R<sup>5</sup>. This function 
-(1) downloads/updates a compiled R<sup>5</sup> JAR file, tailored for the 
-purposes of the package, and stores it locally in the r5r package directory for 
-future use; and (2) combines the OSM and GTFS data sets to build a routable 
-network object. For fast routing analysis, r5r currently has two functions:
+The r5r package has 3 fundamental functions:
+
+- `setup_r5()`: builds a multimodal transport network used for routing in R<sup>5</sup>.
+This function (1) downloads/updates a compiled R<sup>5</sup> JAR file, tailored for the purposes of the package, and stores it locally in the r5r package directory for future use; and (2) combines the OSM and GTFS data sets to build a routable network object.
 
@@
 departure times over a time window.
-The travel_time_matrix function uses an R5-specific extension to the RAPTOR routing algorithm developed by Conway et al.
+This function uses an R5-specific extension to the RAPTOR routing algorithm developed by Conway et al.
 [@conway2017evidencebased], which provides an efficient and systematic sample of
@@
 to the McRAPTOR routing algorithm to find paths that are optimal or less than
-optimal. The specific extension to McRAPTOR to do suboptimal path routing are not documented yet, but a detailed description of base McRAPTOR can be found in Delling et al (2015).
-
-#> incluir essa ref
-#> - Delling, D., Pajor, T., & Werneck, R. F. (2015). Round-based public transit routing. Transportation Science, 49(3), 591-604.
+optimal. The specific extension to McRAPTOR to do suboptimal path routing are not documented yet, but a detailed description of base McRAPTOR can be found in @delling2015roundbased.
 
@@
 To illustrate functionality, the package includes a small sample data set for 
-the city of Porto Alegre (Brazil). It includes four files:
+the city of São Paulo (Brazil). It includes three files:
 
@@
 - The spatial coordinates of points covering the area in `.csv` format.
-- The spatial coordinates of some of the city's points of interest.
 
@@
 ```{r, message = FALSE}
+data_path <- system.file("extdata/poa", package = "r5r")
+
+r5r_core <- setup_r5(data_path)
+
 # load points of interests and set origin and destination
-points <- read.csv(file.path(data_path, "poa_points_of_interest.csv"))
+points <- fread(file.path(data_path, "poa_points_of_interest.csv"))
 origins <- points[c(10, 14),]
@@
 dit <- detailed_itineraries(r5r_core,
-                            origins,
-                            destinations,
-                            mode,
-                            departure_datetime,
-                            max_walk_dist,
+                            origins = origins,
+                            destinations = destinations,
+                            mode = mode,
+                            departure_datetime = departure_datetime,
+                            max_walk_dist = max_walk_dist,
+                            max_trip_duration = max_trip_duration,
                             shortest_path = FALSE)

### 2020-10-22 r-package/paper2020/paper2020.Rmd
fixes paper detailed itineraries example

@@
 
-- `setup_r5()`: builds a multimodal transport network used for routing in R<sup>5</sup>.
-This function (1) downloads/updates a compiled R<sup>5</sup> JAR file, tailored for the purposes of the package, and stores it locally in the r5r package directory for future use; and (2) combines the OSM and GTFS data sets to build a routable network object.
+- `setup_r5()`: builds a multimodal transport network used for routing in 
+R<sup>5</sup>. This function (1) downloads/updates a compiled R<sup>5</sup> JAR 
+file, tailored for the purposes of the package, and stores it locally in the r5r 
+package directory for future use; and (2) combines the OSM and GTFS data sets to 
+build a routable network object.
 
@@
 multiple origin/destination pairs for a single departure time or for multiple
-departure times over a time window.
-This function uses an R5-specific extension to the RAPTOR routing algorithm developed by Conway et al.
-[@conway2017evidencebased], which provides an efficient and systematic sample of
-multiple departures per minute over a `time_window` set by the user. If the user
-passes a `time_window` parameter to the function, r5r will calculate multiple travel 
-time matrices departing each minute and return additional information on the percentile distribution of travel times that allows one to grasp service reliability issues.
+departure times over a time window. This function uses an R5-specific extension 
+to the RAPTOR routing algorithm which provides an efficient and systematic 
+sample of multiple departures per minute over a `time_window` set by the user 
+[@conway2017evidencebased]. If the user passes a `time_window` parameter to the 
+function, r5r calculates multiple travel time matrices departing each minute and 
+return additional information on the percentile distribution of travel times 
+that allows one to grasp service reliability issues.
 
@@
 route alternatives such as the transport mode, waiting time, travel time and 
-distance of each segment of the trip. This function uses an R5-specific extension
-to the McRAPTOR routing algorithm to find paths that are optimal or less than
-optimal. The specific extension to McRAPTOR to do suboptimal path routing are not documented yet, but a detailed description of base McRAPTOR can be found in @delling2015roundbased.
+distance of each segment of the trip. This function uses an R5-specific 
+extension to the McRAPTOR routing algorithm to find optimal or less than optimal 
+paths. The specific extension to McRAPTOR to do suboptimal path routing are not 
+documented yet, but a detailed description of base McRAPTOR can be found in 
+@delling2015roundbased.
 
@@
 package includes the `detailed_itineraries()` function, which also takes travel 
-restrictions as inputs, but outputs a much more detailed data set at the trip 
+restrictions as inputs, but outputs a much more detailed data table at the trip 
 segment level. Another key feature of this function is that it (optionally) 
-provides multiple alternative routes between origin/destination pairs:
+provides multiple alternative routes between origin/destination pairs. The 
+algorithm doesn't yet support frequency-based trips (such as São Paulo's), 
+however, so we use another data sample included in the package that covers the 
+city of Porto Alegre (Brazil):
 
@@
 The picture above can be used to compare the walking-only to the walking and 
-transit itineraries between each origin/destination pair. Given that the sample 
-`GTFS.zip` file includes only very few public transport routes, the resulting 
-plot picture some odd scenarios, such as a fairly lengthy circular-like bus 
-route, in the leftmost facet, and a very long walking access trip to the bus, 
-in the rightmost facet.
+transit itineraries between each origin/destination pair. Given that Porto 
+Alegre's sample `GTFS.zip` file includes only very few public transport routes, 
+the resulting plot picture some odd scenarios, such as a fairly lengthy 
+circular-like bus route, in the leftmost facet, and a very long walking access 
+trip to the bus, in the rightmost facet.
 

### 2020-10-22 r-package/paper2020/paper2020.Rmd
removes paragraph from paper

@@
 
-Since the sample data holds information for few transit routes, many trips rely 
-only on walking to be completed. Consequently, the travel time estimates of such 
-trips are not affected by transit reliability issues. This is the case of the 
-first five rows of `ttm`, which present the same estimate regardless of the 
-percentile they represent. The effect of the `time_window` and `percentiles` 
-parameters can be observed above in the last rows of `ttm`, in which the 
-estimates grow along the increase of the percentile they depict.
-
 ### Calculate detailed itineraries

### 2020-10-23 r-package/R/detailed_itineraries.R
updates poa's gtfs to include more routes

@@ -69,7 +69,7 @@
 #' points <- read.csv(file.path(data_path, "poa_points_of_interest.csv"))
 #'
 #' # inputs
-#' departure_datetime <- as.POSIXct("13-03-2019 14:00:00", format = "%d-%m-%Y %H:%M:%S")
+#' departure_datetime <- as.POSIXct("13-05-2019 14:00:00", format = "%d-%m-%Y %H:%M:%S")
 #'
 #' dit <- detailed_itineraries(r5r_core,
 #'                             origins = points[10,],
@@ -78,6 +78,8 @@
 #'                             departure_datetime = departure_datetime,
 #'                             max_walk_dist = 1000,
 #'                             max_trip_duration = 120L)
+#'
+#' stop_r5(r5r_core)
 #' }
 #' @export
 

### 2020-10-23 r-package/R/street_network_to_sf.R

@@ -18,6 +18,8 @@
 #'
 #' # load origin/destination points
 #' street_net <- street_network_to_sf(r5r_core)
+#'
+#' stop_r5(r5r_core)
 #' }
 #' @export
 

### 2020-10-23 r-package/tests/testthat/test-detailed_itineraries.R

@@ -15,7 +15,7 @@ default_tester <- function(r5r_core,
                            origins = points[1:2, ],
                            destinations = points[2:1, ],
                            mode = c("WALK", "TRANSIT"),
-                           departure_datetime = as.POSIXct("13-03-2019 14:00:00",
+                           departure_datetime = as.POSIXct("13-05-2019 14:00:00",
                                                            format = "%d-%m-%Y %H:%M:%S"),
                            max_walk_dist = Inf,
                            max_trip_duration = 120L,
@@ -92,9 +92,9 @@ test_that("detailed_itineraries adequately raises errors", {
   expect_error(default_tester(r5r_core, mode = "all"))
 
   # errors related to date formatting
-  numeric_datetime <- as.numeric(as.POSIXct("13-03-2019 14:00:00", format = "%d-%m-%Y %H:%M:%S"))
+  numeric_datetime <- as.numeric(as.POSIXct("13-05-2019 14:00:00", format = "%d-%m-%Y %H:%M:%S"))
 
-  expect_error(default_tester(r5r_core, departure_datetime = "13-03-2019 14:00:00"))
+  expect_error(default_tester(r5r_core, departure_datetime = "13-05-2019 14:00:00"))
   expect_error(default_tester(r5r_core, numeric_datetime))
 
   # errors related to max_walk_dist

### 2020-10-26 r-package/paper2020/paper2020.Rmd
back at using poa data in paper

@@
 To illustrate functionality, the package includes a small sample data set for 
-the city of São Paulo (Brazil). It includes three files:
+the city of Porto Alegre (Brazil). It includes three files:
 
@@
 - The spatial coordinates of points covering the area in `.csv` format.
+- The spatial coordinates of some of the city's points of interest.
 
@@
 # points to directory with data
-data_path <- system.file("extdata/spo", package = "r5r")
+data_path <- system.file("extdata/poa", package = "r5r")
 
@@
 # read points of origin and destination
-points <- fread(file.path(data_path, "spo_hexgrid.csv"))
+points <- fread(file.path(data_path, "poa_hexgrid.csv"))
 points <- points[c(sample(1:nrow(points), 100))]
@@
 segment level. Another key feature of this function is that it (optionally) 
-provides multiple alternative routes between origin/destination pairs. The 
-algorithm doesn't yet support frequency-based trips (such as São Paulo's), 
-however, so we use another data sample included in the package that covers the 
-city of Porto Alegre (Brazil):
+provides multiple alternative routes between origin/destination pairs:
 
 ```{r, message = FALSE}
-data_path <- system.file("extdata/poa", package = "r5r")
-
-r5r_core <- setup_r5(data_path)
-
 # load points of interests and set origin and destination

### 2020-10-26 r-package/paper2020/paper2020.Rmd
adds plot showing travel time distribution to paper

@@
   collapse = TRUE,
-  comment = "#>"
+  comment = "#>",
+  fig.width = 7, 
+  fig.height = 5
 )
@@
 The first step when using r5r is to set the amount of memory available to Java. 
-By default `R` allocates only 512 MB of memory for Java processes~~, which is not 
-enough for large queries using r5r~~. To increase available memory to 2GB, for 
-example, we need to set the `java.parameters` option at the beginning of the 
-script as follows:
+By default `R` allocates only 512 MB of memory for Java processes. To increase 
+available memory to 2GB, for example, we need to set the `java.parameters` 
+option at the beginning of the script as follows:
 
@@
 ```{r, message = FALSE}
-time_window <- 60
+time_window <- 120
 percentiles <- c(25, 50, 75)
@@
 ```
+
+The image below depicts the travel time distribution for each origin/destination in which a trip could be completed within the specified parameters.
+The pairs are ordered in the x axis by their median travel time and each point in the plot area pictures a distinct distribution percentile.
+
+```{r, message = FALSE}
+order_data <- ttm[, .(fromId, toId, travel_time_p025, travel_time_p050, travel_time_p075)]
+order_data <- order_data[order(travel_time_p050, travel_time_p025, travel_time_p075)]
+id_order   <- order_data[, id := paste0(fromId, "_", toId)]$id
+
+pctls <- paste0("travel_time_p", formatC(percentiles, width = 3, flag = "0"))
+
+plot_data <- melt(ttm, measure = pctls, variable = "percentil", value = "travel_time")
+plot_data[, percentil := as.integer(regmatches(percentil, regexpr("\\d+", percentil)))]
+plot_data[, travel_time := fifelse(travel_time > max_trip_duration, NA_integer_, travel_time)]
+plot_data[, id := factor(paste0(fromId, "_", toId), levels = id_order)]
+
+ggplot(plot_data) + 
+   geom_point(aes(id, travel_time, color = percentil), alpha = 0.1, size = 1) + 
+   theme_minimal() +
+   theme(axis.title.x = element_blank(), axis.text.x = element_blank())
+```
+
+Notice that the leftmost origin/destination pairs do not exhibit any variation in the distribution of their travel time along the time window.
+This is because the trips between such pairs rely solely on walking to be completed, therefore are not susceptible to transit reliability issues.
 
@@
 
-```{r, message = FALSE, fig.width = 7, fig.height = 5}
+```{r, message = FALSE}
 # extract OSM network

### 2020-10-26 r-package/paper2020/paper2020.Rmd
changes paper output formats

@@
 output:
-  rmarkdown::html_vignette: default
-  github_document: default
+  html_document: default
+  pdf_document: default
 abstract: "Routing is a key step in transport planning.
@@
 urlcolor: blue
-vignette: |
-  %\VignetteIndexEntry{r5r-intro} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
 bibliography: references.json
@@
 paths. The specific extension to McRAPTOR to do suboptimal path routing are not 
-documented yet, but a detailed description of base McRAPTOR can be found in 
-@delling2015roundbased.
+documented yet, but a detailed description of base McRAPTOR can be found in @delling2015roundbased.
 
@@
 ```{r, message = FALSE}
+percentiles <- 1:100
+
+ttm <- travel_time_matrix(r5r_core,
+                          origins = points,
+                          destinations = points,
+                          mode = mode,
+                          departure_datetime = departure_datetime,
+                          max_walk_dist = max_walk_dist,
+                          max_trip_duration = max_trip_duration,
+                          time_window = time_window,
+                          percentiles = percentiles)
+
 order_data <- ttm[, .(fromId, toId, travel_time_p025, travel_time_p050, travel_time_p075)]
@@
 ggplot(plot_data) + 
-   geom_point(aes(id, travel_time, color = percentil), alpha = 0.1, size = 1) + 
+   geom_point(aes(id, travel_time), alpha = 0.02, size = 1) + 
    theme_minimal() +
@@
 
-Notice that the leftmost origin/destination pairs do not exhibit any variation in the distribution of their travel time along the time window.
+Note that the leftmost origin/destination pairs do not exhibit any variation in the distribution of their travel time along the time window.
 This is because the trips between such pairs rely solely on walking to be completed, therefore are not susceptible to transit reliability issues.

### 2020-10-28 r-package/paper2020/paper2020.Rmd
tweaks paper's introduction second paragraph

@@
 process large transport networks. More importantly, these softwares generate
-deterministic estimates of travel times without accounting for issues of service
-reliability (REF) and they only return the fastest route option between a set of
-origins and destinations. This simple approach overlooks how often people's 
+deterministic estimates of travel times without accounting for variability in 
+results (due to service reliability issues, for example) and only return the 
+fastest route option between a set of origins and destinations (see @conway2018accounting 
+for a more in-depth discussion). This simple approach overlooks how often people's 
 actual route choices deviate from shortest travel time paths, what

### 2020-11-06 r-package/paper2020/paper2020.Rmd
General review of r5r paper.

--- 2020-10-26 r-package/paper2020/paper2020.Rmd
@@
 <!-- Examples of similar papers -->
+
 <!--  - https://transportfindings.org/article/8416-accessibility-toolbox-for-r-and-arcgis -->
@@
 
-Transport routing is the process of finding the alternative routes and costs 
-that connect places on a given transport network. Routing often involves complex
-mathematical algorithms and it is a fundamental step required for transport
-accessibility analysis, fleet allocation and transport simulation and planning
-more broadly. Despite the long-standing importance of routing analysis to the 
-field of transport and mobility studies, researchers and practitioners still 
-frequently face practical challenges with this task. These challenges involve for
-example the monetary costs of licensed software, data availability issues for data
-hungry models, and the long computation times to run multiple routing scenarios, 
-particularly in  large and complex multimodal transport networks.
+Transport routing is the process of finding alternative routes and costs that connect places on a given transport network.
+Routing often involves complex mathematical algorithms and it is a fundamental step required for transport accessibility analysis, fleet allocation and transport simulation and planning more broadly.
+Despite the long-standing importance of routing analysis to the field of transport and mobility studies, researchers and practitioners still frequently face practical challenges when carrying out this task.
+These challenges involve, for example, the monetary costs of licensed software, limited data availability for data hungry models, and the long computation times required to run multiple routing scenarios, particularly in large and complex multimodal transport networks.
 
-While there is a growing number of open source routing models [@opentripplanneropentripplanner; 
-@lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently 
-process large transport networks. More importantly, these softwares generate
-deterministic estimates of travel times without accounting for issues of service
-reliability (REF) and they only return the fastest route option between a set of
-origins and destinations. This simple approach overlooks how often people's 
-actual route choices deviate from shortest travel time paths, what
-can occur due to multiple factors such as imperfect information, monetary costs 
-and safety issues [@larsen2011travel; @zhu2015people; @conway2019getting].
+While there is a growing number of open source routing models [@opentripplanneropentripplanner; @lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently process large transport networks.
+More importantly, these softwares generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) and only return the fastest route option between a set of origins and destinations (see [@conway2018accounting] for a more in-depth discussion).
+This simple approach overlooks how often people's actual route choices deviate from shortest travel time paths, what can occur due to multiple factors such as imperfect information, monetary costs and safety issues [@larsen2011travel; @zhu2015people; @conway2019getting].
 
-This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open source 
-R package for rapid realistic routing on multimodal transport networks based on
-the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5). 
-R<sup>5</sup> is a powerful next-generation routing engine written in Java and
-developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to
-replace its predecessor OpenTripPlanner. The r5r package provides a simple and
-friendly interface to run R<sup>5</sup> locally from within R, what allows users
-to efficiently calculate travel time matrices or generate multiple route 
-alternatives between origins and destinations using seamless parallel computing.
+This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open source R package for rapid realistic routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R](https://github.com/conveyal/r5)<sup>5</sup>).
+R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to replace its predecessor OpenTripPlanner.
+The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R, which allows users to efficiently calculate travel time matrices or generate multiple route alternatives between origins and destinations using seamless parallel computing.
 
@@
 
-r5r has low data requirements, enabling routing analysis on different contexts 
-wherever data on street networks from OpenStreetMap (OSM) is available. The 
-package is also easily scalable allowing for fast computations at either the 
-city or country level analysis. For multimodal transport routing, r5r will 
-automatically combine OSM data with one or multiple public transport data sets 
-provided by the user in  the standard GTFS format to create a routable transport
-network.
+The r5r package has low data requirements, enabling routing analysis on different contexts wherever data on street networks from [OpenStreetMap](https://www.openstreetmap.org/) (OSM) is available.
+The package is also easily scalable allowing for fast computation at either the city or country levels of analysis.
+For multimodal transport routing, r5r will automatically combine OSM data with one or multiple public transport data sets provided by the user in the standard [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format to create a routable transport network.
 
@@
 
-- `setup_r5()`: builds a multimodal transport network used for routing in 
-R<sup>5</sup>. This function (1) downloads/updates a compiled R<sup>5</sup> JAR 
-file, tailored for the purposes of the package, and stores it locally in the r5r 
-package directory for future use; and (2) combines the OSM and GTFS data sets to 
-build a routable network object.
+-   `setup_r5()`: builds a multimodal transport network used for routing in R<sup>5</sup>.
+This function (1) downloads/updates a compiled R<sup>5</sup> JAR file, tailored for the purposes of the package, and stores it locally in the r5r package directory for future use; and (2) combines the OSM and GTFS data sets to build a routable network object.
 
-- `travel_time_matrix()`: computes travel time estimates between one or 
-multiple origin/destination pairs for a single departure time or for multiple
-departure times over a time window. This function uses an R5-specific extension 
-to the RAPTOR routing algorithm which provides an efficient and systematic 
-sample of multiple departures per minute over a `time_window` set by the user 
-[@conway2017evidencebased]. If the user passes a `time_window` parameter to the 
-function, r5r calculates multiple travel time matrices departing each minute and 
-return additional information on the percentile distribution of travel times 
-that allows one to grasp service reliability issues.
+-   `travel_time_matrix()`: computes travel time estimates between one or multiple origin/destination pairs for a single departure time or for multiple departure times over a time window.
+This function uses an R5-specific extension to the RAPTOR routing algorithm which provides an efficient and systematic sample of multiple departures per minute over a `time_window` set by the user [@conway2017evidencebased].
+If the user passes a `time_window` parameter to the function, r5r calculates multiple travel time matrices departing each minute and return additional information on the percentile distribution of travel times that allows one to grasp service reliability issues.
 
-- `detailed_itineraries()`: computes detailed information on either the
-shortest or multiple alternative routes between one or multiple origin/destination
-pairs for a single departure time. The output includes detailed information on
-route alternatives such as the transport mode, waiting time, travel time and 
-distance of each segment of the trip. This function uses an R5-specific 
-extension to the McRAPTOR routing algorithm to find optimal or less than optimal 
-paths. The specific extension to McRAPTOR to do suboptimal path routing are not 
-documented yet, but a detailed description of base McRAPTOR can be found in @delling2015roundbased.
+-   `detailed_itineraries()`: computes detailed information on either the shortest or multiple alternative routes between one or multiple origin/destination pairs for a single departure time.
+The output includes detailed information on route alternatives such as the transport mode, waiting time, travel time and distance of each segment of the trip.
+This function uses an R5-specific extension to the McRAPTOR routing algorithm to find optimal or less than optimal paths.
+The specific extension to McRAPTOR to do suboptimal path routing are not documented yet, but a detailed description of base McRAPTOR can be found in [@delling2015roundbased].
 
-Both routing functions are versatile so users can easily set customized inputs
-such as transport modes, departure date and times, walking and cycling speeds, 
-maximum trip duration, walking distances and number of public transport 
-transfers, etc.
+Both routing functions are versatile so users can easily set customized inputs such as transport modes, departure date and times, walking and cycling speeds, maximum trip duration, walking distances and number of public transport transfers.
 
@@
 The first step when using r5r is to set the amount of memory available to Java.
-By default `R` allocates only 512 MB of memory for Java processes. To increase 
-available memory to 2GB, for example, we need to set the `java.parameters` 
-option at the beginning of the script as follows:
+By default `R` allocates only 512 MB of memory to the Java Virtual Machine (JVM), which is not enough for the tasks r5r is designed to carry out, such as calculating travel time matrices on large multimodal networks.
+To increase available memory to 2GB[^1], for example, the user needs to set the `java.parameters` option at the beginning of the script as follows:
+
+[^1]: Note that large study areas may require significantly more memory to run properly.
 
@@
 
-The package can be attached (with some other packages used in this article) with 
-the following command:
+The package can be attached (alogside some other packages used in this article) with the following command:
 
@@
 
-To illustrate functionality, the package includes a small sample data set for 
-the city of Porto Alegre (Brazil). It includes three files:
+To illustrate its functionalities, the package includes sample datasets for the cities of São Paulo and Porto Alegre (both in Brazil). Each dataset includes three files:
 
@@
 -   The spatial coordinates of points covering the area in `.csv` format.
-- The spatial coordinates of some of the city's points of interest.
+
+Additionally, the Porto Alegre dataset includes a `.csv` file containing the spatial coordinates of some of the city's points of interest.
+A sample of the Porto Alegre dataset, which will be used in the examples of this paper, is presented below:
 
@@
 
-To build a routable transport network with r5r, the user only needs one command
-line pointing to where the input data is stored:
+To build a routable transport network with r5r and load it in memory, the user needs one command line pointing to the location on disk where the input data is stored:
 
 ```{r, message = FALSE}
-r5r_core <- setup_r5(data_path)
+r5r_core <- setup_r5(data_path, verbose = FALSE)
 ```
 
-The function uses the `.pbf` and the `GTFS.zip` files in the directory pointed 
-by `data_path` to create a multimodal transport network used for routing by 
-R<sup>5</sup>. The resulting `network.dat` as well as some other files used by 
-R<sup>5</sup> are saved inside the supplied directory.
+The function uses the `.pbf` and the `GTFS.zip` files in the directory pointed by `data_path` to create a multimodal transport network used for routing by R<sup>5</sup>.
+The resulting `network.dat` as well as some other files used by R<sup>5</sup> are saved inside the supplied directory.
 
-The presence of a `.pbf` file with OpenStreetMap data is mandatory because it is
-required to create a street network. Meanwhile, a `GTFS.zip` file is only used to
-calculate public transport trips, thus may not be necessary if that's not intended.
-Multiple GTFS feeds can be used simultaneously, in which case R<sup>5</sup> will
-merge all files into a single network automatically.
+The `.pbf` file with OpenStreetMap data is mandatory because it is required to create a street network for the study area.
+The `GTFS.zip` file, conversely, is only required for calculating public transport trips, thus it is not necessary if that's not intended.
+Multiple GTFS feeds can be used simultaneously, in which case R<sup>5</sup> will merge all files into a single public transport network automatically.
 
@@
 
-Calculating a travel time matrix with r5r can be done very efficiently with the
-`travel_time_matrix()` function. The function takes origins/destinations spatial
-location (either as a spatial `sf POINT` object, or as a `data.frame` containing
-the columns `id`, `lon` and `lat`) and travel parameters as inputs and outputs a
-`data.table` with the travel time estimates between all combinations of origins 
-and destinations that can be reached with the given parameters.
-
-The `time_window` parameter is set to 1 by default, which means that travel time 
-estimates will be calculated using a single departure time (specified by 
-`departure_datetime`):
+Calculating a travel time matrix with r5r can be done with the `travel_time_matrix()` function.
+This function takes, as inputs, origins/destinations spatial location (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and travel parameters and outputs a `data.table` with the travel time estimates between all combinations of origins and destinations that can be reached with the given parameters.
+An example of the `travel_time_matrix()` function's usage is presented as follows:
 
@@
 max_trip_duration <- 120
-departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
+departure_datetime <- as.POSIXct("13-05-2019 14:00:00",
                                  format = "%d-%m-%Y %H:%M:%S")
-
 # calculate travel time matrix
@@
                           max_walk_dist = max_walk_dist,
-                          max_trip_duration = max_trip_duration)
+                          max_trip_duration = max_trip_duration,
+                          verbose = FALSE)
 
-head(ttm)
+ttm
 ```
-
-Alternatively, a bigger time window can be specified. In such cases several 
-travel time matrices are calculated using every minute from `departure_datetime` 
-to the end of the given window as a distinct departure time. By default the 
-function outputs the 50th percentile (i.e. the median) travel time between each 
-origin and destination pair, which can be overruled by the `percentiles` 
-parameter:
+The `time_window` parameter is set to 1 by default, which means that travel time estimates are calculated using a single departure time (specified by the `departure_datetime` parameter).
+Alternatively, a bigger time window can be specified.
+In such cases several travel time matrices are calculated using every minute from `departure_datetime` to the end of the given window as a distinct departure time.
+By default the function outputs the 50th percentile (i.e.
+the median) travel time between each origin and destination pair, which can be overruled by the `percentiles` parameter.
 
 ```{r, message = FALSE}
+mode <- c("WALK", "TRANSIT")
+max_walk_dist <- 1000
+max_trip_duration <- 120
+departure_datetime <- as.POSIXct("13-05-2019 14:00:00",
+                                 format = "%d-%m-%Y %H:%M:%S")
 time_window <- 120
-percentiles <- c(25, 50, 75)
+percentiles <- 1:100
 
@@
                           time_window = time_window,
-                          percentiles = percentiles)
+                          percentiles = percentiles,
+                          verbose = FALSE)
 
-ttm
+ttm[, .(fromId, toId, travel_time_p025, travel_time_p050, travel_time_p075)]
 ```
 
-The image below depicts the travel time distribution for each origin/destination in which a trip could be completed within the specified parameters.
-The pairs are ordered in the x axis by their median travel time and each point in the plot area pictures a distinct distribution percentile.
+The plot below depicts the travel time distribution for each origin/destination in which a trip could be completed within the specified parameters.
+The pairs are ordered in the x axis by their median travel time and each point in the plot area represents a distinct distribution percentile.
 
 ```{r, message = FALSE}
-percentiles <- 1:100
-
-ttm <- travel_time_matrix(r5r_core,
-                          origins = points,
-                          destinations = points,
-                          mode = mode,
-                          departure_datetime = departure_datetime,
-                          max_walk_dist = max_walk_dist,
-                          max_trip_duration = max_trip_duration,
-                          time_window = time_window,
-                          percentiles = percentiles)
-
 order_data <- ttm[, .(fromId, toId, travel_time_p025, travel_time_p050, travel_time_p075)]
@@
 
-plot_data <- melt(ttm, measure = pctls, variable = "percentil", value = "travel_time")
+plot_data <- melt(ttm, measure = pctls, variable = "percentil", value = "travel_time", 
+                  na.rm = TRUE)
 plot_data[, percentil := as.integer(regmatches(percentil, regexpr("\\d+", percentil)))]
-plot_data[, travel_time := fifelse(travel_time > max_trip_duration, NA_integer_, travel_time)]
 plot_data[, id := factor(paste0(fromId, "_", toId), levels = id_order)]
@@
 ggplot(plot_data) + 
-   geom_point(aes(id, travel_time), alpha = 0.02, size = 1) + 
-   theme_minimal() +
-   theme(axis.title.x = element_blank(), axis.text.x = element_blank())
+   geom_point(aes(id, travel_time), alpha = 0.005, size = 0.5) +
+   expand_limits(y = 120) +
+   scale_y_continuous(breaks = c(0, 30, 60, 90, 120)) +
+   theme(axis.title.x = element_blank(), 
+         axis.text.x = element_blank(),
+         axis.ticks = element_blank(),
+         panel.border = element_rect(fill = NA, colour = "grey80", size=1)) +
+   labs(title = "Travel time variation among percentiles, per OD pair",
+        y = "Travel Time (min)")
 ```
@@
 
-Often enough one wants more information on the itineraries between 
-origin/destination pairs than simply the travel time. For such cases, the 
-package includes the `detailed_itineraries()` function, which also takes travel 
-restrictions as inputs, but outputs a much more detailed data table at the trip 
-segment level. Another key feature of this function is that it (optionally) 
-provides multiple alternative routes between origin/destination pairs:
+Often enough one wants more information on the itineraries between origin/destination pairs than simply the travel time.
+For such cases, r5r includes the `detailed_itineraries()` function.
+This function also takes travel restrictions as inputs, but outputs a more detailed table of trips disaggregated by segment.
+Similarly to the `travel_time_matrix()` function, the `origins` and `destinations` parameters of `detailed_itineraries()` can be either spatial `sf POINT` objects, or `data.frames` containing the columns `id`, `lon` and `lat`.
+Itineraries are calculated between matching pairs of locations in both datasets, so `origins` and `destinations` inputs must contain the same number of observations.
+However, if the input datasets do not match but `origins` contains only one observation, it is automatically expanded to match the number of observations in `destinations`.
+The calculations are automatically made in parallel, thus optimizing the use of the available computational resources.
 
@@
 max_trip_duration <- 120
-departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
+departure_datetime <- as.POSIXct("13-05-2019 14:00:00",
                                  format = "%d-%m-%Y %H:%M:%S")
@@
 
-dit
+st_set_geometry(dit, NULL)
 ```
-
-Since the `shortest_path` argument has been set to `FALSE`, `dit` holds 
-information for multiple itineraries between the same origin/destination pairs:
+A key feature of `detailed_itineraries()` is that it (optionally) provides multiple alternative routes between origin/destination pairs, when the `shortest_path` parameter is set to `FALSE`.
+This is the case in the example above, so `dit` holds information for multiple itineraries between the same origin/destination pairs, as can be seen below.
 
@@
 
-### Visualize results
+#### Visualize results
 
-`detailed_itineraries()` output is a `sf` object, which means that each trip 
-segment is associated to a spatial geometry that can be plotted with common `R` 
-packages, such as `ggplot2`. The package also offers a handy 
-`street_network_to_sf()` function which extracts the OSM street network used by 
-R<sup>5</sup> as a `sf` object and that can be used to provide geographic 
-context for the visualization.
+`detailed_itineraries()` output is a `sf` object, which means that each trip segment is associated to a spatial geometry that can be plotted with common `R` packages, such as `ggplot2`.
+The package also offers a handy `street_network_to_sf()` function which extracts the OSM street network used by R<sup>5</sup> as a `sf` object and that can be used to provide geographic context for the visualization.
 
@@
 
-The picture above can be used to compare the walking-only to the walking and 
-transit itineraries between each origin/destination pair. Given that Porto 
-Alegre's sample `GTFS.zip` file includes only very few public transport routes, 
-the resulting plot picture some odd scenarios, such as a fairly lengthy 
-circular-like bus route, in the leftmost facet, and a very long walking access 
-trip to the bus, in the rightmost facet.
+The picture above can be used to compare the walking-only to the walking and transit itineraries between each origin/destination pair.
+Given that Porto Alegre's sample `GTFS.zip` file includes only very few public transport routes, the resulting plot picture some odd scenarios, such as a fairly lengthy circular-like bus route, in the leftmost facet, and a very long walking access trip to the bus, in the rightmost facet.
 
@@
 
-The [R<sup>5</sup> routing engine](https://github.com/conveyal/r5) is developed 
-at [Conveyal](https://www.conveyal.com/) with contributions from several developers. 
-This work was supported by the Brazilian Institute for Applied Economic Research 
-(Ipea).
+The [R](https://github.com/conveyal/r5)<sup>5</sup> routing engine is developed at [Conveyal](https://www.conveyal.com/) with contributions from several developers.
+This work was supported by the Brazilian Institute for Applied Economic Research (Ipea).
 
@@
 <!-- Rafa's notes -->
+
 <!-- jean: #>-  mass of assumptions regarding decision-making and routing (Miller, 2018).  -->
+
 <!-- Haugen, 2011; Schwanen, 2008 -->
@@
 <!-- This article presents **r5r**, a new open-source `R` package for rapid  -->
+
 <!-- realistic routing on multimodal transport networks. -->

### 2020-11-11 r-package/paper2020/paper2020.Rmd
Added isochrone section to the paper.

@@
 points <- fread(file.path(data_path, "poa_hexgrid.csv"))
-points <- points[c(sample(1:nrow(points), 100))]
 head(points)
@@
 ```
+
+#### Visualise Isochrones
+
+The sample points dataset provided form a grid covering the study area, so they can easily be used to build ishochrone-like visualizations. An example of how this can be achieved is presented below:
+
+```{r}
+cutoffs <- c(0, 30, 60, 90, 120)
+
+# get a sample of origins
+ids <- sample(ttm$fromId, size = 2)
+ttm_iso <- ttm[fromId %in% ids] 
+origins <- points[id %in% ids]
+setnames(origins, c("fromId", "lon", "lat"))
+
+# aggregate travel-times
+ttm_iso[, isochrones := cut(x=travel_time, breaks=cutoffs), by=.(fromId)]
+
+# join ttm results to points
+ttm_iso[points, on=c('toId' ='id'), `:=`(lon = i.lon, lat = i.lat)]
+ttm_iso <- na.omit(ttm_iso)
+
+ggplot(ttm_iso, aes(x=lon, y=lat, colour = isochrones)) +
+   geom_point(size = 2) +
+   geom_point(data = origins, aes(x=lon, y=lat), colour="blue", size = 2) +
+   coord_map() +
+   scale_colour_brewer(palette = "Reds") +
+   facet_wrap(~fromId) +
+   theme_minimal() +
+   theme(axis.title = element_blank(),
+         panel.border = element_rect(fill = NA, colour = "grey40"))
+```
+
+#### Time window feature
+
 The `time_window` parameter is set to 1 by default, which means that travel time estimates are calculated using a single departure time (specified by the `departure_datetime` parameter).
@@
 ```{r, message = FALSE}
+# sampling destination points
+points <- points[c(sample(1:nrow(points), 100))]
+
 mode <- c("WALK", "TRANSIT")
@@
 Note that the leftmost origin/destination pairs do not exhibit any variation in the distribution of their travel time along the time window.
-This is because the trips between such pairs rely solely on walking to be completed, therefore are not susceptible to transit reliability issues.
+This happens because those pairs are in walking-distance from each other, and walking travel times are not affected by variations in the trip's departure time.
 
@@
                             max_trip_duration = max_trip_duration,
-                            shortest_path = FALSE)
+                            shortest_path = FALSE,
+                            verbose = FALSE)
 
@@
 ```
+
 A key feature of `detailed_itineraries()` is that it (optionally) provides multiple alternative routes between origin/destination pairs, when the `shortest_path` parameter is set to `FALSE`.
@@
 
-#### Visualize results
+#### Visualize itineraries
 
-`detailed_itineraries()` output is a `sf` object, which means that each trip segment is associated to a spatial geometry that can be plotted with common `R` packages, such as `ggplot2`.
+The output of `detailed_itineraries()` is a `sf` object, which means that each trip segment is associated to a spatial geometry that can be plotted with common `R` packages, such as `ggplot2`.
 The package also offers a handy `street_network_to_sf()` function which extracts the OSM street network used by R<sup>5</sup> as a `sf` object and that can be used to provide geographic context for the visualization.
@@
 
-The picture above can be used to compare the walking-only to the walking and transit itineraries between each origin/destination pair.
-Given that Porto Alegre's sample `GTFS.zip` file includes only very few public transport routes, the resulting plot picture some odd scenarios, such as a fairly lengthy circular-like bus route, in the leftmost facet, and a very long walking access trip to the bus, in the rightmost facet.
+The maps above demonstrate how the multiple itinerary options that R5 produces differ from each other.
+The walking-only trips can also be compared to the walking and transit itineraries between each origin/destination pair.
 

### 2020-11-11 r-package/paper2020/paper2020.Rmd
changes to paper text

--- 2020-11-06 r-package/paper2020/paper2020.Rmd
@@
 output:
+  pdf_document: default
   html_document: default
-  pdf_document: default
 abstract: "Routing is a key step in transport planning.
-Nonetheless, researchers and practitioners still face challenges regarding this task, such as the cost of licensed software and the long computation times to run complex scenarios.
-Conveyal's R<sup>5</sup> is a multimodal transport network router that offers meaningful planning features, such as accounting for service reliability issues and returning multiple travel time estimates and itineraries for a single origin/destination pair.
-This paper describes r5r, an open source R package built on top of R<sup>5</sup> that allow users to efficiently compute travel time matrices and generate detailed itineraries between a set of origins and destinations at no expense.
-The development of this package contributes to the transport field as it offers an easy-to-use powerful tool for robust planning and research."
+Nonetheless, researchers and practitioners still face challenges to perform this task due the costs of licensed software or the long computation times to run multiple routing analysis in complex transport networks.
+Conveyal's R<sup>5</sup> is a multimodal transport network router that offers meaningful planning features, such as accounting for service reliability issues and returning multiple travel time estimates and itineraries for a single or multiple origin/destination pairs.
+This paper describes r5r, an open source R package that leverages on R<sup>5</sup> to allow users to efficiently compute travel time matrices and generate detailed itineraries between a set of origins and destinations at no expense.
+The development of this package contributes to the transport field as by offerin an easy-to-use powerful tool for robust planning and research."
 urlcolor: blue
@@
 -   `setup_r5()`: builds a multimodal transport network used for routing in R<sup>5</sup>.
-This function (1) downloads/updates a compiled R<sup>5</sup> JAR file, tailored for the purposes of the package, and stores it locally in the r5r package directory for future use; and (2) combines the OSM and GTFS data sets to build a routable network object.
+This function automatically (1) downloads/updates a compiled R<sup>5</sup> JAR file, tailored for the purposes of the package, and stores it locally in the r5r package directory for future use; and (2) combines the OSM and GTFS data sets to build a routable network object.
 
@@
 
-The package can be installed with the following command:
+The package can be installed with the following command[^1]:
 
@@
 The first step when using r5r is to set the amount of memory available to Java.
-By default `R` allocates only 512 MB of memory to the Java Virtual Machine (JVM), which is not enough for the tasks r5r is designed to carry out, such as calculating travel time matrices on large multimodal networks.
-To increase available memory to 2GB[^1], for example, the user needs to set the `java.parameters` option at the beginning of the script as follows:
+By default `R` allocates only 512 MB of memory to the Java Virtual Machine (JVM), which might not be enough to perform tasks on large multimodal networks.
+To increase available memory to 2GB, for example, the user needs to set the `java.parameters` option at the beginning of the script as follows:
 
-[^1]: Note that large study areas may require significantly more memory to run properly.
+[^1]: To use r5r, one needs to have locally installed Java SE Development Kit, version 11.0.8 or higher. Available at <https://www.oracle.com/java/technologies/javase-jdk11-downloads.html>.
 

### 2020-11-12 r-package/paper2020/paper2020.Rmd
update paper manuscript

@@
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 abstract: "Routing is a key step in transport planning.
@@
 The `.pbf` file with OpenStreetMap data is mandatory because it is required to create a street network for the study area.
-The `GTFS.zip` file, conversely, is only required for calculating public transport trips, thus it is not necessary if that's not intended.
-Multiple GTFS feeds can be used simultaneously, in which case R<sup>5</sup> will merge all files into a single public transport network automatically.
+The `GTFS.zip` file, conversely, is optional and it is only required for calculating public transport trips.
+Multiple GTFS feeds can be used simultaneously, in which case R<sup>5</sup> will automatically merge all files into a single public transport network.
 
@@
 Calculating a travel time matrix with r5r can be done with the `travel_time_matrix()` function.
-This function takes, as inputs, origins/destinations spatial location (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and travel parameters and outputs a `data.table` with the travel time estimates between all combinations of origins and destinations that can be reached with the given parameters.
+This function takes, as inputs, the spatial location of origins/destinations  (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and a few travel parameters such as maximum trip duration or walking distance.
+The function outputs `data.table` with the travel time estimates between all combinations of origins and destinations that can be reached with the given parameters.
 An example of the `travel_time_matrix()` function's usage is presented as follows:
@@
 
-ttm
+head(ttm)
 ```
 
-#### Visualise Isochrones
+#### Visualize Isochrones
 
-The sample points dataset provided form a grid covering the study area, so they can easily be used to build ishochrone-like visualizations. An example of how this can be achieved is presented below:
+If the destination points cover the entire study area, the output can easily be used to build ishochrone-like visualizations.
+In our example, we can visualize the isochrone departing from the central bus station as follows:
 
 ```{r}
-cutoffs <- c(0, 30, 60, 90, 120)
+cutoffs <- seq(0, 120, 20)
 
-# get a sample of origins
-ids <- sample(ttm$fromId, size = 2)
-ttm_iso <- ttm[fromId %in% ids] 
-origins <- points[id %in% ids]
-setnames(origins, c("fromId", "lon", "lat"))
+# select the travel time matrix departing from a given origin
+central_bus_stn <- points[291,]
+ttm_iso <- ttm[fromId %in% central_bus_stn$id ]
 
 # aggregate travel-times
-ttm_iso[, isochrones := cut(x=travel_time, breaks=cutoffs), by=.(fromId)]
+ttm_iso[, isochrones := cut(x=travel_time, breaks=cutoffs), by= fromId]
 
-# join ttm results to points
+# join travel-matrix results to points
 ttm_iso[points, on=c('toId' ='id'), `:=`(lon = i.lon, lat = i.lat)]
-ttm_iso <- na.omit(ttm_iso)
 
@@
    geom_point(size = 2) +
-   geom_point(data = origins, aes(x=lon, y=lat), colour="blue", size = 2) +
+   geom_point(data = central_bus_stn, aes(x=lon, y=lat), colour="blue", size = 2) +
    coord_map() +
@@
 
-#### Time window feature
+#### Accounting for travel-time uncertainty
 
-The `time_window` parameter is set to 1 by default, which means that travel time estimates are calculated using a single departure time (specified by the `departure_datetime` parameter).
-Alternatively, a bigger time window can be specified.
-In such cases several travel time matrices are calculated using every minute from `departure_datetime` to the end of the given window as a distinct departure time.
+By default, `r5r` will return a single travel time estimate for each origin-destination pair at a set `departure_datetime`.
+However, this can provide biased estimates due to how service levels varies across the day (reference to https://www.sciencedirect.com/science/article/pii/S0966692318305209).
+To account
+for these variations, one can set a `time_window` in the `travel_time_matrix` function. In this case, several matrices are calculated using multiple departure
+times every minute within the set time window.
 By default the function outputs the 50th percentile (i.e.
-the median) travel time between each origin and destination pair, which can be overruled by the `percentiles` parameter.
+the median) travel time between each origin and destination pair in within that time window.
+Nonetheless, it is possible to retrieve information on travel time estimates at different points of the distribution using the `percentiles` parameter, as follows.
 
 ```{r, message = FALSE}
-# sampling destination points
-points <- points[c(sample(1:nrow(points), 100))]
 
-mode <- c("WALK", "TRANSIT")
-max_walk_dist <- 1000
-max_trip_duration <- 120
-departure_datetime <- as.POSIXct("13-05-2019 14:00:00",
-                                 format = "%d-%m-%Y %H:%M:%S")
 time_window <- 120
@@
 # calculate travel time matrix
-ttm <- travel_time_matrix(r5r_core,
-                          origins = points,
-                          destinations = points,
+ttm_tw <- travel_time_matrix(r5r_core,
+                          origins = points[1:800],
+                          destinations = central_bus_stn,
                           mode = mode,
@@
 
-ttm[, .(fromId, toId, travel_time_p025, travel_time_p050, travel_time_p075)]
 ```
 
-The plot below depicts the travel time distribution for each origin/destination in which a trip could be completed within the specified parameters.
-The pairs are ordered in the x axis by their median travel time and each point in the plot area represents a distinct distribution percentile.
+In the example above we calculated how the estimates of travel-time to arrive at the central bus station vary by place of trip origin.
+Now we can visualize how these estimates are more uncertain when leaving from some places than others.
+
+As expected, there is little to no uncertainty in travel times from places that are very close (walking distance) to the central bus station.
+As the figure shows, travel time uncertainties tend to be smaller for places that are very close to the central bus station.
+This happens because shorter trips are less affected by variations in the trip's departure time and service frequency levels.
 
 ```{r, message = FALSE}
-order_data <- ttm[, .(fromId, toId, travel_time_p025, travel_time_p050, travel_time_p075)]
-order_data <- order_data[order(travel_time_p050, travel_time_p025, travel_time_p075)]
-id_order   <- order_data[, id := paste0(fromId, "_", toId)]$id
+# reshape data
+plot_data <- melt(ttm_tw, measure = patterns("^travel_time_p"), variable = "percentile", value = "travel_time")
+plot_data[, mediantt := median(travel_time), by=fromId]
 
-pctls <- paste0("travel_time_p", formatC(percentiles, width = 3, flag = "0"))
-
-plot_data <- melt(ttm, measure = pctls, variable = "percentil", value = "travel_time", 
-                  na.rm = TRUE)
-plot_data[, percentil := as.integer(regmatches(percentil, regexpr("\\d+", percentil)))]
-plot_data[, id := factor(paste0(fromId, "_", toId), levels = id_order)]
-
-ggplot(plot_data) + 
-   geom_point(aes(id, travel_time), alpha = 0.005, size = 0.5) +
+ggplot(data=plot_data, aes(y = travel_time, x = reorder(fromId, mediantt))) +
+  geom_point(alpha = 0.05, size = .7) +
+  geom_line(aes(y=mediantt, group=toId), color='blue') +
   expand_limits(y = 120) +
   scale_y_continuous(breaks = c(0, 30, 60, 90, 120)) +
+  theme_minimal() +
   theme(axis.title.x = element_blank(), 
@@
        y = "Travel Time (min)")
+
 ```
 
-Note that the leftmost origin/destination pairs do not exhibit any variation in the distribution of their travel time along the time window.
-This happens because those pairs are in walking-distance from each other, and walking travel times are not affected by variations in the trip's departure time.
+<!-- alternative plot -->
+
+<!-- ```{r, message = FALSE} -->
+<!-- # reshape data -->
+<!-- plot_data <- melt(ttm_tw, measure = patterns("^travel_time_p"), variable = "percentile", value = "travel_time") -->
+<!-- plot_data[, mediantt := travel_time[which(percentile=='travel_time_p050')], by=fromId] -->
+
+<!-- ggplot(data=plot_data, aes(y = travel_time, x = reorder(fromId, mediantt))) + -->
+<!--   geom_point(alpha = 0.005, size = 0.5) + -->
+<!--   geom_line(aes(group=toId), color='blue') -->
+
+<!-- # plot -->
+<!-- ggplot(data=plot_data, aes(x = travel_time, y = reorder(fromId, mediantt))) + -->
+<!--   geom_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha=.9) + -->
+<!--   expand_limits(x = 120) + -->
+<!--   scale_x_continuous(breaks = c(0, 30, 60, 90, 120)) + -->
+<!--   theme(axis.title.y = element_blank(),  -->
+<!--         axis.text.y = element_blank(), -->
+<!--         axis.ticks = element_blank(), -->
+<!--         panel.border = element_rect(fill = NA, colour = "grey80", size=1)) + -->
+<!--   labs(title = "Travel time variation among percentiles, per OD pair", -->
+<!--        x = "Travel Time (min)") -->
+
+<!-- ``` -->
 
@@
 
-Often enough one wants more information on the itineraries between origin/destination pairs than simply the travel time.
+Often enough one wants more information than simply the travel time between origin/destination pairs.
 For such cases, r5r includes the `detailed_itineraries()` function.
-This function also takes travel restrictions as inputs, but outputs a more detailed table of trips disaggregated by segment.
+This function outputs a more detailed table of alternative routes disaggregated by trip segment.
 Similarly to the `travel_time_matrix()` function, the `origins` and `destinations` parameters of `detailed_itineraries()` can be either spatial `sf POINT` objects, or `data.frames` containing the columns `id`, `lon` and `lat`.
@@
 # set restrictions
-mode <- c("WALK", "TRANSIT")
 max_walk_dist <- Inf
-max_trip_duration <- 120
-departure_datetime <- as.POSIXct("13-05-2019 14:00:00",
-                                 format = "%d-%m-%Y %H:%M:%S")
 
@@
 
-st_set_geometry(dit, NULL)
+head(st_set_geometry(dit, NULL))
 ```
 
-A key feature of `detailed_itineraries()` is that it (optionally) provides multiple alternative routes between origin/destination pairs, when the `shortest_path` parameter is set to `FALSE`.
+A key feature of `detailed_itineraries()` is that it (optionally) provides multiple alternative routes between origin/destination pairs when the `shortest_path` parameter is set to `FALSE`.
 This is the case in the example above, so `dit` holds information for multiple itineraries between the same origin/destination pairs, as can be seen below.
@@
 
-The output of `detailed_itineraries()` is a `sf` object, which means that each trip segment is associated to a spatial geometry that can be plotted with common `R` packages, such as `ggplot2`.
-The package also offers a handy `street_network_to_sf()` function which extracts the OSM street network used by R<sup>5</sup> as a `sf` object and that can be used to provide geographic context for the visualization.
+The output of `detailed_itineraries()` is an `sf` object, which means that each trip segment is associated to a spatial geometry that can be plotted with common `R` packages, such as `ggplot2`.
+The r5r package also offers a handy `street_network_to_sf()` function which extracts the OSM street network used by R<sup>5</sup> as a `sf` object and that can be used to provide geographic context for the visualization.
 

### 2020-11-13 r-package/paper2020/paper2020.Rmd
update paper plot

@@
   theme_minimal() +
-  theme(axis.title.x = element_blank(), 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.title.x = element_blank(), 
         axis.text.x = element_blank(),
@@
         panel.border = element_rect(fill = NA, colour = "grey80", size=1)) +
-  labs(title = "Travel time variation among percentiles, per OD pair",
-       y = "Travel Time (min)")
+  labs(title = "Travel time uncertainty per destination",
+       y = "Travel Time (min)", x='Destinations')
 

### 2020-11-13 r-package/paper2020/paper2020.Rmd
update paper

--- 2020-11-12 r-package/paper2020/paper2020.Rmd
@@
 
-Additionally, the Porto Alegre dataset includes a `.csv` file containing the spatial coordinates of some of the city's points of interest.
 A sample of the Porto Alegre dataset, which will be used in the examples of this paper, is presented below:
@@
 plot_data <- melt(ttm_tw, measure = patterns("^travel_time_p"), variable = "percentile", value = "travel_time")
-plot_data[, mediantt := median(travel_time), by=fromId]
+plot_data[, mediantt := median(travel_time, na.rm=T), by=fromId]
 
 ggplot(data=plot_data, aes(y = travel_time, x = reorder(fromId, mediantt))) +
-  geom_point(alpha = 0.05, size = .7) +
-  geom_line(aes(y=mediantt, group=toId), color='blue') +
+  geom_point(alpha = 0.01, size = .7) +
+  geom_line(aes(y=mediantt, group=toId), color="#FE9F45", size=1.5) +
   expand_limits(y = 120) +
@@
 ```{r, message = FALSE}
-# load points of interests and set origin and destination
-points <- fread(file.path(data_path, "poa_points_of_interest.csv"))
-origins <- points[c(10, 14),]
-destinations <- points[c(12, 15),]
-
-# set restrictions
-max_walk_dist <- Inf
-
 dit <- detailed_itineraries(r5r_core,
-                            origins = origins,
-                            destinations = destinations,
+                            origins = points[612],# 793
+                            destinations = central_bus_stn,
                             mode = mode,
@@
 
-dit$facet_labels <- paste0(dit$fromId, " to\n", dit$toId)
+dit$facet_labels <- paste0("Alternative #", dit$option)
 
@@
   geom_sf(data = dit, aes(color = mode), size = 1) +
-   facet_wrap(~ facet_labels, nrow = 1) +
+  facet_wrap(.~facet_labels) +
   theme_minimal() + 

### 2021-01-06 r-package/paper2020/paper2020.Rmd
Added a garbage collector call at the end of stop_r5 to clean up memory.

@@
 output:
+  word_document: default
+  pdf_document: default
   html_document: default
-  pdf_document: default
-abstract: "Routing is a key step in transport planning.
-Nonetheless, researchers and practitioners still face challenges to perform this task due the costs of licensed software or the long computation times to run multiple routing analysis in complex transport networks.
-Conveyal's R<sup>5</sup> is a multimodal transport network router that offers meaningful planning features, such as accounting for service reliability issues and returning multiple travel time estimates and itineraries for a single or multiple origin/destination pairs.
-This paper describes r5r, an open source R package that leverages on R<sup>5</sup> to allow users to efficiently compute travel time matrices and generate detailed itineraries between a set of origins and destinations at no expense.
-The development of this package contributes to the transport field as by offerin an easy-to-use powerful tool for robust planning and research."
+abstract: Routing is a key step in transport planning. Nonetheless, researchers and
+  practitioners still face challenges to perform this task due the costs of licensed
+  software or the long computation times to run multiple routing analysis in complex
+  transport networks. Conveyal's R<sup>5</sup> is a multimodal transport network router
+  that offers meaningful planning features, such as accounting for service reliability
+  issues and returning multiple travel time estimates and itineraries for a single
+  or multiple origin/destination pairs. This paper describes r5r, an open source R
+  package that leverages on R<sup>5</sup> to allow users to efficiently compute travel
+  time matrices and generate detailed itineraries between a set of origins and destinations
+  at no expense. The development of this package contributes to the transport field
+  as by offerin an easy-to-use powerful tool for robust planning and research.
 urlcolor: blue
@@
 While there is a growing number of open source routing models [@opentripplanneropentripplanner; @lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently process large transport networks.
-More importantly, these softwares generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) and only return the fastest route option between a set of origins and destinations (see [@conway2018accounting] for a more in-depth discussion).
+More importantly, those tools generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) and only return the fastest route option between a set of origins and destinations (see [@conway2018accounting] for a more in-depth discussion).
 This simple approach overlooks how often people's actual route choices deviate from shortest travel time paths, what can occur due to multiple factors such as imperfect information, monetary costs and safety issues [@larsen2011travel; @zhu2015people; @conway2019getting].
@@
 This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open source R package for rapid realistic routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R](https://github.com/conveyal/r5)<sup>5</sup>).
-R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to replace its predecessor OpenTripPlanner.
-The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R, which allows users to efficiently calculate travel time matrices or generate multiple route alternatives between origins and destinations using seamless parallel computing.
+R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to provide an efficient backend for analytic applications - such as accessibility assessment The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R, which allows users to efficiently calculate travel time matrices or generate multiple route alternatives between origins and destinations using seamless parallel computing.
 
@@
 Both routing functions are versatile so users can easily set customized inputs such as transport modes, departure date and times, walking and cycling speeds, maximum trip duration, walking distances and number of public transport transfers.
+We will focus on results obtained from `travel_time_matrix()` on the findings discussed in the following section.
 
@@
 
-The first step when using r5r is to set the amount of memory available to Java.
 By default `R` allocates only 512 MB of memory to the Java Virtual Machine (JVM), which might not be enough to perform tasks on large multimodal networks.
-To increase available memory to 2GB, for example, the user needs to set the `java.parameters` option at the beginning of the script as follows:
+To increase the amount of memory available to the package, the user needs to set the `java.parameters` option at the beginning of the script.
+After that, the package can be attached (alongside some other packages used in this article), as follows:
 
@@
 options(java.parameters = "-Xmx2G")
-```
 
-The package can be attached (alogside some other packages used in this article) with the following command:
-
-```{r, message = FALSE}
 library(r5r)
@@
 
-To illustrate its functionalities, the package includes sample datasets for the cities of São Paulo and Porto Alegre (both in Brazil). Each dataset includes three files:
+The package includes sample datasets for the cities of São Paulo and Porto Alegre (both in Brazil), which are used in the package's vignette and examples to illustrate its functionalities.
+Each dataset includes:
 
@@
 
-A sample of the Porto Alegre dataset, which will be used in the examples of this paper, is presented below:
+### Build a routable transport network
 
-```{r}
+To build a routable transport network with r5r and load it in memory, the user needs one command line pointing to the location on disk where the input data is stored.
+In the examples of this paper, we will be using the provided Porto Alegre dataset.
+
+```{r, message = FALSE}
 # points to directory with data
@@
 
-list.files(data_path)
-
-# read points of origin and destination
-points <- fread(file.path(data_path, "poa_hexgrid.csv"))
-head(points)
-```
-
-### Build a routable transport network
-
-To build a routable transport network with r5r and load it in memory, the user needs one command line pointing to the location on disk where the input data is stored:
-
-```{r, message = FALSE}
 r5r_core <- setup_r5(data_path, verbose = FALSE)
@@
 The function uses the `.pbf` and the `GTFS.zip` files in the directory pointed by `data_path` to create a multimodal transport network used for routing by R<sup>5</sup>.
-The resulting `network.dat` as well as some other files used by R<sup>5</sup> are saved inside the supplied directory.
-
-The `.pbf` file with OpenStreetMap data is mandatory because it is required to create a street network for the study area.
-The `GTFS.zip` file, conversely, is optional and it is only required for calculating public transport trips.
 Multiple GTFS feeds can be used simultaneously, in which case R<sup>5</sup> will automatically merge all files into a single public transport network.
+The resulting `network.dat` as well as some other files used by R<sup>5</sup> are saved inside the supplied directory for later reuse.
 
@@
 This function takes, as inputs, the spatial location of origins/destinations  (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and a few travel parameters such as maximum trip duration or walking distance.
-The function outputs `data.table` with the travel time estimates between all combinations of origins and destinations that can be reached with the given parameters.
-An example of the `travel_time_matrix()` function's usage is presented as follows:
+
+By default, `r5r` will return a single travel time estimate for each origin-destination pair at a set `departure_datetime`.
+However, this can provide biased estimates due to how service levels varies across the day (reference to https://www.sciencedirect.com/science/article/pii/S0966692318305209).
+To account
+for these variations, the user can set a `time_window` in the `travel_time_matrix()` function. In this case, several matrices are calculated using multiple departure
+times every minute within the set time window.
+By default the function outputs the 50th percentile (i.e.
+the median) travel time between each origin and destination pair  within that time window.
+Nonetheless, it is possible to retrieve information on travel time estimates at different points of the distribution using the `percentiles` parameter.
+The function outputs a `data.table` with the travel time estimates between all combinations of origins and destinations that can be reached with the given parameters.
+An example of the `travel_time_matrix()` function's usage is presented below:
 
 ```{r, message = FALSE}
+# read points of origin and destination
+points <- fread(file.path(data_path, "poa_hexgrid.csv"))
+
 mode <- c("WALK", "TRANSIT")
@@
                                  format = "%d-%m-%Y %H:%M:%S")
+
+time_window <- 120
+percentiles <- c(20, 40, 60, 80, 99)
+
 # calculate travel time matrix
@@
                           max_trip_duration = max_trip_duration,
+                          time_window = time_window,
+                          percentiles = percentiles,
                           verbose = FALSE)
@@
 head(ttm)
+```
+
+#### Accounting for travel-time uncertainty
+
+In the example above we calculated how the estimates of travel-time to arrive at the central bus station vary by place of trip origin.
+
+We can visualize how the estimated travel times are more uncertain when leaving from some places than others.
+The plot below shows travel times from all origin points to the central bus station at five percentiles (20th, 40th, 60th, 80th, 99th).
+As expected, there is little to no uncertainty in travel times from places that are very close (walking distance) to the central bus station.
+Conversely, travel times from places that are far away are more affected by variations in the trip's departure time and service frequency levels.
+
+```{r, message = FALSE}
+# select the travel time matrix departing from a given origin
+central_bus_stn <- points[291,]
+ttm_tw <- ttm[toId %in% central_bus_stn$id ]
+
+# reshape data
+plot_data <- melt(ttm_tw, measure = patterns("^travel_time_p"), variable = "percentile", value = "travel_time")
+plot_data[, mediantt := median(travel_time, na.rm=T), by=fromId]
+
+ggplot(data=plot_data, aes(y = travel_time, x = reorder(fromId, mediantt))) +
+  geom_point(alpha = 0.1, size = .7) +
+  geom_line(aes(y=mediantt, group=toId), color="#FE9F45", size=1.5) +
+  expand_limits(y = 120) +
+  scale_y_continuous(breaks = c(0, 30, 60, 90, 120)) +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        axis.title.x = element_blank(), 
+        axis.text.x = element_blank(),
+        axis.ticks = element_blank(),
+        panel.border = element_rect(fill = NA, colour = "grey80", size=1)) +
+  labs(title = "Travel time uncertainty per destination",
+       y = "Travel Time (min)", x='Destinations')
 ```
@@
 # aggregate travel-times
-ttm_iso[, isochrones := cut(x=travel_time, breaks=cutoffs), by= fromId]
+ttm_iso[, isochrones := cut(x=travel_time_p099, breaks=cutoffs), by= fromId]
 
@@
 ```
-
-#### Accounting for travel-time uncertainty
-
-By default, `r5r` will return a single travel time estimate for each origin-destination pair at a set `departure_datetime`.
-However, this can provide biased estimates due to how service levels varies across the day (reference to https://www.sciencedirect.com/science/article/pii/S0966692318305209).
-To account
-for these variations, one can set a `time_window` in the `travel_time_matrix` function. In this case, several matrices are calculated using multiple departure
-times every minute within the set time window.
-By default the function outputs the 50th percentile (i.e.
-the median) travel time between each origin and destination pair in within that time window.
-Nonetheless, it is possible to retrieve information on travel time estimates at different points of the distribution using the `percentiles` parameter, as follows.
-
-```{r, message = FALSE}
-
-time_window <- 120
-percentiles <- 1:100
-
-# calculate travel time matrix
-ttm_tw <- travel_time_matrix(r5r_core,
-                          origins = points[1:800],
-                          destinations = central_bus_stn,
-                          mode = mode,
-                          departure_datetime = departure_datetime,
-                          max_walk_dist = max_walk_dist,
-                          max_trip_duration = max_trip_duration,
-                          time_window = time_window,
-                          percentiles = percentiles,
-                          verbose = FALSE)
-
-```
-
-In the example above we calculated how the estimates of travel-time to arrive at the central bus station vary by place of trip origin.
-Now we can visualize how these estimates are more uncertain when leaving from some places than others.
-
-As expected, there is little to no uncertainty in travel times from places that are very close (walking distance) to the central bus station.
-As the figure shows, travel time uncertainties tend to be smaller for places that are very close to the central bus station.
-This happens because shorter trips are less affected by variations in the trip's departure time and service frequency levels.
-
-```{r, message = FALSE}
-# reshape data
-plot_data <- melt(ttm_tw, measure = patterns("^travel_time_p"), variable = "percentile", value = "travel_time")
-plot_data[, mediantt := median(travel_time, na.rm=T), by=fromId]
-
-ggplot(data=plot_data, aes(y = travel_time, x = reorder(fromId, mediantt))) +
-  geom_point(alpha = 0.01, size = .7) +
-  geom_line(aes(y=mediantt, group=toId), color="#FE9F45", size=1.5) +
-  expand_limits(y = 120) +
-  scale_y_continuous(breaks = c(0, 30, 60, 90, 120)) +
-  theme_minimal() +
-  theme(panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(),
-        axis.title.x = element_blank(), 
-        axis.text.x = element_blank(),
-        axis.ticks = element_blank(),
-        panel.border = element_rect(fill = NA, colour = "grey80", size=1)) +
-  labs(title = "Travel time uncertainty per destination",
-       y = "Travel Time (min)", x='Destinations')
-
-```
-
-<!-- alternative plot -->
-
-<!-- ```{r, message = FALSE} -->
-<!-- # reshape data -->
-<!-- plot_data <- melt(ttm_tw, measure = patterns("^travel_time_p"), variable = "percentile", value = "travel_time") -->
-<!-- plot_data[, mediantt := travel_time[which(percentile=='travel_time_p050')], by=fromId] -->
-
-<!-- ggplot(data=plot_data, aes(y = travel_time, x = reorder(fromId, mediantt))) + -->
-<!--   geom_point(alpha = 0.005, size = 0.5) + -->
-<!--   geom_line(aes(group=toId), color='blue') -->
-
-<!-- # plot -->
-<!-- ggplot(data=plot_data, aes(x = travel_time, y = reorder(fromId, mediantt))) + -->
-<!--   geom_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha=.9) + -->
-<!--   expand_limits(x = 120) + -->
-<!--   scale_x_continuous(breaks = c(0, 30, 60, 90, 120)) + -->
-<!--   theme(axis.title.y = element_blank(),  -->
-<!--         axis.text.y = element_blank(), -->
-<!--         axis.ticks = element_blank(), -->
-<!--         panel.border = element_rect(fill = NA, colour = "grey80", size=1)) + -->
-<!--   labs(title = "Travel time variation among percentiles, per OD pair", -->
-<!--        x = "Travel Time (min)") -->
-
-<!-- ``` -->
-
-### Calculate detailed itineraries
-
-Often enough one wants more information than simply the travel time between origin/destination pairs.
-For such cases, r5r includes the `detailed_itineraries()` function.
-This function outputs a more detailed table of alternative routes disaggregated by trip segment.
-Similarly to the `travel_time_matrix()` function, the `origins` and `destinations` parameters of `detailed_itineraries()` can be either spatial `sf POINT` objects, or `data.frames` containing the columns `id`, `lon` and `lat`.
-Itineraries are calculated between matching pairs of locations in both datasets, so `origins` and `destinations` inputs must contain the same number of observations.
-However, if the input datasets do not match but `origins` contains only one observation, it is automatically expanded to match the number of observations in `destinations`.
-The calculations are automatically made in parallel, thus optimizing the use of the available computational resources.
-
-```{r, message = FALSE}
-dit <- detailed_itineraries(r5r_core,
-                            origins = points[612],# 793
-                            destinations = central_bus_stn,
-                            mode = mode,
-                            departure_datetime = departure_datetime,
-                            max_walk_dist = max_walk_dist,
-                            max_trip_duration = max_trip_duration,
-                            shortest_path = FALSE,
-                            verbose = FALSE)
-
-head(st_set_geometry(dit, NULL))
-```
-
-A key feature of `detailed_itineraries()` is that it (optionally) provides multiple alternative routes between origin/destination pairs when the `shortest_path` parameter is set to `FALSE`.
-This is the case in the example above, so `dit` holds information for multiple itineraries between the same origin/destination pairs, as can be seen below.
-
-```{r, message = FALSE}
-setDT(copy(dit))[, .(n_itineraries = uniqueN(option)), by = .(fromId, toId)]
-```
-
-#### Visualize itineraries
-
-The output of `detailed_itineraries()` is an `sf` object, which means that each trip segment is associated to a spatial geometry that can be plotted with common `R` packages, such as `ggplot2`.
-The r5r package also offers a handy `street_network_to_sf()` function which extracts the OSM street network used by R<sup>5</sup> as a `sf` object and that can be used to provide geographic context for the visualization.
-
-```{r, message = FALSE}
-# extract OSM network
-street_net <- street_network_to_sf(r5r_core)
-
-dit$facet_labels <- paste0("Alternative #", dit$option)
-
-ggplot() +
-  geom_sf(data = street_net$edges, color = "gray85") +
-  geom_sf(data = dit, aes(color = mode), size = 1) +
-  facet_wrap(.~facet_labels) +
-  theme_minimal() + 
-  theme(legend.position = "bottom", axis.text.x = element_text(angle = 30, hjust = 1))
-```
-
-The maps above demonstrate how the multiple itinerary options that R5 produces differ from each other.
-The walking-only trips can also be compared to the walking and transit itineraries between each origin/destination pair.
 

### 2021-01-14 r-package/paper2020/paper2020.Rmd
Revised the paper to comply with Transport Findings' rules. - The paper's body now has 993 words (limit of 1000 words), not counting source code, tables, and references. - The abstract is 98 words long (limit of 100 words).
Matt's comments were addressed as well.

@@
   html_document: default
-abstract: Routing is a key step in transport planning. Nonetheless, researchers and
-  practitioners still face challenges to perform this task due the costs of licensed
-  software or the long computation times to run multiple routing analysis in complex
-  transport networks. Conveyal's R<sup>5</sup> is a multimodal transport network router
-  that offers meaningful planning features, such as accounting for service reliability
-  issues and returning multiple travel time estimates and itineraries for a single
-  or multiple origin/destination pairs. This paper describes r5r, an open source R
-  package that leverages on R<sup>5</sup> to allow users to efficiently compute travel
-  time matrices and generate detailed itineraries between a set of origins and destinations
-  at no expense. The development of this package contributes to the transport field
-  as by offerin an easy-to-use powerful tool for robust planning and research.
+abstract: Routing is a key step in transport planning.
+Nonetheless, researchers and practitioners face challenges to perform this task, such as the costs of licensed software and long computation times.
+Conveyal’s R<sup>5<\sup> is a multimodal transport network router that offers meaningful planning features, such as calculating travel times over a time window and returning multiple itineraries for a single or multiple origin/destination pairs.
+This paper describes r5r, an open source R package that leverages on R<sup>5<\sup> to allow users to efficiently compute travel time matrices and generate detailed itineraries between a set of origins and destinations at no expense.
 urlcolor: blue
@@
 
-Transport routing is the process of finding alternative routes and costs that connect places on a given transport network.
-Routing often involves complex mathematical algorithms and it is a fundamental step required for transport accessibility analysis, fleet allocation and transport simulation and planning more broadly.
-Despite the long-standing importance of routing analysis to the field of transport and mobility studies, researchers and practitioners still frequently face practical challenges when carrying out this task.
-These challenges involve, for example, the monetary costs of licensed software, limited data availability for data hungry models, and the long computation times required to run multiple routing scenarios, particularly in large and complex multimodal transport networks.
+Transport routing is the process of finding alternative routes and costs that connect places on a given transport network, and it is a key step in transport accessibility analysis, fleet allocation and transport simulation and planning more broadly.
+Researchers and practitioners still frequently face practical challenges when carrying out routing analysis tasks, such as the monetary costs of licensed software, limited data availability for data hungry models, and the long computation times required to run multiple routing scenarios, particularly in large and complex multimodal transport networks.
 
 While there is a growing number of open source routing models [@opentripplanneropentripplanner; @lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently process large transport networks.
-More importantly, those tools generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) and only return the fastest route option between a set of origins and destinations (see [@conway2018accounting] for a more in-depth discussion).
+This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open source R package for routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5).
+R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to provide an efficient backend for analytic applications - such as accessibility assessment.
+The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R, which allows users to efficiently calculate travel time matrices or generate multiple route alternatives between origins and destinations using seamless parallel computing.
+
+<!-- More importantly, those tools generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) and only return the fastest route option between a set of origins and destinations (see [@conway2018accounting] for a more in-depth discussion).
 This simple approach overlooks how often people's actual route choices deviate from shortest travel time paths, what can occur due to multiple factors such as imperfect information, monetary costs and safety issues [@larsen2011travel; @zhu2015people; @conway2019getting].
-
-This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open source R package for rapid realistic routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R](https://github.com/conveyal/r5)<sup>5</sup>).
-R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to provide an efficient backend for analytic applications - such as accessibility assessment The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R, which allows users to efficiently calculate travel time matrices or generate multiple route alternatives between origins and destinations using seamless parallel computing.
+-->
 
@@
 
-The r5r package has low data requirements, enabling routing analysis on different contexts wherever data on street networks from [OpenStreetMap](https://www.openstreetmap.org/) (OSM) is available.
-The package is also easily scalable allowing for fast computation at either the city or country levels of analysis.
-For multimodal transport routing, r5r will automatically combine OSM data with one or multiple public transport data sets provided by the user in the standard [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format to create a routable transport network.
+The r5r package has low data requirements and is easily scalable, allowing fast computation of routes and travel times at either the city or country levels of analysis.
+Street networks from [OpenStreetMap](https://www.openstreetmap.org/) (OSM) are automatically combined with one or multiple public transport data sets provided by the user in the standard [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format to create a routable transport network.
 
@@
 -   `setup_r5()`: builds a multimodal transport network used for routing in R<sup>5</sup>.
-This function automatically (1) downloads/updates a compiled R<sup>5</sup> JAR file, tailored for the purposes of the package, and stores it locally in the r5r package directory for future use; and (2) combines the OSM and GTFS data sets to build a routable network object.
+This function automatically (1) downloads/updates a compiled R<sup>5</sup> JAR file and stores it locally in the r5r package directory for future use; and (2) combines the OSM and GTFS datasets to build a routable network object.
 
--   `travel_time_matrix()`: computes travel time estimates between one or multiple origin/destination pairs for a single departure time or for multiple departure times over a time window.
-This function uses an R5-specific extension to the RAPTOR routing algorithm which provides an efficient and systematic sample of multiple departures per minute over a `time_window` set by the user [@conway2017evidencebased].
-If the user passes a `time_window` parameter to the function, r5r calculates multiple travel time matrices departing each minute and return additional information on the percentile distribution of travel times that allows one to grasp service reliability issues.
+-   `travel_time_matrix()`: computes travel time estimates between one or multiple origin/destination pairs for a single departure time or for multiple departure times over a `time_window` set by the user.
+This function uses an R5-specific extension to the RAPTOR routing algorithm which provides an efficient and systematic sampling of multiple simulated schedules when using frequency-based GTFS data  [@conway2017evidencebased].
+
+<!-- If the user passes a `time_window` parameter to the function, r5r calculates multiple travel time matrices departing each minute and returns additional information on the percentile distribution of travel times that allows one to grasp service reliability issues.
+-->
 
@@
 The output includes detailed information on route alternatives such as the transport mode, waiting time, travel time and distance of each segment of the trip.
-This function uses an R5-specific extension to the McRAPTOR routing algorithm to find optimal or less than optimal paths.
-The specific extension to McRAPTOR to do suboptimal path routing are not documented yet, but a detailed description of base McRAPTOR can be found in [@delling2015roundbased].
+This function uses an R5-specific extension[^1] to the McRAPTOR [@delling2015roundbased] routing algorithm to find optimal or less than optimal paths.
+
+[^1]: The specific extension to McRAPTOR to do suboptimal path routing is not documented yet.
 
@@
 
-The package can be installed with the following command[^1]:
-
-```{r, message = FALSE, eval = FALSE}
-# stable version
-install.packages("r5r")
-```
-
-By default `R` allocates only 512 MB of memory to the Java Virtual Machine (JVM), which might not be enough to perform tasks on large multimodal networks.
-To increase the amount of memory available to the package, the user needs to set the `java.parameters` option at the beginning of the script.
-After that, the package can be attached (alongside some other packages used in this article), as follows:
-
-[^1]: To use r5r, one needs to have locally installed Java SE Development Kit, version 11.0.8 or higher. Available at <https://www.oracle.com/java/technologies/javase-jdk11-downloads.html>.
+After installed with the `install.packages("r5r")` command, the package can be attached (alongside some other packages used in this article), as follows:
 
 ```{r, message = FALSE}
-options(java.parameters = "-Xmx2G")
-
 library(r5r)
@@
 library(ggplot2)
+library(akima)
 ```
 
-The package includes sample datasets for the cities of São Paulo and Porto Alegre (both in Brazil), which are used in the package's vignette and examples to illustrate its functionalities.
-Each dataset includes:
+The package includes sample datasets for the cities of São Paulo and Porto Alegre (both in Brazil). Each dataset includes:
 
@@
 
-Calculating a travel time matrix with r5r can be done with the `travel_time_matrix()` function.
-This function takes, as inputs, the spatial location of origins/destinations  (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and a few travel parameters such as maximum trip duration or walking distance.
+The `travel_time_matrix()` function takes, as inputs, the spatial location of origins/destinations  (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and a few travel parameters such as _maximum trip duration_ or _walking distance_, and outputs travel time estimates for each origin-destination pair at a set `departure_datetime`.
 
-By default, `r5r` will return a single travel time estimate for each origin-destination pair at a set `departure_datetime`.
-However, this can provide biased estimates due to how service levels varies across the day (reference to https://www.sciencedirect.com/science/article/pii/S0966692318305209).
-To account
-for these variations, the user can set a `time_window` in the `travel_time_matrix()` function. In this case, several matrices are calculated using multiple departure
-times every minute within the set time window.
+To account for variations on service levels accross the day [@stepniak2019impact], the user can use the `time_window` parameter.
+In this case, several matrices are calculated using one departure time every minute within the set time window.
 By default the function outputs the 50th percentile (i.e.
 the median) travel time between each origin and destination pair  within that time window.
-Nonetheless, it is possible to retrieve information on travel time estimates at different points of the distribution using the `percentiles` parameter.
-The function outputs a `data.table` with the travel time estimates between all combinations of origins and destinations that can be reached with the given parameters.
+The `percentiles` parameter allows the user to retrieve travel time estimates at different points of the distribution.
 An example of the `travel_time_matrix()` function's usage is presented below:
+
+<!-- However, this can provide biased estimates due to how service levels varies across the day (reference to https://www.sciencedirect.com/science/article/pii/S0966692318305209).  -->
+
+<!-- To account for these variations,  -->
 
@@
 
-#### Accounting for travel-time uncertainty
+#### Visualizing travel-time uncertainty
 
-In the example above we calculated how the estimates of travel-time to arrive at the central bus station vary by place of trip origin.
+<!-- In the example above we calculated how the estimates of travel-time to arrive at the central bus station vary by place of trip origin.  -->
 
-We can visualize how the estimated travel times are more uncertain when leaving from some places than others.
-The plot below shows travel times from all origin points to the central bus station at five percentiles (20th, 40th, 60th, 80th, 99th).
-As expected, there is little to no uncertainty in travel times from places that are very close (walking distance) to the central bus station.
-Conversely, travel times from places that are far away are more affected by variations in the trip's departure time and service frequency levels.
+The plot below shows travel times from all origin points to the central bus station at five percentiles (20th, 40th, 60th, 80th, 99th), thus showing how the estimated travel times are more uncertain when leaving from some places than others.
+As expected, there is little to no uncertainty in travel times from places that are very close (walking distance) to the central bus station, while travel times from places that are far away are more affected by variations in the trip's departure time and service frequency levels.
 
@@
 
-#### Visualize Isochrones
+#### Visualizing Isochrones
 
-If the destination points cover the entire study area, the output can easily be used to build ishochrone-like visualizations.
+<!-- If the destination points cover the entire study area, the output can easily be used to build ishochrone-like visualizations.  -->
 In our example, we can visualize the isochrone departing from the central bus station as follows:
@@
 ```{r}
-cutoffs <- seq(0, 120, 20)
+# extract OSM network
+street_net <- street_network_to_sf(r5r_core)
 
-# select the travel time matrix departing from a given origin
-central_bus_stn <- points[291,]
-ttm_iso <- ttm[fromId %in% central_bus_stn$id ]
+travel_times <- ttm[fromId %in% central_bus_stn$id ]
+travel_times[points, on=c('toId' ='id'), `:=`(lon = i.lon, lat = i.lat)]
+travel_times[, mediantt := median(travel_time_p020, travel_time_p040, 
+                                  travel_time_p060, travel_time_p080,
+                                  travel_time_p099), by = toId]
 
-# aggregate travel-times
-ttm_iso[, isochrones := cut(x=travel_time_p099, breaks=cutoffs), by= fromId]
+travel_times.interp <- with(travel_times, interp(lon, lat, mediantt)) %>% 
+  with(cbind(travel_time=as.vector(z),  # Column-major order
+             x=rep(x, times=length(y)),
+             y=rep(y, each=length(x)))) %>% 
+  as.data.frame()
 
-# join travel-matrix results to points
-ttm_iso[points, on=c('toId' ='id'), `:=`(lon = i.lon, lat = i.lat)]
+ggplot(na.omit(travel_times.interp)) +
+  geom_contour_filled(aes(x=x, y=y, z=travel_time)) +
+  geom_sf(data = street_net$edges, color = "gray55", size=0.1, alpha = 0.7) +
+  coord_sf() +
+  labs(fill = "cutoff") +
+  theme_minimal() +
+  theme(axis.title = element_blank())
 
-ggplot(ttm_iso, aes(x=lon, y=lat, colour = isochrones)) +
-   geom_point(size = 2) +
-   geom_point(data = central_bus_stn, aes(x=lon, y=lat), colour="blue", size = 2) +
-   coord_map() +
-   scale_colour_brewer(palette = "Reds") +
-   facet_wrap(~fromId) +
-   theme_minimal() +
-   theme(axis.title = element_blank(),
-         panel.border = element_rect(fill = NA, colour = "grey40"))
 ```
@@
 <!-- realistic routing on multimodal transport networks. -->
+
+<!-- old isochrone code
+# ```{r}
+# cutoffs <- seq(0, 120, 20)
+# 
+# # select the travel time matrix departing from a given origin
+# central_bus_stn <- points[291,]
+# ttm_iso <- ttm[fromId %in% central_bus_stn$id ] 
+# 
+# # aggregate travel-times
+# ttm_iso[, isochrones := cut(x=travel_time_p099, breaks=cutoffs), by= fromId]
+#   
+# # join travel-matrix results to points
+# ttm_iso[points, on=c('toId' ='id'), `:=`(lon = i.lon, lat = i.lat)]
+# 
+# ggplot(ttm_iso, aes(x=lon, y=lat, colour = isochrones)) +
+#    geom_point(size = 2) +
+#    geom_point(data = central_bus_stn, aes(x=lon, y=lat), colour="blue", size = 2) +
+#    coord_map() +
+#    scale_colour_brewer(palette = "Reds") +
+#    facet_wrap(~fromId) +
+#    theme_minimal() +
+#    theme(axis.title = element_blank(),
+#          panel.border = element_rect(fill = NA, colour = "grey40"))
+# ```
+ -->

### 2021-01-15 r-package/paper2020/paper2020.Rmd
minor paper tweaks

@@
 output:
+   html_document: default
    word_document: default
    pdf_document: default
-  html_document: default
 abstract: Routing is a key step in transport planning.
 Nonetheless, researchers and practitioners face challenges to perform this task, such as the costs of licensed software and long computation times.
-Conveyal’s R<sup>5<\sup> is a multimodal transport network router that offers meaningful planning features, such as calculating travel times over a time window and returning multiple itineraries for a single or multiple origin/destination pairs.
-This paper describes r5r, an open source R package that leverages on R<sup>5<\sup> to allow users to efficiently compute travel time matrices and generate detailed itineraries between a set of origins and destinations at no expense.
+Conveyal’s R<sup>5</sup> is a multimodal transport network router that offers meaningful planning features, such as calculating travel times over a time window and returning multiple itineraries for a single or multiple origin/destination pairs.
+This paper describes r5r, an open source R package that leverages on R<sup>5</sup> to allow users to efficiently compute travel time matrices and generate detailed itineraries between a set of origins and destinations at no expense.
 urlcolor: blue
@@
 
-```{r, message = FALSE}
+```{r, message = FALSE, warning = FALSE}
 library(r5r)
@@
 
-To account for variations on service levels accross the day [@stepniak2019impact], the user can use the `time_window` parameter.
-In this case, several matrices are calculated using one departure time every minute within the set time window.
+To account for variations on service levels across the day [@stepniak2019impact], the user can use the `time_window` parameter.
+In this case, several matrices are calculated usinh every minute within the set time window as a departure time.
 By default the function outputs the 50th percentile (i.e.
@@
 The plot below shows travel times from all origin points to the central bus station at five percentiles (20th, 40th, 60th, 80th, 99th), thus showing how the estimated travel times are more uncertain when leaving from some places than others.
-As expected, there is little to no uncertainty in travel times from places that are very close (walking distance) to the central bus station, while travel times from places that are far away are more affected by variations in the trip's departure time and service frequency levels.
+As expected, there is little to no uncertainty in travel times from places that are very close (walking distance) to the central bus station, while travel times from places that are far away are more affected by variations in the departure time and service frequency levels.
 
@@
 plot_data <- melt(ttm_tw, measure = patterns("^travel_time_p"), variable = "percentile", value = "travel_time")
-plot_data[, mediantt := median(travel_time, na.rm=T), by=fromId]
+plot_data[, mediantt := as.integer(median(travel_time, na.rm=T)), by=fromId]
 
@@
 
-travel_times <- ttm[fromId %in% central_bus_stn$id ]
+travel_times <- ttm[fromId %in% central_bus_stn$id]
 travel_times[points, on=c('toId' ='id'), `:=`(lon = i.lon, lat = i.lat)]
@@
 
-The [R](https://github.com/conveyal/r5)<sup>5</sup> routing engine is developed at [Conveyal](https://www.conveyal.com/) with contributions from several developers.
+The [R<sup>5</sup>](https://github.com/conveyal/r5) routing engine is developed at [Conveyal](https://www.conveyal.com/) with contributions from several developers.
 This work was supported by the Brazilian Institute for Applied Economic Research (Ipea).

### 2021-01-18 r-package/paper2020/paper2020.Rmd
add Matt as co-author. Change abstract

@@
 title: 'r5r: rapid realistic routing on multimodal transport networks with R5 in R'
-author: "Rafael H. M. Pereira, Marcus Saraiva, Daniel Herszenhut, Carlos Kaue Braga"
+author: "Rafael H. M. Pereira, Marcus Saraiva, Daniel Herszenhut, Carlos Kaue Braga, Matthew Wigginton Conway"
 date: "`r Sys.Date()`"
@@
    pdf_document: default
-abstract: Routing is a key step in transport planning.
-Nonetheless, researchers and practitioners face challenges to perform this task, such as the costs of licensed software and long computation times.
-Conveyal’s R<sup>5</sup> is a multimodal transport network router that offers meaningful planning features, such as calculating travel times over a time window and returning multiple itineraries for a single or multiple origin/destination pairs.
-This paper describes r5r, an open source R package that leverages on R<sup>5</sup> to allow users to efficiently compute travel time matrices and generate detailed itineraries between a set of origins and destinations at no expense.
+abstract: Routing is a key step in transport planning and research.
+Nonetheless, researchers and practitioners often face challenges to perform this task due to the cost of licensed software and long computation times.
+Conveyal’s R<sup>5</sup> is a multimodal transport network router that offers multiple planning features, such as calculating travel times over a time window, and returning multiple itineraries for origin/destination pairs.
+This paper describes r5r, an open-source R package that leverages on R<sup>5</sup> to allow users to efficiently compute travel time matrices and generate detailed itineraries between sets of origins and destinations at no expense using seamless parallel computing.
 urlcolor: blue
@@
 
-The `travel_time_matrix()` function takes, as inputs, the spatial location of origins/destinations  (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and a few travel parameters such as _maximum trip duration_ or _walking distance_, and outputs travel time estimates for each origin-destination pair at a set `departure_datetime`.
+The `travel_time_matrix()` function takes, as inputs, the spatial location of origins/destinations (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and a few travel parameters such as *maximum trip duration* or *walking distance*, and outputs travel time estimates for each origin-destination pair at a set `departure_datetime`.
 
@@
 <!-- If the destination points cover the entire study area, the output can easily be used to build ishochrone-like visualizations.  -->
+
 In our example, we can visualize the isochrone departing from the central bus station as follows:
@@
 
+```{=html}
 <!-- old isochrone code
@@
  -->
+```

### 2021-01-18 r-package/paper2020/paper2020.Rmd
add referecen to Levinson et al, plus text edits

@@
 
-Transport routing is the process of finding alternative routes and costs that connect places on a given transport network, and it is a key step in transport accessibility analysis, fleet allocation and transport simulation and planning more broadly.
-Researchers and practitioners still frequently face practical challenges when carrying out routing analysis tasks, such as the monetary costs of licensed software, limited data availability for data hungry models, and the long computation times required to run multiple routing scenarios, particularly in large and complex multimodal transport networks.
+Transport routing is the process of finding the alternative routes and costs that connect places on a given transport network, and it is a key step in transport accessibility analysis, fleet allocation and transport simulation and planning more broadly [@levinson2020manual].
+Researchers and practitioners still frequently face practical challenges when carrying out routing analysis tasks due to the monetary costs of licensed software, limited data availability for data hungry models, and the long computation times required to run multiple routing scenarios, particularly in large and complex multimodal transport networks.
 
@@
 This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open source R package for routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5).
-R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to provide an efficient backend for analytic applications - such as accessibility assessment.
+R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to provide an efficient backend for analytic applications - such as accessibility analysis.
 The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R, which allows users to efficiently calculate travel time matrices or generate multiple route alternatives between origins and destinations using seamless parallel computing.
 
-<!-- More importantly, those tools generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) and only return the fastest route option between a set of origins and destinations (see [@conway2018accounting] for a more in-depth discussion).
-This simple approach overlooks how often people's actual route choices deviate from shortest travel time paths, what can occur due to multiple factors such as imperfect information, monetary costs and safety issues [@larsen2011travel; @zhu2015people; @conway2019getting].
--->
+<!-- More importantly, those tools generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) ... 666666666666
 
@@
 
-The r5r package has low data requirements and is easily scalable, allowing fast computation of routes and travel times at either the city or country levels of analysis.
-Street networks from [OpenStreetMap](https://www.openstreetmap.org/) (OSM) are automatically combined with one or multiple public transport data sets provided by the user in the standard [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format to create a routable transport network.
+The r5r package has low data requirements and is easily scalable, allowing fast computation of routes and travel times at either city or country-level analysis.
+It creates a routable transport network using street network data from [OpenStreetMap](https://www.openstreetmap.org/) (OSM), and it  automatically combines public transport data if they are provided by the user in the standard [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format .
 
@@
 
-<!-- Rafa's notes -->
-
-<!-- jean: #>-  mass of assumptions regarding decision-making and routing (Miller, 2018).  -->
-
-<!-- Haugen, 2011; Schwanen, 2008 -->
-
-<!-- Miller, E.J., 2018. Accessibility: measurement and application in transportation planning. Transp. Rev. 38, #>551–555. https://doi.org/10.1080/01441647.2018.1492778 -->
-
-<!-- This article presents **r5r**, a new open-source `R` package for rapid  -->
-
-<!-- realistic routing on multimodal transport networks. -->
-
 ```{=html}

### 2021-01-18 r-package/paper2020/paper2020.Rmd
paper edits, plus fix stepniak citation

@@
 
-<!-- More importantly, those tools generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) ... 666666666666
+<!-- More importantly, those tools generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) ...
+[rafa: the idea here is to emphasize the power of R5 with the time_window parameter.
+We certainly need to rephrase things, but perhaps we should stress this idea here -->
 
@@
 The r5r package has low data requirements and is easily scalable, allowing fast computation of routes and travel times at either city or country-level analysis.
-It creates a routable transport network using street network data from [OpenStreetMap](https://www.openstreetmap.org/) (OSM), and it  automatically combines public transport data if they are provided by the user in the standard [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format .
+It creates a routable transport network using street network data from [OpenStreetMap](https://www.openstreetmap.org/) (OSM), and it automatically combines public transport data if they are provided by the user in the standard [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format.
 
@@
 -   `setup_r5()`: builds a multimodal transport network used for routing in R<sup>5</sup>.
-This function automatically (1) downloads/updates a compiled R<sup>5</sup> JAR file and stores it locally in the r5r package directory for future use; and (2) combines the OSM and GTFS datasets to build a routable network object.
+This function automatically (1) downloads/updates a compiled R<sup>5</sup> JAR file and stores it locally for future use; and (2) combines the OSM and GTFS datasets to build a routable network object.
 
@@
 Both routing functions are versatile so users can easily set customized inputs such as transport modes, departure date and times, walking and cycling speeds, maximum trip duration, walking distances and number of public transport transfers.
-We will focus on results obtained from `travel_time_matrix()` on the findings discussed in the following section.
+In the following section, we will focus on results obtained from `travel_time_matrix()`.
 
@@
 
-After installed with the `install.packages("r5r")` command, the package can be attached (alongside some other packages used in this article), as follows:
+After installed with the `install.packages("r5r")` command, the package can be attached (alongside some other packages to reproduce this article), as follows:
 
@@
 
-### Build a routable transport network
+### Building a routable transport network
 
-To build a routable transport network with r5r and load it in memory, the user needs one command line pointing to the location on disk where the input data is stored.
+To build a routable transport network with r5r and load it in memory, the user need to run a command line pointing to the location on disk where the input data is stored.
 In the examples of this paper, we will be using the provided Porto Alegre dataset.
@@
 
-### Calculate a travel time matrix
+### Calculating a travel time matrix
 
-The `travel_time_matrix()` function takes, as inputs, the spatial location of origins/destinations (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and a few travel parameters such as *maximum trip duration* or *walking distance*, and outputs travel time estimates for each origin-destination pair at a set `departure_datetime`.
+The `travel_time_matrix()` function takes, as inputs, the spatial location of origins/destinations (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and a few travel parameters such as *maximum trip duration* or *walking distance*.
+It outputs travel time estimates for each origin-destination pair at a set `departure_datetime`.
 
-To account for variations on service levels across the day [@stepniak2019impact], the user can use the `time_window` parameter.
-In this case, several matrices are calculated usinh every minute within the set time window as a departure time.
+Since service levels can significantly vary across the day [@stepniak2019impact], users can account for this variation using the `time_window` parameter.
+In this case, several matrices are calculated using every minute within the set time window as a departure time.
 By default the function outputs the 50th percentile (i.e.
-the median) travel time between each origin and destination pair within that time window.
+the median) travel time between each origin and destination pair within the time window.
 The `percentiles` parameter allows the user to retrieve travel time estimates at different points of the distribution.
-An example of the `travel_time_matrix()` function's usage is presented below:
-
-<!-- However, this can provide biased estimates due to how service levels varies across the day (reference to https://www.sciencedirect.com/science/article/pii/S0966692318305209).  -->
-
-<!-- To account for these variations,  -->
+An example of the function's usage is presented below:
 

### 2021-01-18 r-package/paper2020/paper2020.Rmd
paper, edit text travel uncertainty

@@
 
+# routing inputs
 mode <- c("WALK", "TRANSIT")
-max_walk_dist <- 1000
-max_trip_duration <- 120
+max_walk_dist <- 1000 # in meters
+max_trip_duration <- 120 # in minutes
 departure_datetime <- as.POSIXct("13-05-2019 14:00:00",
@@
 
-time_window <- 120
+time_window <- 120 # in minutes
 percentiles <- c(20, 40, 60, 80, 99)
@@
 
-The plot below shows travel times from all origin points to the central bus station at five percentiles (20th, 40th, 60th, 80th, 99th), thus showing how the estimated travel times are more uncertain when leaving from some places than others.
-As expected, there is little to no uncertainty in travel times from places that are very close (walking distance) to the central bus station, while travel times from places that are far away are more affected by variations in the departure time and service frequency levels.
+The plot below shows how the travel times to arrive at the central bus station from several origin points vary within the time window (percentiles 20th, 40th, 60th, 80th, 99th), reflecting that these travel times are more uncertain when leaving from some places than others.
+While there is little to no uncertainty when departing from places that are very close (walking distance) to the central bus station, travel times from farther away places are more affected by variations departure time and service frequency levels.
 
 ```{r, message = FALSE}
-# select the travel time matrix departing from a given origin
+# subset travel time matrix departing from a given origin
 central_bus_stn <- points[291,]
-ttm_tw <- ttm[toId %in% central_bus_stn$id ]
+ttm_tw <- subset(ttm, toId %in% central_bus_stn$id)
 

### 2021-01-18 r-package/paper2020/paper2020.Rmd
paper, edits to text and plot

@@
 ggplot(data=plot_data, aes(y = travel_time, x = reorder(fromId, mediantt))) +
-  geom_point(alpha = 0.1, size = .7) +
+  geom_point(alpha = 0.3, size = .7) +
   geom_line(aes(y=mediantt, group=toId), color="#FE9F45", size=1.5) +
@@
         panel.grid.minor = element_blank(),
-        axis.title.x = element_blank(), 
         axis.text.x = element_blank(),
@@
   labs(title = "Travel time uncertainty per destination",
-       y = "Travel Time (min)", x='Destinations')
+       y = "Travel Time (min)", x='Destinations ordered by median travel time')
 ```
@@
                                    x=rep(x, times=length(y)),
-             y=rep(y, each=length(x)))) %>% 
-  as.data.frame()
+                                   y=rep(y, each=length(x)))) %>% as.data.frame()
 
 ggplot(na.omit(travel_times.interp)) +
-  geom_contour_filled(aes(x=x, y=y, z=travel_time)) +
+  geom_contour_filled(aes(x=x, y=y, z=travel_time), alpha=.7) +
   geom_sf(data = street_net$edges, color = "gray55", size=0.1, alpha = 0.7) +
+  scale_fill_viridis_d(direction = -1, option = 'B') +
   coord_sf() +
-  labs(fill = "cutoff") +
+  labs(fill = "cutoff in minutes") +
   theme_minimal() +

### 2021-01-18 r-package/paper2020/paper2020.Rmd
paper update

--- 2021-01-18 r-package/paper2020/paper2020.Rmd
@@
   labs(title = "Travel time uncertainty per destination",
-       y = "Travel Time (min)", x='Destinations')
+       y = "Travel Time (min)", x='Origins ordered by median travel time')
 ```
@@
                                    x=rep(x, times=length(y)),
-             y=rep(y, each=length(x)))) %>% 
-  as.data.frame()
+                                   y=rep(y, each=length(x)))) %>% as.data.frame()
 
 ggplot(na.omit(travel_times.interp)) +
-  geom_contour_filled(aes(x=x, y=y, z=travel_time)) +
+  geom_contour_filled(aes(x=x, y=y, z=travel_time), alpha=.8) +
   geom_sf(data = street_net$edges, color = "gray55", size=0.1, alpha = 0.7) +
+  scale_fill_viridis_d(direction = -1, option = 'B') +
   coord_sf() +
-  labs(fill = "cutoff") +
+  labs(fill = "cutoff in minutes") +
   theme_minimal() +

### 2021-01-19 r-package/paper2020/paper2020.Rmd
small paper tweaking and comments removing

@@
 ```
-
-<!-- Examples of similar papers -->
-
-<!--  - https://transportfindings.org/article/8416-accessibility-toolbox-for-r-and-arcgis -->
-
-<!--  - https://transportfindings.org/article/6945-dodgr-an-r-package-for-network-flow-aggregation -->
-
-<!--  - https://joss.theoj.org/papers/10.21105/joss.01926 -->
 
@@
 
-<!-- In the example above we calculated how the estimates of travel-time to arrive at the central bus station vary by place of trip origin.  -->
-
 The plot below shows how the travel times to arrive at the central bus station from several origin points vary within the time window (percentiles 20th, 40th, 60th, 80th, 99th), reflecting that these travel times are more uncertain when leaving from some places than others.
-While there is little to no uncertainty when departing from places that are very close (walking distance) to the central bus station, travel times from farther away places are more affected by variations departure time and service frequency levels.
+While there is little to no uncertainty when departing from places that are very close (walking distance) to the central bus station, travel times from places farther away are more affected by departure time variations and service frequency levels.
 
@@
 #### Visualizing Isochrones
-
-<!-- If the destination points cover the entire study area, the output can easily be used to build ishochrone-like visualizations.  -->
 
@@
 # References
-
-```{=html}
-<!-- old isochrone code
-# ```{r}
-# cutoffs <- seq(0, 120, 20)
-# 
-# # select the travel time matrix departing from a given origin
-# central_bus_stn <- points[291,]
-# ttm_iso <- ttm[fromId %in% central_bus_stn$id ] 
-# 
-# # aggregate travel-times
-# ttm_iso[, isochrones := cut(x=travel_time_p099, breaks=cutoffs), by= fromId]
-#   
-# # join travel-matrix results to points
-# ttm_iso[points, on=c('toId' ='id'), `:=`(lon = i.lon, lat = i.lat)]
-# 
-# ggplot(ttm_iso, aes(x=lon, y=lat, colour = isochrones)) +
-#    geom_point(size = 2) +
-#    geom_point(data = central_bus_stn, aes(x=lon, y=lat), colour="blue", size = 2) +
-#    coord_map() +
-#    scale_colour_brewer(palette = "Reds") +
-#    facet_wrap(~fromId) +
-#    theme_minimal() +
-#    theme(axis.title = element_blank(),
-#          panel.border = element_rect(fill = NA, colour = "grey40"))
-# ```
- -->
-```

### 2021-01-25 r-package/paper2020/paper2020.Rmd
paper edits from MWC

@@
 abstract: Routing is a key step in transport planning and research.
-Nonetheless, researchers and practitioners often face challenges to perform this task due to the cost of licensed software and long computation times.
-Conveyal’s R<sup>5</sup> is a multimodal transport network router that offers multiple planning features, such as calculating travel times over a time window, and returning multiple itineraries for origin/destination pairs.
-This paper describes r5r, an open-source R package that leverages on R<sup>5</sup> to allow users to efficiently compute travel time matrices and generate detailed itineraries between sets of origins and destinations at no expense using seamless parallel computing.
+Nonetheless, researchers and practitioners often face challenges when performing this task due to long computation times and the cost of licensed software.
+Conveyal’s R<sup>5</sup> is a multimodal transport network router that offers multiple routing features, such as calculating travel times over a time window and returning multiple itineraries for origin/destination pairs.
+This paper describes r5r, an open-source R package that leverages R<sup>5</sup> to efficiently compute travel time matrices and generate detailed itineraries between sets of origins and destinations at no expense using seamless parallel computing.
 urlcolor: blue
@@
 
-Transport routing is the process of finding the alternative routes and costs that connect places on a given transport network, and it is a key step in transport accessibility analysis, fleet allocation and transport simulation and planning more broadly [@levinson2020manual].
-Researchers and practitioners still frequently face practical challenges when carrying out routing analysis tasks due to the monetary costs of licensed software, limited data availability for data hungry models, and the long computation times required to run multiple routing scenarios, particularly in large and complex multimodal transport networks.
+<!-- edited this because of grammar - costs don't connect places which was grammatically implied in the previous structure -->
+Transport routing is the process of finding the fastest or lowest-cost routes that connect places in a given transport network, and is a key step in transport accessibility analysis, fleet allocation and transport simulation and planning more broadly [@levinson2020manual].
+However, researchers and practitioners often face practical challenges when carrying out routing tasks due to the costs of licensed software, limited data availability, and the long computation times required to run multiple routing scenarios, particularly in large and complex multimodal transport networks.
 
-While there is a growing number of open source routing models [@opentripplanneropentripplanner; @lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently process large transport networks.
-This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open source R package for routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5).
-R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to provide an efficient backend for analytic applications - such as accessibility analysis.
+<!-- MWC: I'm not sure I agree with this - OTP can handle large networks (and OTP 2.0 actually shares a lot of algorithms with R5), and there are also tools like osrm that can handle really large street routing problems.
+Is the problem more that the tools that can handle really large networks are inaccessible to researchers?
+-->
+While there are a growing number of open-source routing packages [@opentripplanneropentripplanner; @lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently process large transport networks.
+This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open-source R package for routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5) package.
+R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to provide an efficient backend for analytic applications, such as accessibility analysis.
 The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R, which allows users to efficiently calculate travel time matrices or generate multiple route alternatives between origins and destinations using seamless parallel computing.
@@
 
-The r5r package has low data requirements and is easily scalable, allowing fast computation of routes and travel times at either city or country-level analysis.
-It creates a routable transport network using street network data from [OpenStreetMap](https://www.openstreetmap.org/) (OSM), and it automatically combines public transport data if they are provided by the user in the standard [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format.
+<!-- Country-level analysis takes a ton of RAM (well, depending on the country)---Maybe we should say region-level instead?
+Like, the Netherlands is possible on high-end consumer hardware, but I suspect that, say, the US would overwhelm most commodity machines -->
+The r5r package has low data requirements and is easily scalable, allowing fast computation of routes and travel times for either city or country-level analysis.
+It creates a routable transport network using street network data from [OpenStreetMap](https://www.openstreetmap.org/) (OSM) and optionally public transport data in the [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format.
 
@@
 
--   `detailed_itineraries()`: computes detailed information on either the shortest or multiple alternative routes between one or multiple origin/destination pairs for a single departure time.
+-   `detailed_itineraries()`: computes detailed information on routes between one or multiple origin/destination pairs for a single departure time.
 The output includes detailed information on route alternatives such as the transport mode, waiting time, travel time and distance of each segment of the trip.
-This function uses an R5-specific extension[^1] to the McRAPTOR [@delling2015roundbased] routing algorithm to find optimal or less than optimal paths.
+This function uses an R<sup>5</sup>-specific extension[^1] to the McRAPTOR [@delling2015roundbased] routing algorithm to find both optimal and slightly suboptimal paths.
 
@@
 
-Both routing functions are versatile so users can easily set customized inputs such as transport modes, departure date and times, walking and cycling speeds, maximum trip duration, walking distances and number of public transport transfers.
+Both routing functions are versatile so users can easily set customized inputs such as transport modes, departure dates and times, walking and cycling speeds, maximum trip duration, walking distances and number of public transport transfers.
 In the following section, we will focus on results obtained from `travel_time_matrix()`.
@@
 
-After installed with the `install.packages("r5r")` command, the package can be attached (alongside some other packages to reproduce this article), as follows:
+After it is installed with the `install.packages("r5r")` command, the package can be attached (alongside other packages to reproduce this article), as follows:
 
@@
 library(akima)
+library(dplyr)
 ```
+
+For this article, we used r5r version XX and R5 version XX. <!-- TODO I'm running r5r from github, but we should use a prod release here. I do think we should include these version numbers -->
 
@@
 
-To build a routable transport network with r5r and load it in memory, the user need to run a command line pointing to the location on disk where the input data is stored.
-In the examples of this paper, we will be using the provided Porto Alegre dataset.
+To build a routable transport network with r5r and load it into memory, the user needs to call `setup_r5` with the path to the directory where OSM and GTFS data are stored.
+In the examples herein, we use the provided Porto Alegre dataset.
 
+<!-- We're getting an error about OSM data here: Continuing to load but ignoring generalized costs due to exception: java.lang.RuntimeException: All ways are expected to have generalized cost tags.
+Missing: slope_1:forward
+Digging through the R5 code though, it looks like this is harmless and is related to some Los Angeles-specific development work done at Conveyal -->
 ```{r, message = FALSE}
@@
 The function uses the `.pbf` and the `GTFS.zip` files in the directory pointed by `data_path` to create a multimodal transport network used for routing by R<sup>5</sup>.
-Multiple GTFS feeds can be used simultaneously, in which case R<sup>5</sup> will automatically merge all files into a single public transport network.
+If multiple GTFS files are present, R<sup>5</sup> will merge them into a single transport network.
 The resulting `network.dat` as well as some other files used by R<sup>5</sup> are saved inside the supplied directory for later reuse.
@@
 
-The `travel_time_matrix()` function takes, as inputs, the spatial location of origins/destinations (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and a few travel parameters such as *maximum trip duration* or *walking distance*.
+The `travel_time_matrix()` function takes, as inputs, the spatial location of origins/destinations (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and a few travel parameters such as *maximum trip duration*, or *walking distance*.
 It outputs travel time estimates for each origin-destination pair at a set `departure_datetime`.
 
-Since service levels can significantly vary across the day [@stepniak2019impact], users can account for this variation using the `time_window` parameter.
-In this case, several matrices are calculated using every minute within the set time window as a departure time.
-By default the function outputs the 50th percentile (i.e.
-the median) travel time between each origin and destination pair within the time window.
-The `percentiles` parameter allows the user to retrieve travel time estimates at different points of the distribution.
-An example of the function's usage is presented below:
+<!-- It might be worthwhile to consider changing the defaults to use a longer time window, and to return more percentiles, so users don't have to set parameters to see variation -MWC -->
+Since service levels can significantly vary across the day [@stepniak2019impact], r5r provides a `time_window` parameter.
+When this parameter is set, R5 will compute travel times for trips at the specified departure time and every minute for `time_window` minutes after.
+The `percentiles` parameter allows the user to retrieve travel time estimates at different points of the distribution (by default the median).
+These percentiles reflect service variation over the time window, but do not reflect schedule deviation not represented in the GTFS, though tools exist to create GTFS which reflects schedule deviations [@wessel2017constructing].
+
+An example of the function's usage is presented below.
+Computing this 1227x1227 travel time matrix with a 120-minute time window takes less than 5 minutes on a MacBook Pro with a 3.1GHz Intel i7 and 16GB RAM.
+<!-- MWC: I think it makes sense to put performance stats here, but it might make sense to read this off of someone else's machine---I have R running inside Docker which may impose some overhead -->
 
@@
 time_window <- 120 # in minutes
-percentiles <- c(20, 40, 60, 80, 99)
+percentiles <- c(20, 40, 50, 60, 80, 99)
 
 # calculate travel time matrix
-ttm <- travel_time_matrix(r5r_core,
+computation_time <- system.time(ttm <- travel_time_matrix(r5r_core,
                           origins = points,
@@
                           percentiles = percentiles,
-                          verbose = FALSE)
-
+                          verbose = FALSE))
+print(paste('travel time matrix computed in', computation_time[['elapsed']], 'seconds'))
 head(ttm)
@@
 
-The plot below shows how the travel times to arrive at the central bus station from several origin points vary within the time window (percentiles 20th, 40th, 60th, 80th, 99th), reflecting that these travel times are more uncertain when leaving from some places than others.
+The plot below shows how the travel times to arrive at the central bus station from several origin points vary within the time window (20th, 40th, 60th, 80th, and 99th percentiles), reflecting that travel times are more uncertain when leaving from some places than others.
 While there is little to no uncertainty when departing from places that are very close (walking distance) to the central bus station, travel times from places farther away are more affected by departure time variations and service frequency levels.
@@
 # reshape data
-plot_data <- melt(ttm_tw, measure = patterns("^travel_time_p"), variable = "percentile", value = "travel_time")
-plot_data[, mediantt := as.integer(median(travel_time, na.rm=T)), by=fromId]
+plot_data <- ttm_tw %>% rename(mediantt=travel_time_p050) %>% melt(measure = patterns("^travel_time_p"), variable = "percentile", value = "travel_time")
 
@@
 
-In our example, we can visualize the isochrone departing from the central bus station as follows:
+In our example, we can visualize the isochrone (area reachable within a certain amount of time) departing from the central bus station as follows:
+
+<!-- Travel time percentile 50 is the median, we shouldn't be calculating it from the other travel times -MWC -->
+<!-- For this map and the next, I couldn't figure it out but there must be a way to crop the map view so only the shaded
+     portion is visible? -MWC -->
 
@@
 travel_times[points, on=c('toId' ='id'), `:=`(lon = i.lon, lat = i.lat)]
-travel_times[, mediantt := median(travel_time_p020, travel_time_p040, 
-                                  travel_time_p060, travel_time_p080,
-                                  travel_time_p099), by = toId]
 
-travel_times.interp <- with(travel_times, interp(lon, lat, mediantt)) %>%
+travel_times.interp <- with(travel_times, interp(lon, lat, travel_time_p050)) %>%
                         with(cbind(travel_time=as.vector(z),  # Column-major order
@@
   geom_sf(data = street_net$edges, color = "gray55", size=0.1, alpha = 0.7) +
+  geom_point(aes(x=lon, y=lat, color='Central bus\nstation'), data=central_bus_stn) +
   scale_fill_viridis_d(direction = -1, option = 'B') +
+  scale_color_manual(values=c('Central bus\nstation'='black')) +
   coord_sf() +
-  labs(fill = "cutoff in minutes") +
+  labs(fill = "travel time (minutes)", color='') +
   theme_minimal() +
@@
 
+```
+
+#### Creating accessibility metrics
+
+<!-- MWC: I added this section, and I think it's helpful to bring this full circle back to accessibility metrics.
+However, right now the accessibility metric is just assuming random numbers of opportunities at each node - we'd want to pull in some real data if we keep the section.
+-->
+
+Accessibility metrics measure the opportunities, such as jobs, a traveler could reach from a particular location [@levinson2020manual].
+One of the simplest forms is a cumulative-opportunities metric, which sums all of the opportunities accessible from each location in less than a cutoff time.
+Using the travel time matrix and information on the number of opportunities available at each location, we can calculate and map accessibility:
+
+```{r}
+# for now just creating random access
+opportunities <- data.frame(toId=unique(ttm$toId), opportunities=runif(length(unique(ttm$toId)), 50, 1000))
+
+access <- ttm %>%
+  left_join(opportunities, by='toId') %>%
+  group_by(fromId) %>%
+  # sum all opportunities accessible within 20 minutes
+  summarize(acc=sum(opportunities[travel_time_p050 <= 20], na.rm=T))
+
+access.interp <- access %>%
+  inner_join(points, by=c('fromId'='id')) %>%
+  with(interp(lon, lat, acc)) %>%
+                        with(cbind(acc=as.vector(z),  # Column-major order
+                                   x=rep(x, times=length(y)),
+                                   y=rep(y, each=length(x)))) %>% as.data.frame()
+
+ggplot(na.omit(access.interp)) +
+  geom_contour_filled(aes(x=x, y=y, z=acc), alpha=.8) +
+  geom_sf(data = street_net$edges, color = "gray55", size=0.1, alpha = 0.7) +
+  scale_fill_viridis_d(direction = -1, option = 'B') +
+  coord_sf() +
+  labs(fill = "Opportunities within\n20 minutes\n(median travel time)") +
+  theme_minimal() +
+  theme(axis.title = element_blank())
 ```

### 2021-01-25 r-package/paper2020/paper2020.Rmd
edits to paper2020 following Matt's suggestions

@@
 
-<!-- edited this because of grammar - costs don't connect places which was grammatically implied in the previous structure -->
 Transport routing is the process of finding the fastest or lowest-cost routes that connect places in a given transport network, and is a key step in transport accessibility analysis, fleet allocation and transport simulation and planning more broadly [@levinson2020manual].
@@
 
-<!-- MWC: I'm not sure I agree with this - OTP can handle large networks (and OTP 2.0 actually shares a lot of algorithms with R5), and there are also tools like osrm that can handle really large street routing problems.
-Is the problem more that the tools that can handle really large networks are inaccessible to researchers?
--->
-While there are a growing number of open-source routing packages [@opentripplanneropentripplanner; @lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently process large transport networks.
+While there are a growing number of open-source routing packages [@opentripplanneropentripplanner; @lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently process large public transport networks and/or are not user-friendly.
 This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open-source R package for routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5) package.
@@
 
-<!-- More importantly, those tools generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) ...
-[rafa: the idea here is to emphasize the power of R5 with the time_window parameter.
-We certainly need to rephrase things, but perhaps we should stress this idea here -->
-
 # METHODS AND DATA
 
-<!-- Country-level analysis takes a ton of RAM (well, depending on the country)---Maybe we should say region-level instead?
-Like, the Netherlands is possible on high-end consumer hardware, but I suspect that, say, the US would overwhelm most commodity machines -->
-The r5r package has low data requirements and is easily scalable, allowing fast computation of routes and travel times for either city or country-level analysis.
+The r5r package has low data requirements and is easily scalable, allowing fast computation of routes and travel times for either city or region-level analysis.
 It creates a routable transport network using street network data from [OpenStreetMap](https://www.openstreetmap.org/) (OSM) and optionally public transport data in the [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format.
@@
 This function uses an R5-specific extension to the RAPTOR routing algorithm which provides an efficient and systematic sampling of multiple simulated schedules when using frequency-based GTFS data [@conway2017evidencebased].
-
-<!-- If the user passes a `time_window` parameter to the function, r5r calculates multiple travel time matrices departing each minute and returns additional information on the percentile distribution of travel times that allows one to grasp service reliability issues.
--->
 
@@
 
-For this article, we used r5r version XX and R5 version XX. <!-- TODO I'm running r5r from github, but we should use a prod release here. I do think we should include these version numbers -->
+For this article, we used r5r version v0.3-2 and R5 version v6.0.1.
 
@@
 -   A public transport network in `GTFS.zip` format.
--   The spatial coordinates of points covering the area in `.csv` format.
+-   The spatial coordinates of points covering the area in `.csv` format, including information
+on the size of resident population and the number of schools in each location.
 
@@
 
-<!-- We're getting an error about OSM data here: Continuing to load but ignoring generalized costs due to exception: java.lang.RuntimeException: All ways are expected to have generalized cost tags.
-Missing: slope_1:forward
-Digging through the R5 code though, it looks like this is harmless and is related to some Los Angeles-specific development work done at Conveyal -->
 ```{r, message = FALSE}
@@
 An example of the function's usage is presented below.
-Computing this 1227x1227 travel time matrix with a 120-minute time window takes less than 5 minutes on a MacBook Pro with a 3.1GHz Intel i7 and 16GB RAM.
-<!-- MWC: I think it makes sense to put performance stats here, but it might make sense to read this off of someone else's machine---I have R running inside Docker which may impose some overhead -->
+Computing this 1227x1227 travel time matrix with a 120-minute time window takes less than two minutes on a Windows machine with a 1.9GHz Intel i7 and 16GB RAM.
 
@@
 time_window <- 120 # in minutes
-percentiles <- c(20, 40, 50, 60, 80, 99)
+percentiles <- c(5, 25, 50, 75, 95)
 
@@
 
-The plot below shows how the travel times to arrive at the central bus station from several origin points vary within the time window (20th, 40th, 60th, 80th, and 99th percentiles), reflecting that travel times are more uncertain when leaving from some places than others.
+The plot below shows how the travel times to arrive at the central bus station from several origin points vary within the time window (20th, 40th, 50th, 60th, 80th, and 99th percentiles), reflecting that travel times are more uncertain when leaving from some places than others.
 While there is little to no uncertainty when departing from places that are very close (walking distance) to the central bus station, travel times from places farther away are more affected by departure time variations and service frequency levels.
@@
 # reshape data
-plot_data <- ttm_tw %>% rename(mediantt=travel_time_p050) %>% melt(measure = patterns("^travel_time_p"), variable = "percentile", value = "travel_time")
+plot_data <- setnames(ttm_tw, 'travel_time_p050', 'mediantt') %>%
+             melt(., measure = patterns("^travel_time_p"), 
+                     variable = "percentile", 
+                     value = "travel_time")
 
+# plot
 ggplot(data=plot_data, aes(y = travel_time, x = reorder(fromId, mediantt))) +
@@
 
-<!-- Travel time percentile 50 is the median, we shouldn't be calculating it from the other travel times -MWC -->
 <!-- For this map and the next, I couldn't figure it out but there must be a way to crop the map view so only the shaded
@@
 
+# select trips departing the bus central station and add coordinates of destinations
 travel_times <- ttm[fromId %in% central_bus_stn$id]
@@
 
-travel_times.interp <- with(travel_times, interp(lon, lat, travel_time_p050)) %>%
+# interpolate estimates to get spatially smooth result
+travel_times.interp <- with(na.omit(travel_times), interp(lon, lat, travel_time_p050)) %>%
                         with(cbind(travel_time=as.vector(z),  # Column-major order
@@
 
+# plot
 ggplot(na.omit(travel_times.interp)) +
@@
 
-<!-- MWC: I added this section, and I think it's helpful to bring this full circle back to accessibility metrics.
-However, right now the accessibility metric is just assuming random numbers of opportunities at each node - we'd want to pull in some real data if we keep the section.
--->
-
 Accessibility metrics measure the opportunities, such as jobs, a traveler could reach from a particular location [@levinson2020manual].
 One of the simplest forms is a cumulative-opportunities metric, which sums all of the opportunities accessible from each location in less than a cutoff time.
-Using the travel time matrix and information on the number of opportunities available at each location, we can calculate and map accessibility:
+Using the travel time matrix and information on the number of opportunities available at each location, we can calculate and map accessibility.
+In the example below we estimate the number of schools accessible by public transport in less than 20 minutes.
 
 ```{r}
-# for now just creating random access
-opportunities <- data.frame(toId=unique(ttm$toId), opportunities=runif(length(unique(ttm$toId)), 50, 1000))
+# merge schools information to travel time matrix
+ttm[points, on=c('toId' ='id'), schools := i.schools]
 
-access <- ttm %>%
-  left_join(opportunities, by='toId') %>%
-  group_by(fromId) %>%
-  # sum all opportunities accessible within 20 minutes
-  summarize(acc=sum(opportunities[travel_time_p050 <= 20], na.rm=T))
+# calculate number of schools accessible
+access <- ttm[travel_time_p050 <= 20, .(acc = sum(schools)), by=fromId]
 
+# interpolate estimates to get spatially smooth result
 access.interp <- access %>%
@@
 
+# plot
 ggplot(na.omit(access.interp)) +
@@
   coord_sf() +
-  labs(fill = "Opportunities within\n20 minutes\n(median travel time)") +
+  labs(fill = "Schools within\n20 minutes\n(median travel time)") +
   theme_minimal() +

### 2021-01-25 r-package/paper2020/paper2020.Rmd
[paper] Fixing isochrone and accessibility plots to remove weird edges at the limits of the road network.

@@
                                    x=rep(x, times=length(y)),
-                                   y=rep(y, each=length(x)))) %>% as.data.frame()
+                                   y=rep(y, each=length(x)))) %>% 
+  as.data.frame() %>% na.omit()
+
+# find isochrone's bounding box to crop the map below
+bb_x <- c(min(travel_times.interp$x), max(travel_times.interp$x))
+bb_y <- c(min(travel_times.interp$y), max(travel_times.interp$y))
 
 # plot
-ggplot(na.omit(travel_times.interp)) +
+ggplot(travel_times.interp) +
   geom_contour_filled(aes(x=x, y=y, z=travel_time), alpha=.8) +
@@
   scale_color_manual(values=c('Central bus\nstation'='black')) +
-  coord_sf() +
+  scale_x_continuous(expand=c(0,0)) +
+  scale_y_continuous(expand=c(0,0)) +
+  coord_sf(xlim = bb_x, ylim = bb_y) +
   labs(fill = "travel time (minutes)", color='') +
@@
   scale_fill_viridis_d(direction = -1, option = 'B') +
-  coord_sf() +
+  scale_x_continuous(expand=c(0,0)) +
+  scale_y_continuous(expand=c(0,0)) +
+  coord_sf(xlim = bb_x, ylim = bb_y) + 
   labs(fill = "Schools within\n20 minutes\n(median travel time)") +

### 2021-01-25 r-package/paper2020/paper2020.Rmd
knit new version of paper2020

@@
 
-<!-- It might be worthwhile to consider changing the defaults to use a longer time window, and to return more percentiles, so users don't have to set parameters to see variation -MWC -->
 Since service levels can significantly vary across the day [@stepniak2019impact], r5r provides a `time_window` parameter.
@@
 In our example, we can visualize the isochrone (area reachable within a certain amount of time) departing from the central bus station as follows:
-
-<!-- For this map and the next, I couldn't figure it out but there must be a way to crop the map view so only the shaded
-     portion is visible? -MWC -->
 

### 2021-01-26 r-package/paper2020/paper2020.Rmd

@@
 ```{r, message = FALSE}
-# points to directory with data
+# system.file returns the directory with example data inside the r5r package
+# set data path to directory containing your own data if not using the examples
 data_path <- system.file("extdata/poa", package = "r5r")
@@
 
-The plot below shows how the travel times to arrive at the central bus station from several origin points vary within the time window (20th, 40th, 50th, 60th, 80th, and 99th percentiles), reflecting that travel times are more uncertain when leaving from some places than others.
+The plot below shows how the travel times to arrive at the central bus station from several origin points vary within the time window (5th, 25th, 50th, 75th, and 95th percentiles), reflecting that travel times are more uncertain when leaving from some places than others.
 While there is little to no uncertainty when departing from places that are very close (walking distance) to the central bus station, travel times from places farther away are more affected by departure time variations and service frequency levels.
@@
 Using the travel time matrix and information on the number of opportunities available at each location, we can calculate and map accessibility.
-In the example below we estimate the number of schools accessible by public transport in less than 20 minutes.
+In the example below we compute the number of schools accessible by public transport in less than 20 minutes.
 

### 2021-01-26 r-package/paper2020/paper2020.Rmd
paper submitted

@@
 ```
+
+*Keywords*: routing, transport networks, travel time, accessibility, travel impedance
 
@@
 ```
+**Code 1: Load required libraries**
 
@@
 ```
+**Code 2: Set up routable transport network**
 
@@
 ```
+**Code 3: Compute travel time matrix**
 
@@
         panel.border = element_rect(fill = NA, colour = "grey80", size=1)) +
-  labs(title = "Travel time uncertainty per destination",
+  labs(title = " ",
        y = "Travel Time (min)", x='Origins ordered by median travel time')
 ```
+**Figure 1: Travel time uncertainty by trip origin**
 
@@
 ```
+**Figure 2: Isochrones by public transport departing from the central bus station**
 
@@
 ```
+**Figure 3: Number of schools accessible by public transport in less than 20 minutes**
 

### 2021-02-24 r-package/paper2020/paper2020.Rmd
update paper with response to reviwers

@@
 While there are a growing number of open-source routing packages [@opentripplanneropentripplanner; @lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently process large public transport networks and/or are not user-friendly.
-This paper presents [r5r](https://ipeagit.github.io/r5r/), a new open-source R package for routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5) package.
+To fill this gap, this paper presents [r5r](https://ipeagit.github.io/r5r/), a new open-source R package for routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5) package.
 R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to provide an efficient backend for analytic applications, such as accessibility analysis.
-The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R, which allows users to efficiently calculate travel time matrices or generate multiple route alternatives between origins and destinations using seamless parallel computing.
+The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R using seamless parallel computing.
+This tool can be used to address a variety of questions that require the efficient calculation of travel matrices or the examination of multimodal transport routes.
 
@@
 
-Since service levels can significantly vary across the day [@stepniak2019impact], r5r provides a `time_window` parameter.
+Since service levels can significantly vary across the day [@stepniak2019impact],
+r5r provides a `time_window` parameter that can help address the aggregation 
+component of the modifiable temporal unit problem (MTUP) [@pereiramtup2019].
 When this parameter is set, R5 will compute travel times for trips at the specified departure time and every minute for `time_window` minutes after.

### 2021-02-24 r-package/paper2020/paper2020.Rmd
add paragraph about 1-1 vs 1-many algorithms

@@
 
-While there are a growing number of open-source routing packages [@opentripplanneropentripplanner; @lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently process large public transport networks and/or are not user-friendly.
+While there are several open-source routing packages available for R, they either do not support public transport networks [@padgham2019dodgr], or focus on providing point-to-point routes rather than origin-destination travel time matrices [@morgan2019opentripplanner; @lovelace2019stplanr].
+Most routing algorithms find paths to all points in the network while finding a single route.
+Storing these paths rather than computing them for one origin-destination pair at a time is orders of magnitude more efficient.
+To our knowledge, no R package exists that supports these efficient many-to-many queries for public transport networks.
+
 To fill this gap, this paper presents [r5r](https://ipeagit.github.io/r5r/), a new open-source R package for routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5) package.

### 2021-02-25 r-package/paper2020/paper2020.Rmd

@@
 
-While there are several open-source routing packages available for R, they either do not support public transport networks [@padgham2019dodgr], or focus on providing point-to-point routes rather than origin-destination travel time matrices [@morgan2019opentripplanner; @lovelace2019stplanr].
+While there are several open-source routing packages available for R, they either do not support public transport networks [@padgham2019dodgr], or primarily focus on providing point-to-point routes rather than origin-destination travel time matrices [@morgan2019opentripplanner; @lovelace2019stplanr].
 Most routing algorithms find paths to all points in the network while finding a single route.

### 2021-03-04 r-package/paper2021/paper2021.Rmd
reorganize paper file on github repo

--- 2021-02-25 r-package/paper2020/paper2020.Rmd
@@
 ```
+
 **Figure 1: Travel time uncertainty by trip origin**
+
+```{r, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
+
+# save plot
+ggsave(file='figure1.jpeg',width = 18, height = 14, units = 'cm', dpi = 400 )
+```
 
@@
 ```
+
 **Figure 2: Isochrones by public transport departing from the central bus station**
+
+```{r, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
+
+# save plot
+ggsave(file='figure2.jpeg',width = 18, height = 14, units = 'cm', dpi = 400 )
+```
 
@@
 ```
+
 **Figure 3: Number of schools accessible by public transport in less than 20 minutes**
+
+```{r, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
+
+# save plot
+ggsave(file='figure3.jpeg',width = 18, height = 14, units = 'cm', dpi = 400 )
+```
 

### 2021-10-19 r-package/paper2021/paper2021.Rmd
Fixed paper Rmd: na.omit() was removing too many rows of the data.frame

@@
 travel_times <- ttm[fromId %in% central_bus_stn$id]
+travel_times <- travel_times[, .(fromId, toId, travel_time_p050)]
 travel_times[points, on=c('toId' ='id'), `:=`(lon = i.lon, lat = i.lat)]

### 2021-10-19 r-package/tests_marcus/test_functions.R

@@ -13,45 +13,45 @@ r5r_core <- setup_r5(data_path = data_path, verbose = TRUE, overwrite = FALSE)
 
 departure_datetime <- as.POSIXct("13-05-2019 14:00:00", format = "%d-%m-%Y %H:%M:%S")
 
-poi <- read.csv(file.path(data_path, "poa_points_of_interest.csv"))
-points <- read.csv(file.path(data_path, "poa_hexgrid.csv"))
+# poi <- read.csv(file.path(data_path, "poa_points_of_interest.csv"))
+# points <- read.csv(file.path(data_path, "poa_hexgrid.csv"))
 
 # points <- points %>%
 #   st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
 #   st_transform(31982)
 
-dest <- points
-
-# points <- r5r_core$getGrid(11L)
-# points <- jdx::convertToR(points)
-# points$schools <- 1
-#
 # dest <- points
+
+points <- r5r_core$getGrid(11L)
+points <- jdx::convertToR(points$getDataFrame())
+points$schools <- 1
+
+dest <- points
 # dest <- dplyr::sample_n(points, 5000)
 
 ## params
 
 # origins = points
 # destinations = dest
-opportunities_colname = "schools"
-mode = "WALK"
-mode_egress = "WALK"
-cutoffs = c(25, 30)
-time_window = 1L
-percentiles = 50L
-decay_function = "step"
-decay_value = 1.0
-max_trip_duration = 30
-max_walk_dist = Inf
-max_bike_dist = Inf
-walk_speed = 3.6
-bike_speed = 12
-max_rides = 3
-max_lts = 2
-n_threads = Inf
-verbose = FALSE
-progress = TRUE
-shortest_path = TRUE
+# opportunities_colname = "schools"
+# mode = "WALK"
+# mode_egress = "WALK"
+# cutoffs = c(25, 30)
+# time_window = 1L
+# percentiles = 50L
+# decay_function = "step"
+# decay_value = 1.0
+# max_trip_duration = 30
+# max_walk_dist = Inf
+# max_bike_dist = Inf
+# walk_speed = 3.6
+# bike_speed = 12
+# max_rides = 3
+# max_lts = 2
+# n_threads = Inf
+# verbose = FALSE
+# progress = TRUE
+# shortest_path = TRUE
 
 
 ## params
@@ -78,81 +78,81 @@ system.time(
                             percentiles = c(25, 50, 75),
                             verbose = FALSE)
 )
-
[…]

### 2022-02-05 r-package/tests_marcus/fare_calculator.Rmd
Added new Fare Calculator functions: `setup_fare_calculator()`, `write_fare_calculator()`, and `read_fare_calculator()`. Added fare calculator to `accessibility()` function.

---
title: "Fare Calculator"
author: "MS"
date: "2/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, include=FALSE}
options(java.parameters = "-Xmx2G")

devtools::load_all(".")
library(sf)
library(data.table)
library(ggplot2)
library(dplyr)
library(mapview)
```

```{r}
# system.file returns the directory with example data inside the r5r package
# set data path to directory containing your own data if not using the examples
data_path <- system.file("extdata/poa", package = "r5r")

r5r_core <- setup_r5(data_path, verbose = FALSE)
```

## Introduction

Considering monetary costs of public transport in travel times and accessibility estimation is a challenge for researchers and planning practitioners. Each public transport system can have its own set of rules for calculating fares, with varying levels of complexity. Fare systems can have features such as: different costs for each leg of travel, depending on transport mode or route; distance- or zone-based fares; different fares for types of riders or time of the day (peak and off-peak, for example); among many others. One common feature among many public transport services is the possibility of integration between modes, when users can get use a single ticket for trip composed of multiple rides. Such trips usually come with a discount in the second or subsequent fares, as well as a limit on the number of discounted rides the user can take and/or a time limit for using that discount. As such, taking all those possible rules into consideration when calculating the monetary cost of multi-modal trips alongside travel times can be quite difficult.

The GTFS format has some features for specifying public transport fares, but those features are quite limited and are not enough for adequately representing many use cases. A new version of that specification is currently being developed (Fares V2 link), but it may take some time for it to be approved and for transport agencies actually start providing GTFS feeds with full fare information.

Meanwhile, R5 has native capabilities and an open architecture for creating and including fare calculators in travel time estimation, making it possible to estimate accessibility considering simultaneously travel time and monetary cost cutoffs. The main challenge, however, is that a specific fare calculator for each city needs to be programmed in Java and tightly integrated into R5. Thus, this is out of reach of many researchers and planners who use R5 through r5r.

To begin tackling this challenge, we are proposing a simple generic rule-based fare calculator that can be configured via a predefined set of properties and rules that can be set directly from R or using external tools such as text editors and spreadsheets. By no means we intend for this to be a robust system that can take into consideration all public transport systems out there. That would be a Herculean task, as can be seen by the long discussions taking place in the Fares V2 forums. The features included in r5r's fare calculator are inspired by our empirical observations of Brazilian public transport systems, and is meant to be used mainly in our own Access to Opportunities project [link]. Everyone else is welcome to use it, if the current features suits their needs.

This vignette aims to show the features of the r5r fare calculator and to demonstrate how to configure it for a specific case. We will be using the sample data set for the city of 
Porto Alegre (Brazil) included in `r5r`.

## The public transport system of Porto Alegre

Porto Alegre has a relatively straightforward public transport system, where the vast majority of the population that rely on transit use buses. The city also has a metropolitan rail service that connects the city center to the neighboring municipalities in the North. That system can be seen in the map below.

```{r}
transit_network <- transit_network_to_sf(r5r_core)

mapview(transit_network$routes, zcol = "mode")
```

The fare rules in the system are as follows:

- Each bus ticket costs R$ 4.80.
- Riding a second bus adds R\$ 2.40 to the total cost. Subsequent bus rides cost the full ticket price of R$ 4.80.
- Each train ticket costs R$ 4.50. Once the user enters a train station, they can take an unlimited amount of trains as long as they do not leave the station. 
- The integrated tariff between bus and train has a 10% discount, which totals R$ 8.37.

In the following sections, we will demonstrate how to implement those rules within r5r's fare calculator.

## Setting up the fare calculator

Three utility functions were added to r5r to allow configuring the fare calculator:

- `setup_fare_calculator()`: analyses the study area's GTFS and builds a 'skeleton' structure of the parameters that need to be set;
- `write_fare_calculator()` and `read_fare_calculator()`: allow saving the current fare calculator settings to disk, and reading them back into memory. The saved settings can be edited outside the R session using external text editors and spreadsheet software, for user's convenience.

First, we need to call `setup_fare_calculator()`, providing three parameters: the current `r5r_core` object, a `base_fare` that represents the default cost of trips that don't fall into the available categories, and the `by` parameters that identifies what is the main property of the route that defines the different fares.

In the example below, the `base_fare` is the standard bus ticket price of R$ 4.80. Prices, however, must be represented as integer values, so here we are saying that the `base_fare` is equivalent to 480 cents. We are also stating that `by = "MODE"`, so that each transport mode has its own fares and integration rules. `by` can also be set to `"AGENCY_ID"` and `"AGENCY_NAME"`, so different rules can be applied to routes ran by different agencies, and to `"GENERIC"` when the entire system follows the same rules.

```{r}
fare_settings <- setup_fare_calculator(r5r_core, 
                                       base_fare = 480,
                                       by = "MODE")
```

Now let's check the contents of the `fare_settings` object. We can see below that it is simply a list with a few properties and data.frames.

```{r}
fare_settings
```

### Global Properties

Let's configure the global properties first, which are the ones that are applied to the entire system. First, we can see that `base_fare` has already been configured to 480, so we leave it as that. `max_discounted_transfers` is set to 1 by default, meaning that the traveler can get 1 integrated discount when riding multiple buses, but after that tickets are charged at full price. Finally, `transfer_time_allowance` is set to 120 minutes, so we have to set it to 60 minutes to fit our use case (travelers have 60 minutes to take the second bus on a discounted fare, otherwise a full fare is charged).

```{r}
fare_settings$base_fare
fare_settings$max_discounted_transfers
fare_settings$transfer_time_allowance <- 60
```

### Transport mode fares configuration

To configure mode-, transfer-, and route-specific properties, we can use the three data.frames inside `fare_settings`. Let's configure the modes first. Below, we can see that the `fare_per_mode` data.frame contains four columns:

- mode: the transport mode to which rules on each row refer to;
- unlimited_transfers: a boolean value that indicates if that transport mode allows unlimited transfers within itself, such as a metro/subway system where the user pays a fare to access a station and then can use as many services as they want as long as they don't exit the system. 
- use_route_fare: another boolean value that indicates if each route will have its own fare, or if all routes in this mode will use the fare indicated in this table.
- fare: the full price fare of this mode.

```{r}
fare_settings$fare_per_mode
```

We need to do two small changes in the `fare_per_mode` table, specifically in the `"RAIL"` mode: setting `unlimited_transfers` to `TRUE`, and updating fare to 450. The `"BUS"` mode does not need any change, so we're leaving it with the default values. We'll do those changes below, using data.table notation.

```{r}
fare_settings$fare_per_mode[mode == "RAIL", unlimited_transfers := TRUE]
fare_settings$fare_per_mode[mode == "RAIL", fare := 450]
```

Checking the results below, everything looks ok:

```{r}
fare_settings$fare_per_mode
```
### Transfer fares configuration

Transfer fares settings are stored in the `fare_per_transfer` data.frame, which is shown below. Each row contains information on how integration between the modes specified in `leg1` and `leg2` work.

```{r}
fare_settings$fare_per_transfer
```

Let's update `fare_per_transfer` to account for the actual integration rules in Porto Alegre.

- The fare for "BUS" to "BUS" integration is composed of 480 for the first leg plus 240 for the second leg, which equals to a total fare of 720. Also, there is a special rule that the discount is not valid when transferring between buses within the same route (for example, from bus T1 to another T1), so we need to set `allow_same_route` to `FALSE`.

```{r}
fare_settings$fare_per_transfer[leg1 == "BUS" & leg2 == "BUS", 
                                allow_same_route := FALSE]

fare_settings$fare_per_transfer[leg1 == "BUS" & leg2 == "BUS", 
                                fare := 720]
```

- Transfers between "RAIL" and "RAIL" are free and unlimited, which is already accounted for in the field `unlimited_transfers` of the `fare_per_mode` table. Thus, the equivalent line of the `fare_per_transfer` data.frame can be completely removed.

```{r}
fare_settings$fare_per_transfer <- 
  fare_settings$fare_per_transfer[!(leg1 == "RAIL" & leg2 == "RAIL")]
```

- Transfers between "BUS" and "RAIL" (in any direction) cost 837, once the 10% discount is applied. Also, `allow_same_route` is irrelevant here, because those transfers are necessarily between different routes. Let's make a final update in the data.frame to account for that.

```{r}
fare_settings$fare_per_transfer[leg1 != leg2, fare := 837]
```

Once all changes are applied, the `fare_per_transfer` data.frame should look like this:

```{r}
fare_settings$fare_per_transfer
```

### Routes configuration

Information on specific routes is stored in the `routes_info` data.frame. Below, we can see a sample of the bus and train routes in Porto Alegre.

```{r}
tail(fare_settings$routes_info)
```

Basic route information is taken directly from the GTFS data (agency, route id and names, mode, etc), but `route_fare` and `fare_type` columns were added specifically for the fare calculator.

- `route_fare`: is used to set a specific fare for each route. This field can be used to represent services that have many unique fares, such as metropolitan / suburban trains and buses. This is used together with the `use_route_fare` column in the `fare_per_mode` table: the `route_fare` field is only read by the fare_calculator when `use_route_fare` of that mode is set to TRUE.

- `fare_type`: is used to link each route with information in the `fare_per_mode` and `fare_per_transfer` tables. In this example, `fare_type` is always the same as `mode`, because that was what we choose when calling `setup_fare_calculator` earlier (we could have chosen to discriminate fares by agency, for example).

We actually don't have any change do to in the `routes_info` table, in this example. It does not matter that the `route_fare` value is wrong for the "RAIL" lines, because we are using the fares set in `fare_per_mode` and `fare_per_transfer` which we already set up correctly before.

Now that our `fare_settings` is complete, we can use it to calculate accessibility with monetary cost cutoffs. Let's see how it's done in the next section.

## Calculating accessibiilty with monetary cost

The `travel_time_matrix()` and `accessibility()` functions have two new parameters to account for monetary costs thresholds:

- `fare_calculator_settings`: the settings object that we've been working on.
- `max_fare`: the maximum total fare that can be used in the trip.

Now, we can answer questions like "how many jobs one can access in 60 minutes using public transport, on a $5 budget?".

```{r}
points_small <- read.csv(system.file("extdata/poa/poa_hexgrid.csv", 
                                     package = "r5r"))
points_small$unit <- 1
departure_datetime <- as.POSIXct("13-05-2019 14:00:00", 
                                 format = "%d-%m-%Y %H:%M:%S")

```

```{r}
r5r_core$setFareCalculatorDebugOutput(
  here::here("tests_marcus", "fare_calculator_output.csv"))
```

```{r}
access_f <- accessibility(r5r_core,
                        origins = points_small,
                        destinations = points_small,
                        departure_datetime = departure_datetime,
                        opportunities_colname = "unit",
                        mode = c("WALK", "TRANSIT"),
                        cutoffs = c(30, 45),
                        fare_calculator_settings = fare_settings,
                        max_fare = 500,
                        max_trip_duration = 45,
                        max_walk_dist = 800,
                        time_window = 1,
                        percentiles = 50,
                        verbose = FALSE)
```

```{r}
debug_output <- fread(here::here("tests_marcus", "fare_calculator_output.csv"))

debug_output[, count := stringr::str_count(debug_output$pattern, 
                                           pattern = "\\|") + 1]
```

```{r}
debug_output[stringr::str_starts(debug_output$pattern, "LINHA")] |> View()
```

```{r}
View(fare_settings$fare_per_transfer)
```

### 2022-02-05 r-package/R/accessibility.R

@@ -241,6 +241,8 @@ accessibility <- function(r5r_core,
                           decay_function = "step",
                           cutoffs = 30L,
                           decay_value = 1.0,
+                          fare_calculator_settings = NULL,
+                          max_fare = Inf,
                           max_walk_dist = Inf,
                           max_bike_dist = Inf,
                           max_trip_duration = 120L,
@@ -322,6 +324,10 @@ accessibility <- function(r5r_core,
   # decay
   decay_list <- assert_decay_function(decay_function, decay_value)
 
+  # max fare
+  checkmate::assert_numeric(max_fare)
+  max_fare <- as.integer(max_fare)
+
   # set r5r_core options ----------------------------------------------------
 
   # time window
@@ -349,6 +355,13 @@ accessibility <- function(r5r_core,
 
   # set progress
   set_progress(r5r_core, progress)
+
+  if (!is.null(fare_calculator_settings)) {
+    fare_settings_json <- jsonlite::toJSON(fare_calculator_settings)
+    json_string <- as.character(fare_settings_json)
+    r5r_core$setFareCalculator(json_string)
+    r5r_core$setMaxFare(max_fare)
+  }
 
 
   # call r5r_core method ----------------------------------------------------

### 2022-02-05 r-package/R/fare_calculator.R

@@ -1,0 +1,153 @@
+#' Title
+#'
+#' @param r5r_core
+#' @param base_fare
+#' @param by
+#'
+#' @return
+#'
+#' @family fare calculator
+#'
+#' @examples
+#'
+#' @export
+setup_fare_calculator <- function(r5r_core,
+                                  base_fare,
+                                  by = "MODE") {
+
+
+  # check inputs ------------------------------------------------------------
+
+  # r5r_core
+  checkmate::assert_class(r5r_core, "jobjRef")
+
+  # max trip duration
+  checkmate::assert_numeric(base_fare, lower = 1)
+  base_fare <- as.integer(base_fare)
+
+  # call r5r_core method ----------------------------------------------------
+
+  f_struct <- r5r_core$buildFareStructure(base_fare, by)
+
+  # collect and convert data returned by r5r_core
+  base_fare <- f_struct$getBaseFare()
+  max_discounted_transfers <- f_struct$getMaxDiscountedTransfers()
+  transfer_time_allowance <- f_struct$getTransferTimeAllowance()
+
+  fare_per_mode <- f_struct$getFarePerModeTable()
+  fare_per_mode <- java_to_dt(fare_per_mode)
+
+  fare_per_transfer <- f_struct$getFarePerTransferTable()
+  fare_per_transfer <- java_to_dt(fare_per_transfer)
+
+  routes_info <- f_struct$getRoutesInfoTable()
+  routes_info <- java_to_dt(routes_info)
+
+  # prepare outputs ----------------------------------------------------
+
+  fare_settings <- list(
+    base_fare = base_fare,
+    max_discounted_transfers = max_discounted_transfers,
+    transfer_time_allowance = transfer_time_allowance,
+    fare_per_mode = fare_per_mode,
+    fare_per_transfer = fare_per_transfer,
+    routes_info = routes_info
+  )
+
+  return(fare_settings)
+
+}
+
+
+#' Write fare calculator settings to file
+#'
+#' @param fare_structure
+#' @param file_path
+#'
+#' @return
+#' @export
+#'
+#' @family fare calculator
+#'
+#' @examples
+write_fare_calculator <- function(fare_structure, file_path) {
+
+  # get temporary folder
+  tmp_dir <- tempdir()
+
+  fare_global_settings <- data.table::data.table(
+    setting = c("base_fare",
+                "max_discounted_transfers",
+                "transfer_time_allowance"),
+    value = c(fare_structure$base_fare,
[…]

### 2022-02-05 r-package/tests_marcus/fares_poa.R

@@ -131,3 +131,13 @@ tn$routes %>% View()
 
 mapview::mapview(tn$routes)
 
+
+## Fare structure -------
+
+f_struct <- fare_calculator_settings(r5r_core, 470, "MODE")
+f_struct$base_fare
+f_struct$max_discounted_transfers
+
+f_struct$fare_per_mode
+f_struct$fare_per_transfer
+f_struct$routes_info

### 2022-02-09 r-package/tests_marcus/fare_calculator.Rmd
Added function `dropFareCalculator()` to R5RCore Added fareCap to FareStructure Atomic values from R are not converted to JSONArrays anymore Fixed bugs in RuleBasedFareCalculator logic Updated Fare Calculator draft vignette

@@
 devtools::load_all(".")
+library(tidyverse)
 library(sf)
@@
 That would be a Herculean task, as can be seen by the long discussions taking place in the Fares V2 forums.
-The features included in r5r's fare calculator are inspired by our empirical observations of Brazilian public transport systems, and is meant to be used mainly in our own Access to Opportunities project [link].
+The features included in r5r's fare calculator are inspired by our empirical observations of Brazilian public transport systems, and is meant to be used mainly in the (Access to Opportunities project)[https://www.ipea.gov.br/acessooportunidades/en/].
 Everyone else is welcome to use it, if the current features suits their needs.
 
-This vignette aims to show the features of the r5r fare calculator and to demonstrate how to configure it for a specific case. We will be using the sample data set for the city of 
-Porto Alegre (Brazil) included in `r5r`.
+This vignette aims to show the features of the r5r fare calculator and to demonstrate how to configure it for a specific case.
+We will be using the sample data set for the city of Porto Alegre (Brazil) included in `r5r`.
 
@@
 
-- Each bus ticket costs R$ 4.80.
-- Riding a second bus adds R\$ 2.40 to the total cost. Subsequent bus rides cost the full ticket price of R$ 4.80.
-- Each train ticket costs R$ 4.50. Once the user enters a train station, they can take an unlimited amount of trains as long as they do not leave the station. 
-- The integrated tariff between bus and train has a 10% discount, which totals R$ 8.37.
+-   Each bus ticket costs R\$ 4.80.
+-   Riding a second bus adds R\$ 2.40 to the total cost. Subsequent bus rides cost the full ticket price of R\$ 4.80.
+-   Each train ticket costs R\$ 4.50. Once the user enters a train station, they can take an unlimited amount of trains as long as they do not leave the station.
+-   The integrated tariff between bus and train has a 10% discount, which totals R\$ 8.37.
 
@@
 
-In the example below, the `base_fare` is the standard bus ticket price of R$ 4.80.
+In the example below, the `base_fare` is the standard bus ticket price of R\$ 4.80.
 Prices, however, must be represented as integer values, so here we are saying that the `base_fare` is equivalent to 480 cents.
@@
 fare_settings$transfer_time_allowance <- 60
+fare_settings$fare_cap
 ```
@@
 We need to do two small changes in the `fare_per_mode` table, specifically in the `"RAIL"` mode: setting `unlimited_transfers` to `TRUE`, and updating fare to 450.
-The `"BUS"` mode does not need any change, so we're leaving it with the default values.
+We also need to change one setting in the `"BUS"` mode.
+There is a discount for transfer between buses (which is set in the following section), but that discount is not valid when transferring between buses within the same route (for example, from bus T1 to another T1).
+We can use the `allow_same_route_transfer` parameter of the `"BUS"` mode for that.
 We'll do those changes below, using data.table notation.
@@
 fare_settings$fare_per_mode[mode == "RAIL", fare := 450]
+fare_settings$fare_per_mode[mode == "BUS", allow_same_route_transfer := FALSE]
 ```
 
-Checking the results below, everything looks ok:
+Checking the results below, everything looks OK:
 
@@
 ```
+
 ### Transfer fares configuration
@@
 -   The fare for "BUS" to "BUS" integration is composed of 480 for the first leg plus 240 for the second leg, which equals to a total fare of 720.
-Also, there is a special rule that the discount is not valid when transferring between buses within the same route (for example, from bus T1 to another T1), so we need to set `allow_same_route` to `FALSE`.
 
 ```{r}
-fare_settings$fare_per_transfer[leg1 == "BUS" & leg2 == "BUS", 
-                                allow_same_route := FALSE]
-
 fare_settings$fare_per_transfer[leg1 == "BUS" & leg2 == "BUS", 
@@
 -   Transfers between "RAIL" and "RAIL" are free and unlimited, which is already accounted for in the field `unlimited_transfers` of the `fare_per_mode` table.
-Thus, the equivalent line of the `fare_per_transfer` data.frame can be completely removed.
+Thus, the equivalent row of the `fare_per_transfer` data.frame needs to be removed.
+If we leave the that row in `fare_per_transfer`, transfers between "RAIL" and "RAIL" will count to the global `max_discounted_transfers` allowance.
 
@@
 
-Now, we can answer questions like "how many jobs one can access in 60 minutes using public transport, on a $5 budget?".
+Now, we can answer questions like "how many jobs one can access in 60 minutes using public transport, on a \$5 budget?".
 
@@
 
-```
+# helper function to calculate accessibility with common settings
+calculate_accessibility <- function(settings, fare) {
+  if (!is.null(settings)) {
+    debug_file <- here::here("tests_marcus/debug_poa", 
+                             paste0("debug_fare_", fare, ".csv" ))
+    settings$debug_settings$output_file = debug_file
+  }
 
-```{r}
-r5r_core$setFareCalculatorDebugOutput(
-  here::here("tests_marcus", "fare_calculator_output.csv"))
-```
-
-```{r}
-access_f <- accessibility(r5r_core,
+  access_df <- accessibility(r5r_core,
                         origins = points_small,
@@
                         mode = c("WALK", "TRANSIT"),
-                        cutoffs = c(30, 45),
-                        fare_calculator_settings = fare_settings,
-                        max_fare = 500,
-                        max_trip_duration = 45,
+                        cutoffs = c(60),
+                        fare_calculator_settings = settings,
+                        max_fare = fare,
+                        max_trip_duration = 60,
                         max_walk_dist = 800,
@@
                         verbose = FALSE)
+}
 ```
@@
 ```{r}
-debug_output <- fread(here::here("tests_marcus", "fare_calculator_output.csv"))
-
-debug_output[, count := stringr::str_count(debug_output$pattern, 
-                                           pattern = "\\|") + 1]
+access_unl <- calculate_accessibility(NULL, 0)
+access_300 <- calculate_accessibility(fare_settings, 300)
+access_500 <- calculate_accessibility(fare_settings, 500)
+access_900 <- calculate_accessibility(fare_settings, 900)
 ```
@@
 ```{r}
-debug_output[stringr::str_starts(debug_output$pattern, "LINHA")] |> View()
+access_unl$max_fare <- "unl"
+access_300$max_fare <- "3.0"
+access_500$max_fare <- "5.0"
+access_900$max_fare <- "9.0"
+
+access_df <- rbind(access_unl, access_300, access_500, access_900)
 ```
@@
 ```{r}
-View(fare_settings$fare_per_transfer)
+access_df$geometry <- h3jsr::h3_to_polygon(access_df$from_id)
+access_df <- st_as_sf(access_df, crs = 4326)
 ```
+
+```{r}
+access_df %>%
+  filter(cutoff == 60) %>%
+  ggplot() +
+  geom_sf(aes(fill=accessibility), color = NA) +
+  scale_fill_distiller(palette = "Spectral") +
+  facet_wrap(~max_fare)
+```
+
+```{r}
+debug_df <- read_csv(here::here("tests_marcus/debug_poa", "debug_fare_300.csv"))
+debug_df <- read_csv(here::here("tests_marcus/debug_poa", "debug_fare_500.csv"))
+debug_df <- read_csv(here::here("tests_marcus/debug_poa", "debug_fare_900.csv"))
+```
+
+```{r}
+access_df %>%
+  st_set_geometry(NULL) %>%
+  pivot_wider(names_from = max_fare, values_from = accessibility) %>%
+  janitor::clean_names()
+```
+
+```{r}
+write_fare_calculator(fare_settings, file_path = here::here("teste_fares_poa.zip"))
+fare_settings_2 <- read_fare_calculator(file_path = here::here("teste_fares_poa.zip"))
+```

### 2022-02-09 java-r5rcore/src/org/ipea/r5r/Fares/FareStructure.java

@@ -6,8 +6,6 @@ import org.json.simple.JSONObject;
 import org.json.simple.parser.JSONParser;
 import org.json.simple.parser.ParseException;
 
-import java.util.Iterator;
-
 public class FareStructure {
 
     /**
@@ -26,6 +24,11 @@ public class FareStructure {
      * Maximum time allowed for discounted transfers, counting from the time of  boarding first transit leg
     */
     private int transferTimeAllowance;
+
+    /**
+     * Maximum charged fare for any combination of trips.
+     */
+    private int fareCap;
 
     private final RDataFrame farePerModeTable;
     private final RDataFrame farePerTransferTable;
@@ -56,6 +59,14 @@ public class FareStructure {
         this.transferTimeAllowance = transferTimeAllowance;
     }
 
+    public int getFareCap() {
+        return fareCap;
+    }
+
+    public void setFareCap(int fareCap) {
+        this.fareCap = fareCap;
+    }
+
     public RDataFrame getFarePerModeTable() {
         return farePerModeTable;
     }
@@ -74,17 +85,18 @@ public class FareStructure {
         this.baseFare = fare;
         this.maxDiscountedTransfers = 1;
         this.transferTimeAllowance = 120;
+        this.fareCap = -1;
 
         this.farePerModeTable = new RDataFrame();
         this.farePerModeTable.addStringColumn("mode", "BUS");
         this.farePerModeTable.addBooleanColumn("unlimited_transfers", false);
+        this.farePerModeTable.addBooleanColumn("allow_same_route_transfer", true);
         this.farePerModeTable.addBooleanColumn("use_route_fare", false);
         this.farePerModeTable.addIntegerColumn("fare", this.baseFare);
 
         this.farePerTransferTable = new RDataFrame();
         this.farePerTransferTable.addStringColumn("leg1", "BUS");
         this.farePerTransferTable.addStringColumn("leg2", "BUS");
-        this.farePerTransferTable.addBooleanColumn("allow_same_route", true);
         this.farePerTransferTable.addIntegerColumn("fare", this.baseFare);
 
         this.routesInfoTable = new RDataFrame();
@@ -106,15 +118,16 @@ public class FareStructure {
             JSONObject json = (JSONObject) parser.parse(data);
 
             // setting global properties
-            // obs: single values come from R as arrays
-            JSONArray tmpArray = (JSONArray) json.get("base_fare");
-            fareStructure.setBaseFare(((Long) tmpArray.get(0)).intValue());
[…]

### 2022-02-09 java-r5rcore/src/org/ipea/r5r/Fares/RuleBasedInRoutingFareCalculator.java

@@ -22,6 +22,7 @@ import java.util.concurrent.ConcurrentSkipListSet;
 class ModeData {
     String mode;
     boolean unlimitedTransfers;
+    boolean allowTransferToSameRoute;
     boolean useRouteFare;
     int fare;
 
@@ -30,6 +31,7 @@ class ModeData {
         return "ModeData{" +
                 "mode='" + mode + '\'' +
                 ", unlimitedTransfers=" + unlimitedTransfers +
+                ", allowTransferTooSameRoute=" + allowTransferToSameRoute +
                 ", useRouteFare=" + useRouteFare +
                 ", fare=" + fare +
                 '}';
@@ -39,7 +41,6 @@ class ModeData {
 class TransferData {
     String leg1;
     String leg2;
-    boolean allowTransferToSameRoute;
     int fare;
 
     @Override
@@ -47,7 +48,6 @@ class TransferData {
         return "TransferData{" +
                 "leg1='" + leg1 + '\'' +
                 ", leg2='" + leg2 + '\'' +
-                ", allowTransferTooSameRoute=" + allowTransferToSameRoute +
                 ", fare=" + fare +
                 '}';
     }
@@ -73,6 +73,7 @@ public class RuleBasedInRoutingFareCalculator extends InRoutingFareCalculator {
     private static final org.slf4j.Logger LOG = LoggerFactory.getLogger(RuleBasedInRoutingFareCalculator.class);
 
     public static String debugFileName = "";
+    public static String debugTripInfo = "";
     public static boolean debugActive = false;
 
     private final FareStructure fareStructure;
@@ -109,6 +110,7 @@ public class RuleBasedInRoutingFareCalculator extends InRoutingFareCalculator {
             ModeData data = new ModeData();
             data.mode = fareStructure.getFarePerModeTable().getStringValue("mode");
             data.unlimitedTransfers = fareStructure.getFarePerModeTable().getBooleanValue("unlimited_transfers");
+            data.allowTransferToSameRoute = fareStructure.getFarePerModeTable().getBooleanValue("allow_same_route_transfer");
             data.useRouteFare = fareStructure.getFarePerModeTable().getBooleanValue("use_route_fare");
             data.fare = fareStructure.getFarePerModeTable().getIntValue("fare");
 
@@ -130,7 +132,6 @@ public class RuleBasedInRoutingFareCalculator extends InRoutingFareCalculator {
             TransferData data = new TransferData();
[…]

### 2022-02-09 java-r5rcore/src/org/ipea/r5r/R5RCore.java

@@ -159,6 +159,10 @@ public class R5RCore {
         this.routingProperties.setFareCalculatorJson(fareCalculatorSettingsJson);
     }
 
+    public void dropFareCalculator() {
+        this.routingProperties.setFareCalculator("");
+    }
+
     public void setFareCalculatorDebugOutput(String fileName) {
         RuleBasedInRoutingFareCalculator.debugFileName = fileName;
         RuleBasedInRoutingFareCalculator.debugActive = !fileName.equals("");

### 2022-02-09 r-package/R/accessibility.R

@@ -357,10 +357,13 @@ accessibility <- function(r5r_core,
   set_progress(r5r_core, progress)
 
   if (!is.null(fare_calculator_settings)) {
-    fare_settings_json <- jsonlite::toJSON(fare_calculator_settings)
+    fare_settings_json <- jsonlite::toJSON(fare_calculator_settings, auto_unbox = TRUE)
     json_string <- as.character(fare_settings_json)
     r5r_core$setFareCalculator(json_string)
     r5r_core$setMaxFare(max_fare)
+  } else {
+    r5r_core$dropFareCalculator()
+    r5r_core$setMaxFare(0L)
   }
 
 

### 2022-02-09 r-package/R/fare_calculator.R

@@ -33,6 +33,8 @@ setup_fare_calculator <- function(r5r_core,
   base_fare <- f_struct$getBaseFare()
   max_discounted_transfers <- f_struct$getMaxDiscountedTransfers()
   transfer_time_allowance <- f_struct$getTransferTimeAllowance()
+  fare_cap <- f_struct$getFareCap()
+  if (fare_cap <= 0) { fare_cap <- Inf }
 
   fare_per_mode <- f_struct$getFarePerModeTable()
   fare_per_mode <- java_to_dt(fare_per_mode)
@@ -43,15 +45,24 @@ setup_fare_calculator <- function(r5r_core,
   routes_info <- f_struct$getRoutesInfoTable()
   routes_info <- java_to_dt(routes_info)
 
+  debug <- list(
+    output_file = "",
+    trip_info = "MODE"
+  )
+
   # prepare outputs ----------------------------------------------------
 
   fare_settings <- list(
     base_fare = base_fare,
     max_discounted_transfers = max_discounted_transfers,
     transfer_time_allowance = transfer_time_allowance,
+    fare_cap = fare_cap,
+
     fare_per_mode = fare_per_mode,
     fare_per_transfer = fare_per_transfer,
-    routes_info = routes_info
+    routes_info = routes_info,
+
+    debug_settings = debug
   )
 
   return(fare_settings)
@@ -78,14 +89,24 @@ write_fare_calculator <- function(fare_structure, file_path) {
   fare_global_settings <- data.table::data.table(
     setting = c("base_fare",
                 "max_discounted_transfers",
-                "transfer_time_allowance"),
+                "transfer_time_allowance",
+                "fare_cap"),
     value = c(fare_structure$base_fare,
               fare_structure$max_discounted_transfers,
-              fare_structure$transfer_time_allowance)
+              fare_structure$transfer_time_allowance,
+              fare_structure$fare_cap)
+  )
+
+  fare_debug_settings <- data.table::data.table(
+    setting = c("output_file",
+                "trip_info"),
+    value = c(fare_structure$debug_settings$output_file,
+              fare_structure$debug_settings$trip_info)
   )
 
   data.table::fwrite(x = fare_global_settings,
                      file = file.path(tmp_dir, "global_settings.csv"))
+
 
   data.table::fwrite(x = fare_structure$fare_per_mode,
                      file = file.path(tmp_dir, "fare_per_mode.csv"))
@@ -96,12 +117,16 @@ write_fare_calculator <- function(fare_structure, file_path) {
[…]

### 2022-02-09 r-package/tests_marcus/fares_rio.R

@@ -9,6 +9,7 @@ library(h3jsr)
 library(data.table)
 library(viridis)
 library(patchwork)
+library(googlesheets4)
 
 # build transport network
 data_path <- "~/Repos/r5r_fares/rio"
@@ -32,8 +33,13 @@ points <- read_csv("~/Repos/r5r_fares/rio/points_rio_09_2019.csv") %>%
   mutate(unit = 1)
 departure_datetime <- as.POSIXct("13-05-2019 14:00:00", format = "%d-%m-%Y %H:%M:%S")
 
-area_sf <- h3jsr::h3_to_polygon(points$id, simple = FALSE) %>%
-  summarise()
+points_r8 <- h3jsr::get_parent(points$id, res = 8, simple = TRUE)
+points_r8 <- unique(points_r8)
+points_r8 <- h3jsr::h3_to_point(points_r8, simple = FALSE)
+points_r8$id <- points_r8$h3_address
+points_r8$unit <- 1
+# area_sf <- h3jsr::h3_to_polygon(points$id, simple = FALSE) %>%
+#   summarise()
 
 
 # Setup fare calculator ---------------------------------------------------
@@ -42,35 +48,45 @@ fare_settings <- setup_fare_calculator(r5r_core,
                                        base_fare = 405,
                                        by = "MODE")
 
-write_fare_calculator(fare_settings, file_path = here::here("tests_marcus", "fares_rio_base.zip"))
+
+write_fare_calculator(fare_settings, file_path = here::here("tests_marcus", "rio_fares_v2.zip"))
+fare_settings <- read_fare_calculator(file_path = here::here("tests_marcus", "rio_fares_v2.zip"))
+
+
 
 fare_settings$base_fare
 fare_settings$max_discounted_transfers
 fare_settings$transfer_time_allowance
+fare_settings$fare_cap
 fare_settings$fare_per_mode %>% View()
 fare_settings$fare_per_transfer %>% View()
 fare_settings$routes_info %>% View()
+fare_settings$debug_settings
 
-### Fares per mode --------------------------------------------------------
 
-fare_settings <- read_fare_calculator(file_path = here::here("tests_marcus", "rio_fares_v1.zip"))
+clipr::write_clip()
+clipr::write_clip(fare_settings$fare_per_mode)
+clipr::write_clip(fare_settings$fare_per_transfer)
+clipr::write_clip(fare_settings$routes_info)
 
-json <- jsonlite::toJSON(fare_settings)
[…]

### 2022-02-21 r-package/tests_marcus/fare_calculator.Rmd
Many changes in fare calculator feature: - fares now accept float values, which are automatically converted to integer for R5 compatibility; - JSON serialisation is now made using Jackson; - removed support for other fare calculators (only RuleBasedFareCalculator is allowed) - function `set_fare_calculator()` created in fare_calculator.R, to pass fare_calculator_settings from R to Java - added fare calculator feature to `travel_time_matrix()` - lots of code cleanup

@@
 fare_settings <- setup_fare_calculator(r5r_core, 
-                                       base_fare = 480,
+                                       base_fare = 4.8,
                                        by = "MODE")
@@
 ```{r}
-fare_settings$fare_per_mode
+fare_settings$fares_per_mode
 ```
@@
 ```{r}
-fare_settings$fare_per_mode[mode == "RAIL", unlimited_transfers := TRUE]
-fare_settings$fare_per_mode[mode == "RAIL", fare := 450]
-fare_settings$fare_per_mode[mode == "BUS", allow_same_route_transfer := FALSE]
+fare_settings$fares_per_mode[mode == "RAIL", unlimited_transfers := TRUE]
+fare_settings$fares_per_mode[mode == "RAIL", fare := 4.5]
+fare_settings$fares_per_mode[mode == "BUS", allow_same_route_transfer := FALSE]
 ```
@@
 ```{r}
-fare_settings$fare_per_transfer
+fare_settings$fares_per_transfer
 ```
@@
 ```{r}
-fare_settings$fare_per_transfer[leg1 == "BUS" & leg2 == "BUS", 
-                                fare := 720]
+fare_settings$fares_per_transfer[first_leg == "BUS" & second_leg == "BUS", 
+                                fare := 7.2]
 ```
@@
 ```{r}
-fare_settings$fare_per_transfer <- 
-  fare_settings$fare_per_transfer[!(leg1 == "RAIL" & leg2 == "RAIL")]
+fare_settings$fares_per_transfer <- 
+  fare_settings$fares_per_transfer[!(first_leg == "RAIL" & second_leg == "RAIL")]
 ```
@@
 ```{r}
-fare_settings$fare_per_transfer[leg1 != leg2, fare := 837]
+fare_settings$fares_per_transfer[first_leg != second_leg, fare := 8.37]
 ```
@@
 ```{r}
-tail(fare_settings$routes_info)
+tail(fare_settings$fares_per_route)
 ```
@@
 calculate_accessibility <- function(settings, fare) {
+
   if (!is.null(settings)) {
-    debug_file <- here::here("tests_marcus/debug_poa", 
-                             paste0("debug_fare_", fare, ".csv" ))
-    settings$debug_settings$output_file = debug_file
+    settings$debug_settings$output_file = 
+      here::here("tests_marcus", paste0("debug_fare_", fare, ".csv" ))
   }
+
+  # if (!is.null(settings)) {
+  #   debug_file <- here::here("tests_marcus/debug_poa", 
+  #                            )
+  #   settings$debug_settings$output_file = debug_file
+  # }
 
@@
 access_unl <- calculate_accessibility(NULL, 0)
-access_300 <- calculate_accessibility(fare_settings, 300)
-access_500 <- calculate_accessibility(fare_settings, 500)
-access_900 <- calculate_accessibility(fare_settings, 900)
+access_300 <- calculate_accessibility(fare_settings, 3.00)
+access_500 <- calculate_accessibility(fare_settings, 5.00)
+access_900 <- calculate_accessibility(fare_settings, 9.00)
+```
+
+```{r}
+clipr::write_clip(r5r_core$getFareStructure())
+clipr::write_clip(r5r_core$getFareCalculatorDebugOutputSettings())
+r5r_core$setFareCalculatorDebugOutputSettings(here::here("tests_marcus", "dbg2.csv"),
+                                              "MODE")
 ```

### 2022-02-21 java-r5rcore/src/org/ipea/r5r/Fares/FarePerMode.java

@@ -1,11 +1,15 @@
 package org.ipea.r5r.Fares;
 
+import com.fasterxml.jackson.annotation.JsonIgnore;
+
 public class FarePerMode {
     private String mode;
-    private boolean unlimitedTransfers = false;
-    private boolean allowSameRouteTransfer = false;
-    private boolean useRouteFare = false;
-    private float fare = 0.0f;
+    private boolean unlimitedTransfers;
+    private boolean allowSameRouteTransfer;
+    private boolean useRouteFare;
+    private float fare;
+    @JsonIgnore
+    private int integerFare;
 
     public FarePerMode(String mode, boolean unlimitedTransfers, boolean allowSameRouteTransfer, boolean useRouteFare, float fare) {
         this.mode = mode;
@@ -13,6 +17,7 @@ public class FarePerMode {
         this.allowSameRouteTransfer = allowSameRouteTransfer;
         this.useRouteFare = useRouteFare;
         this.fare = fare;
+        this.integerFare = Math.round(fare * 100.0f);
     }
 
     public FarePerMode() {
@@ -55,8 +60,13 @@ public class FarePerMode {
         return fare;
     }
 
+    public int getIntegerFare() {
+        return integerFare;
+    }
+
     public void setFare(float fare) {
         this.fare = fare;
+        this.integerFare = Math.round(fare * 100.0f);
     }
 
     @Override

### 2022-02-21 java-r5rcore/src/org/ipea/r5r/Fares/FarePerRoute.java

@@ -1,6 +1,8 @@
 package org.ipea.r5r.Fares;
 
-public class RouteInfo {
+import com.fasterxml.jackson.annotation.JsonIgnore;
+
+public class FarePerRoute {
 
     private String agencyId;
     private String agencyName;
@@ -10,9 +12,11 @@ public class RouteInfo {
     private String mode;
     private float routeFare;
     private String fareType;
+    @JsonIgnore
+    private int integerFare;
 
-    public RouteInfo(String agencyId, String agencyName, String routeId, String routeShortName, String routeLongName,
-                     String mode, float routeFare, String fareType) {
+    public FarePerRoute(String agencyId, String agencyName, String routeId, String routeShortName, String routeLongName,
+                        String mode, float routeFare, String fareType) {
         this.agencyId = agencyId;
         this.agencyName = agencyName;
         this.routeId = routeId;
@@ -20,10 +24,12 @@ public class RouteInfo {
         this.routeLongName = routeLongName;
         this.mode = mode;
         this.routeFare = routeFare;
+        this.integerFare = Math.round(routeFare * 100.0f);
         this.fareType = fareType;
     }
 
-    public RouteInfo() {
+    public FarePerRoute() {
+        this("", "", "", "", "", "", 0.0f, "GENERIC");
     }
 
     public String getAgencyId() {
@@ -78,8 +84,13 @@ public class RouteInfo {
         return routeFare;
     }
 
+    public int getIntegerFare() {
+        return this.integerFare;
+    }
+
     public void setRouteFare(float routeFare) {
         this.routeFare = routeFare;
+        this.integerFare = Math.round(routeFare * 100.0f);
     }
 
     public String getFareType() {
@@ -92,7 +103,7 @@ public class RouteInfo {
 
     @Override
     public String toString() {
-        return "RouteInfo{" +
+        return "FarePerRoute{" +
                 "agencyId='" + agencyId + '\'' +
                 ", agencyName='" + agencyName + '\'' +
                 ", routeId='" + routeId + '\'' +

### 2022-02-21 java-r5rcore/src/org/ipea/r5r/Fares/FarePerTransfer.java

@@ -1,14 +1,19 @@
 package org.ipea.r5r.Fares;
 
+import com.fasterxml.jackson.annotation.JsonIgnore;
+
 public class FarePerTransfer {
     private String firstLeg;
     private String secondLeg;
     private float fare;
+    @JsonIgnore
+    private int integerFare;
 
     public FarePerTransfer(String firstLeg, String secondLeg, float fare) {
         this.firstLeg = firstLeg;
         this.secondLeg = secondLeg;
         this.fare = fare;
+        this.integerFare = Math.round(fare * 100.0f);
     }
 
     public FarePerTransfer() {
@@ -35,8 +40,13 @@ public class FarePerTransfer {
         return fare;
     }
 
+    public int getIntegerFare() {
+        return this.integerFare;
+    }
+
     public void setFare(float fare) {
         this.fare = fare;
+        this.integerFare = Math.round(fare * 100.0f);
     }
 
     @Override

### 2022-02-21 java-r5rcore/src/org/ipea/r5r/Fares/FareStructure.java

@@ -1,17 +1,24 @@
 package org.ipea.r5r.Fares;
 
-import org.ipea.r5r.RDataFrame;
-import org.json.simple.JSONArray;
-import org.json.simple.JSONObject;
-import org.json.simple.parser.JSONParser;
-import org.json.simple.parser.ParseException;
+import com.fasterxml.jackson.annotation.JsonIgnore;
+import com.fasterxml.jackson.annotation.JsonIgnoreProperties;
+import com.fasterxml.jackson.core.JsonProcessingException;
+import com.fasterxml.jackson.databind.ObjectMapper;
+import com.fasterxml.jackson.databind.PropertyNamingStrategy;
+import java.util.ArrayList;
+import java.util.HashMap;
+import java.util.List;
+import java.util.Map;
 
+@JsonIgnoreProperties(ignoreUnknown = true)
 public class FareStructure {
 
     /**
      * When fares are not found in farePerMode or farePerTransfer, the value of baseFare is used.
      */
-    private int baseFare;
+    private float baseFare;
+    @JsonIgnore
+    private int integerBaseFare;
 
     /**
      * Maximum number of transfers that can have a discount. The transfer fare is obtained from the farePerTransfer
@@ -30,17 +37,22 @@ public class FareStructure {
      */
     private int fareCap;
 
-    private final RDataFrame farePerModeTable;
-    private final RDataFrame farePerTransferTable;
-    private final RDataFrame routesInfoTable;
+    private final List<FarePerMode> faresPerMode;
+    private final List<FarePerTransfer> faresPerTransfer;
+    private final List <FarePerRoute> faresPerRoute;
 
     // Getters and Setters
-    public int getBaseFare() {
+    public float getBaseFare() {
         return baseFare;
     }
 
-    public void setBaseFare(int baseFare) {
+    public int getIntegerBaseFare() {
+        return integerBaseFare;
+    }
+
+    public void setBaseFare(float baseFare) {
         this.baseFare = baseFare;
+        this.integerBaseFare = Math.round(baseFare * 100.0f);
     }
 
     public int getMaxDiscountedTransfers() {
@@ -67,50 +79,62 @@ public class FareStructure {
         this.fareCap = fareCap;
     }
 
-    public RDataFrame getFarePerModeTable() {
-        return farePerModeTable;
+    public List<FarePerMode> getFaresPerMode() {
+        return faresPerMode;
     }
-    public RDataFrame getFarePerTransferTable() {
[…]

### 2022-02-21 java-r5rcore/src/org/ipea/r5r/Fares/FareStructureBuilder.java

@@ -3,9 +3,9 @@ package org.ipea.r5r.Fares;
 import com.conveyal.r5.transit.RouteInfo;
 import com.conveyal.r5.transit.TransitLayer;
 import com.conveyal.r5.transit.TransportNetwork;
-import org.ipea.r5r.RDataFrame;
 
 import java.util.HashSet;
+import java.util.List;
 import java.util.Set;
 
 public class FareStructureBuilder {
@@ -21,7 +21,7 @@ public class FareStructureBuilder {
         this.transportNetwork = transportNetwork;
     }
 
-    public FareStructure build(int baseFare, String routeType) {
+    public FareStructure build(float baseFare, String routeType) {
         this.fareStructure = new FareStructure(baseFare);
 
         populateFareStructure(routeType);
@@ -33,36 +33,40 @@ public class FareStructureBuilder {
         Set<String> modes = new HashSet<>();
         Set<String> transfers = new HashSet<>();
 
-        RDataFrame routes = this.fareStructure.getRoutesInfoTable();
+        List<FarePerRoute> routes = this.fareStructure.getFaresPerRoute();
         for (RouteInfo route : transportNetwork.transitLayer.routes) {
 
-            routes.append();
-            routes.set("agency_id", route.agency_id);
-            routes.set("agency_name", route.agency_name);
-            routes.set("route_id", route.route_id);
-            routes.set("route_short_name", route.route_short_name);
-            routes.set("route_long_name", route.route_long_name);
-            routes.set("mode", TransitLayer.getTransitModes(route.route_type).toString());
+            FarePerRoute newRoute = new FarePerRoute();
+
+            newRoute.setAgencyId(route.agency_id);
+            newRoute.setAgencyName(route.agency_name);
+            newRoute.setRouteId(route.route_id);
+            newRoute.setRouteShortName(route.route_short_name);
+            newRoute.setRouteLongName(route.route_long_name);
+            newRoute.setMode(TransitLayer.getTransitModes(route.route_type).toString());
+            newRoute.setRouteFare(this.fareStructure.getBaseFare());
 
             switch (type) {
                 case "MODE":
-                    routes.set("fare_type", TransitLayer.getTransitModes(route.route_type).toString());
+                    newRoute.setFareType(TransitLayer.getTransitModes(route.route_type).toString());
                     modes.add(TransitLayer.getTransitModes(route.route_type).toString());
                     break;
                 case "AGENCY":
                 case "AGENCY_ID":
[…]

### 2022-02-21 java-r5rcore/src/org/ipea/r5r/Fares/RuleBasedInRoutingFareCalculator.java

@@ -15,74 +15,25 @@ import java.util.Map;
 import java.util.Set;
 import java.util.concurrent.ConcurrentSkipListSet;
 
-/*
- * support data classes
- */
-
-class ModeData {
-    String mode;
-    boolean unlimitedTransfers;
-    boolean allowTransferToSameRoute;
-    boolean useRouteFare;
-    int fare;
-
-    @Override
-    public String toString() {
-        return "ModeData{" +
-                "mode='" + mode + '\'' +
-                ", unlimitedTransfers=" + unlimitedTransfers +
-                ", allowTransferTooSameRoute=" + allowTransferToSameRoute +
-                ", useRouteFare=" + useRouteFare +
-                ", fare=" + fare +
-                '}';
-    }
-}
-
-class TransferData {
-    String leg1;
-    String leg2;
-    int fare;
-
-    @Override
-    public String toString() {
-        return "TransferData{" +
-                "leg1='" + leg1 + '\'' +
-                ", leg2='" + leg2 + '\'' +
-                ", fare=" + fare +
-                '}';
-    }
-}
-
-class RouteInfoData {
-    String routeId;
-    String mode;
-    int fare;
-
-    @Override
-    public String toString() {
-        return "RouteInfoData{" +
-                "routeId='" + routeId + '\'' +
-                ", mode='" + mode + '\'' +
-                ", fare=" + fare +
-                '}';
-    }
-}
-
 public class RuleBasedInRoutingFareCalculator extends InRoutingFareCalculator {
 
     private static final org.slf4j.Logger LOG = LoggerFactory.getLogger(RuleBasedInRoutingFareCalculator.class);
 
     public static String debugFileName = "";
-    public static String debugTripInfo = "";
+    public static String debugTripInfo = "MODE";
     public static boolean debugActive = false;
+
+    public FareStructure getFareStructure() {
+        return fareStructure;
+    }
 
     private final FareStructure fareStructure;
 
-    private Map<String, RouteInfoData> routeInfo;
-    private Map<String, ModeData> farePerMode;
-    private Map<String, TransferData> farePerTransfer;
+    private Map<String, FarePerRoute> routeInfo;
+    private Map<String, FarePerMode> farePerMode;
+    private Map<String, FarePerTransfer> farePerTransfer;
 
-    private Set<String> debugOutput;
+    private final Set<String> debugOutput;
 
     public Set<String> getDebugOutput() {
[…]

### 2022-02-21 java-r5rcore/src/org/ipea/r5r/ParetoFrontierCalculator.java

@@ -22,10 +22,10 @@ public class ParetoFrontierCalculator  extends R5MultiDestinationProcess {
         RegionalTask request = buildRequest(index);
         TravelTimeComputer computer = new TravelTimeComputer(request, transportNetwork);
 
-        Map<Integer, OneOriginResult> travelTimeResults = new HashMap<>();
+        Map<Float, OneOriginResult> travelTimeResults = new HashMap<>();
 
-        for (int fareCutoff : this.routingProperties.fareCutoffs) {
-            request.maxFare = fareCutoff;
+        for (float fareCutoff : this.routingProperties.fareCutoffs) {
+            request.maxFare = Math.round(fareCutoff * 100.0f);
             OneOriginResult results = computer.computeTravelTimes();
 
             travelTimeResults.put(fareCutoff, results);
@@ -49,13 +49,13 @@ public class ParetoFrontierCalculator  extends R5MultiDestinationProcess {
         travelTimesTable.addStringColumn("to_id", "");
         travelTimesTable.addIntegerColumn("percentile", 0);
         travelTimesTable.addIntegerColumn("travel_time", 0);
-        travelTimesTable.addIntegerColumn("monetary_cost", 0);
-        travelTimesTable.addIntegerColumn("monetary_cost_upper", 0);
+        travelTimesTable.addDoubleColumn("monetary_cost", 0.0);
+        travelTimesTable.addDoubleColumn("monetary_cost_upper", 0.0);
 
         return travelTimesTable;
     }
 
-    private void populateDataFrame(Map<Integer, OneOriginResult> travelTimeResults, RDataFrame travelTimesTable) {
+    private void populateDataFrame(Map<Float, OneOriginResult> travelTimeResults, RDataFrame travelTimesTable) {
         for (int destination = 0; destination < destinationPoints.featureCount(); destination++) {
             for (int percentileIndex = 0; percentileIndex < this.routingProperties.percentiles.length; percentileIndex++) {
                 // add new row to data frame
@@ -67,20 +67,20 @@ public class ParetoFrontierCalculator  extends R5MultiDestinationProcess {
 
                 boolean first = true;
                 int previousTT = -1;
-                for (int fare : this.routingProperties.fareCutoffs) {
+                for (float fare : this.routingProperties.fareCutoffs) {
                     OneOriginResult travelTimesByFare = travelTimeResults.get(fare);
                     int tt = travelTimesByFare.travelTimes.getValues()[percentileIndex][destination];
 
[…]

### 2022-02-21 java-r5rcore/src/org/ipea/r5r/R5Process.java

@@ -198,6 +198,9 @@ public abstract class R5Process {
 
         request.percentiles = routingProperties.percentiles;
 
+        request.inRoutingFareCalculator = routingProperties.fareCalculator;
+        request.maxFare = Math.round(routingProperties.maxFare * 100.0f);
+
         return request;
     }
 }

### 2022-02-21 java-r5rcore/src/org/ipea/r5r/RoutingProperties.java

@@ -2,9 +2,6 @@ package org.ipea.r5r;
 
 import com.conveyal.r5.analyst.cluster.PathResult;
 import com.conveyal.r5.analyst.fare.*;
-import com.conveyal.r5.analyst.fare.nyc.NYCInRoutingFareCalculator;
-import org.ipea.r5r.Fares.PortoAlegreInRoutingFareCalculator;
-import org.ipea.r5r.Fares.RioDeJaneiroInRoutingFareCalculator;
 import org.ipea.r5r.Fares.RuleBasedInRoutingFareCalculator;
 
 public class RoutingProperties {
@@ -21,40 +18,9 @@ public class RoutingProperties {
     public boolean travelTimesBreakdown;
     public PathResult.Stat travelTimesBreakdownStat;
 
-    public int maxFare;
-    public int[] fareCutoffs = {-1};
+    public float maxFare;
+    public float[] fareCutoffs = {-1.0f};
     public InRoutingFareCalculator fareCalculator;
-
-    public void setFareCalculator(String fareCalculatorName) {
-
-        switch (fareCalculatorName) {
-            case "boston":
-                this.fareCalculator = new BostonInRoutingFareCalculator();
-                break;
-            case "bogota":
-                this.fareCalculator = new BogotaInRoutingFareCalculator();
-                break;
-            case "chicago":
-                this.fareCalculator = new ChicagoInRoutingFareCalculator();
-                break;
-            case "simple":
-                this.fareCalculator = new SimpleInRoutingFareCalculator();
-                break;
-            case "bogota-mixed":
-                this.fareCalculator = new BogotaMixedInRoutingFareCalculator();
-                break;
-            case "nyc":
-                this.fareCalculator = new NYCInRoutingFareCalculator();
-                break;
-            case "porto-alegre":
-                this.fareCalculator = new PortoAlegreInRoutingFareCalculator();
-                break;
-            case "rio-de-janeiro":
-                this.fareCalculator = new RioDeJaneiroInRoutingFareCalculator();
-                break;
-            default: this.fareCalculator = null;
-        }
-    }
 
     public void setFareCalculatorJson(String fareCalculatorJson) {
         this.fareCalculator = new RuleBasedInRoutingFareCalculator(fareCalculatorJson);
@@ -72,7 +38,7 @@ public class RoutingProperties {
         travelTimesBreakdown = false;
         travelTimesBreakdownStat = PathResult.Stat.MEAN;
[…]

### 2022-02-21 r-package/R/pareto_frontier.R

@@ -265,10 +265,6 @@ pareto_frontier <- function(r5r_core,
   checkmate::assert_numeric(percentiles)
   percentiles <- as.integer(percentiles)
 
-  # monetary costs
-  checkmate::assert_numeric(monetary_cost_cutoffs)
-  monetary_cost_cutoffs <- as.integer(monetary_cost_cutoffs)
-
   # set r5r_core options ----------------------------------------------------
 
   # time window
@@ -296,15 +292,7 @@ pareto_frontier <- function(r5r_core,
   set_progress(r5r_core, progress)
 
   # fare calculator settings
-  if (!is.null(fare_calculator_settings)) {
-    fare_settings_json <- jsonlite::toJSON(fare_calculator_settings, auto_unbox = TRUE)
-    json_string <- as.character(fare_settings_json)
-    r5r_core$setFareCalculator(json_string)
-    r5r_core$setFareCutoffs(monetary_cost_cutoffs)
-  } else {
-    r5r_core$dropFareCalculator()
-    r5r_core$setFareCutoffs(0L)
-  }
+  set_fare_calculator(r5r_core, fare_calculator_settings, fare_cutoffs = monetary_cost_cutoffs)
 
   # call r5r_core method ----------------------------------------------------
 

### 2022-02-21 r-package/tests_marcus/fares_poa.R

@@ -134,10 +134,42 @@ mapview::mapview(tn$routes)
 
 ## Fare structure -------
 
-f_struct <- fare_calculator_settings(r5r_core, 470, "MODE")
+
+
+f_struct <- setup_fare_calculator(r5r_core, 4.7, "MODE")
+
+View(f_struct$fares_per_mode)
+
 f_struct$base_fare
 f_struct$max_discounted_transfers
 
 f_struct$fare_per_mode
 f_struct$fare_per_transfer
 f_struct$routes_info
+
+access <- accessibility(r5r_core,
+                        origins = points_small,
+                        destinations = points_small,
+                        departure_datetime = departure_datetime,
+                        opportunities_colname = "unit",
+                        mode = c("WALK", "TRANSIT"),
+                        cutoffs = c(30, 45),
+                        fare_calculator_settings = f_struct,
+                        max_fare = 10,
+                        max_trip_duration = 45,
+                        max_walk_dist = 800,
+                        time_window = 1,
+                        percentiles = 50,
+                        verbose = FALSE) %>%
+  left_join(points, by = c("from_id" = "id"))
+
+
+access %>%
+  filter(cutoff == 30) %>%
+  drop_na() %>%
+  ggplot(aes(x=lon, y=lat)) +
+  geom_point(size=1, aes(color=accessibility)) +
+  geom_point(data=poi[c(1, 10),], color = "blue") +
+  facet_grid(~cutoff) +
+  scale_color_distiller(palette = "Spectral") +
+  coord_map()

### 2022-05-11 r-package/tests/tests_rafa/test_rafa.R
temporary fix of pareto_frontier vignette

@@ -657,3 +657,19 @@ tt <- r5r::travel_time_matrix(r5r_core,
                               origins = vertices_buff,
                               destinations = city_center_df,
                               mode = 'walk')
+
+
+
+vignete pareto
+library(tibble)
+
+df <- tibble::tribble(~option, ~modes, ~time, ~cost,
+                      1, 'walk',       50,    0,
+                      2, 'bus',        35,    4,
+                      3, 'bus + bus',  29,    6,
+                      4, 'subway',     15,    8)
+
+
+ggplot() +
+        geom_path(data=df, aes(x=cost, y=time)) +
+        geom_point(data=df, aes(x=cost, y=time))

### 2022-05-15 r-package/tests/tests_rafa/test_rafa.R
1st full version of pareto_frontier vignette

@@ -551,14 +551,20 @@ r5r_core <- setup_r5(path)
 
 ### CMD Check ----------------
 # Check package errors
+library(tictoc)
 
 # LOCAL
+tic()
 Sys.setenv(NOT_CRAN = "true")
 devtools::check(pkg = ".",  cran = FALSE, env_vars = c(NOT_CRAN = "true"))
+toc()
 
 # CRAN
+tic()
 Sys.setenv(NOT_CRAN = "false")
 devtools::check(pkg = ".",  cran = TRUE, env_vars = c(NOT_CRAN = "false"))
+toc()
+
 
 devtools::check_win_release(pkg = ".")
 

### 2022-05-16 r-package/tests/tests_rafa/test_rafa.R
fix pareto_frontier.Rmd

@@ -505,6 +505,10 @@ Sys.setenv(NOT_CRAN = "true")
 r5r_cov2 <- covr::package_coverage(path = ".", type = "tests")
 r5r_cov
 
+as.data.frame(r5r_cov2)[, c(1:3, 5, 11)]
+
+zeroCov <- covr::zero_coverage(r5r_cov2)
+
 
 x <- as.data.frame(r5r_cov)
 covr::codecov( coverage = r5r_cov, token ='2a7013e9-6562-4011-beb9-168e922c4c84' )

### 2022-05-19 r-package/tests/tests_marcus/test_elevation.R
Updated package dependencies: added gtfstools, removed tibble Added eval = FALSE to last chunk of time window vignette

@@ -6,21 +6,18 @@ library(data.table)
 library(tidyverse)
 
 # build transport network
-data_path <- system.file("extdata/poa", package = "r5r")
-r5r_core <- setup_r5(data_path = data_path, verbose = TRUE, overwrite = FALSE,
-                     temp_dir = FALSE,
-                     use_elevation = FALSE,
-                     use_native_elevation = FALSE)
+data_path <- system.file("extdata/spo", package = "r5r")
+r5r_core <- setup_r5(data_path = data_path, verbose = TRUE, overwrite = TRUE,
+                     temp_dir = FALSE)
 
 # load origin/destination points
-points <- read.csv(file.path(data_path, "poa_hexgrid.csv"))
+points <- read.csv(file.path(data_path, "spo_hexgrid.csv"))
 
 # r5r_core$setTravelTimesBreakdown(FALSE)
 ttm_flat_walk <- travel_time_matrix(r5r_core,
                             origins = points,
                             destinations = points,
-                            breakdown = FALSE,
-                            mode = c("WALK"),
+                            mode = c("WALK", "TRANSIT"),
                             max_trip_duration = 60,
                             max_walk_dist = Inf,
                             time_window = 1,

### 2022-06-21 r-package/tests/tests_rafa/test_rafa.R
Closes #280

@@ -573,7 +573,6 @@ tictoc::toc()
 
 
 
-
 # CRAN
 tictoc::tic()
 Sys.setenv(NOT_CRAN = "false")

### 2022-11-24 r-package/R/pareto_frontier.R
closes #268

@@ -20,7 +20,7 @@
 #'   combination. Due to upstream restrictions, only 5 percentiles can be
 #'   specified at a time. For more details, please see R5 documentation at
 #'   <https://docs.conveyal.com/analysis/methodology#accounting-for-variability>.
-#' @param monetary_cost_cutoffs A numeric vector. The monetary cutoffs that
+#' @param fare_cutoffs A numeric vector. The monetary cutoffs that
 #'   should be considered when calculating the Pareto frontier. Most of the
 #'   time you'll want this parameter to be the combination of all possible
 #'   fares listed in you `fare_structure`. Choosing a coarse distribution of
@@ -80,7 +80,7 @@
 #'   mode = c("WALK", "TRANSIT"),
 #'   departure_datetime = departure_datetime,
 #'   fare_structure = fare_structure,
-#'   monetary_cost_cutoffs = c(4.5, 4.8, 9, 9.3, 9.6)
+#'   fare_cutoffs = c(4.5, 4.8, 9, 9.3, 9.6)
 #' )
 #' head(pf)
 #'
@@ -99,7 +99,7 @@ pareto_frontier <- function(r5r_core,
                             max_car_time = Inf,
                             max_trip_duration = 120L,
                             fare_structure = NULL,
-                            monetary_cost_cutoffs = -1L,
+                            fare_cutoffs = -1L,
                             walk_speed = 3.6,
                             bike_speed = 12,
                             max_rides = 3,
@@ -163,7 +163,7 @@ pareto_frontier <- function(r5r_core,
   set_progress(r5r_core, progress)
   set_fare_structure(r5r_core, fare_structure)
   set_output_dir(r5r_core, output_dir)
-  set_monetary_cutoffs(r5r_core, monetary_cost_cutoffs)
+  set_fare_cutoffs(r5r_core, fare_cutoffs)
 
   # call r5r_core method and process result -------------------------------
 

### 2022-11-24 r-package/R/set.R

@@ -398,23 +398,23 @@ set_cutoffs <- function(r5r_core, cutoffs, decay_function) {
 #' Pareto frontier.
 #'
 #' @template r5r_core
-#' @param monetary_cutoffs A path.
+#' @param fare_cutoffs A path.
 #'
 #' @return Invisibly returns `TRUE`.
 #'
 #' @family setting functions
 #'
 #' @keywords internal
-set_monetary_cutoffs <- function(r5r_core, monetary_cutoffs) {
+set_fare_cutoffs <- function(r5r_core, fare_cutoffs) {
   checkmate::assert_numeric(
-    monetary_cutoffs,
+    fare_cutoffs,
     lower = 0,
     any.missing = FALSE,
     min.len = 1,
     unique = TRUE
   )
 
-  r5r_core$setFareCutoffs(rJava::.jfloat(monetary_cutoffs))
+  r5r_core$setFareCutoffs(rJava::.jfloat(fare_cutoffs))
 
   return(invisible(TRUE))
 }

### 2022-11-24 r-package/tests/tests_rafa/test_rafa.R

@@ -261,7 +261,7 @@ cat(speeds_js)
 # export json to path file
 jsonlite::write_json(speeds_js, path = paste0('build-configsssss.json'))
 
-}
+
 
 
 
@@ -509,6 +509,9 @@ covr::function_coverage(fun=r5r::detailed_itineraries, test_file("tests/testthat
 a <- covr::function_coverage(fun=r5r::expanded_travel_time_matrix, test_file("tests/testthat/test-expanded_travel_time_matrix.R"))
 a <- covr::function_coverage(fun=r5r::pareto_frontier, test_file("tests/testthat/test-pareto_frontier.R"))
 a <- covr::function_coverage(fun=r5r::accessibility, test_file("tests/testthat/test-accessibility.R"))
+
+a <- covr::function_coverage(fun=r5r:::set_verbose, test_file("tests/testthat/test-set_verbose.R"))
+
 
 covr::function_coverage(fun=r5r::street_network_to_sf, test_file("tests/testthat/test-street_network_to_sf.R"))
 covr::function_coverage(fun=r5r::transit_network_to_sf, test_file("tests/testthat/test-transit_network_to_sf.R"))

### 2022-11-24 r-package/tests/testthat/test-pareto_frontier.R

@@ -14,7 +14,7 @@ default_tester <- function(r5r_core,
                            time_window = 1L,
                            percentiles = 50,
                            fare_structure = NULL,
-                           monetary_cost_cutoffs = 0L,
+                           fare_cutoffs = 0L,
                            max_walk_time = Inf,
                            max_bike_time = Inf,
                            max_trip_duration = 120L,
@@ -34,7 +34,7 @@ default_tester <- function(r5r_core,
     time_window = time_window,
     percentiles = percentiles,
     fare_structure = fare_structure,
-    monetary_cost_cutoffs = monetary_cost_cutoffs,
+    fare_cutoffs = fare_cutoffs,
     max_walk_time = max_walk_time,
     max_bike_time = max_bike_time,
     max_trip_duration = max_trip_duration,

### 2022-11-24 r-package/tests/testthat/test-set_max_lts.R

@@ -54,7 +54,7 @@ test_that("max_lts argument works in pareto_frontier()", {
     mode = \"BICYCLE\",
     departure_datetime = departure_datetime,
     fare_structure = fare_structure,
-    monetary_cost_cutoffs = 0
+    fare_cutoffs = 0
   )"
 
   fast_expr <- sub("\\)$", ", max_lts = 4\\)", expr)

### 2022-11-24 r-package/tests/testthat/test-set_max_rides.R

@@ -89,7 +89,7 @@ test_that("max_rides argument works in expanded_travel_time_matrix()", {
         } else {
           s
         }
-      } 
+      }
     )
     count <- vapply(walkless_vector, length, integer(1))
   }
@@ -112,7 +112,7 @@ test_that("max_rides argument works in pareto_frontier()", {
     mode = c('WALK', 'TRANSIT'),
     departure_datetime = departure_datetime,
     fare_structure = fare_structure,
-    monetary_cost_cutoffs = c(0, 5, 10)
+    fare_cutoffs  = c(0, 5, 10)
   )"
 
   one_ride_expr <- sub("\\)$", ", max_rides = 1\\)", expr)
@@ -145,7 +145,7 @@ test_that("max_rides argument works in detailed_itineraries()", {
 
   one_ride <- eval(parse(text = one_ride_expr))
   many_rides <- eval(parse(text = many_rides_expr))
-  
+
   count_transit <- function(routes) {
     non_walk_legs <- routes != ""
     sum(non_walk_legs)

### 2022-11-24 r-package/tests/testthat/test-set_monte_carlo_draws.R

@@ -271,7 +271,7 @@ test_that("draws_per_minute arg works in pareto_frontier() - poa", {
     time_window = 30,
     percentiles = c(1, 50, 99),
     fare_structure = fare_structure,
-    monetary_cost_cutoffs = c(0, 5, 10)
+    fare_cutoffs = c(0, 5, 10)
   )
 
   small_draws_expr <- big_draws_expr <- basic_expr
@@ -296,7 +296,7 @@ test_that("draws_per_minute arg works in pareto_frontier() - poa", {
 #     time_window = 30,
 #     percentiles = c(1, 50, 99),
 #     fare_structure = spo_fare_struc,
-#     monetary_cost_cutoffs = c(0, 5, 10, 15, Inf)
+#     fare_cutoffs = c(0, 5, 10, 15, Inf)
 #   )
 #
 #   small_draws_expr <- big_draws_expr <- basic_expr

### 2022-11-24 r-package/tests/testthat/test-set_percentiles.R

@@ -114,7 +114,7 @@ test_that("set_percentiles argument works in pareto_frontier()", {
     mode = c("TRANSIT", "WALK"),
     departure_datetime = departure_datetime,
     fare_structure = fare_structure,
-    monetary_cost_cutoffs = c(0, 5, 10),
+    fare_cutoffs = c(0, 5, 10),
     draws_per_minute = 1,
     time_window = 30,
     max_trip_duration = 60

### 2022-11-24 r-package/tests/testthat/test-set_progress.R

@@ -26,7 +26,7 @@ test_that("progress argument works in routing functions", {
         pois[1:5],
         departure_datetime = departure_datetime,
         fare_structure = fare_structure,
-        monetary_cost_cutoffs = c(0, 5, 10),
+        fare_cutoffs = c(0, 5, 10),
         max_trip_duration = 60
       )"
     } else if (identical(f, accessibility)) {

### 2022-11-24 r-package/tests/testthat/test-set_speed.R

@@ -143,7 +143,7 @@ test_that("set_speed argument works in pareto_frontier()", {
     points[1:5],
     mode = "WALK",
     fare_structure = fare_structure,
-    monetary_cost_cutoffs = 0
+    fare_cutoffs = 0
   )
 
   slow_walk_expr <- fast_walk_expr <- walk_expr

### 2022-11-24 r-package/tests/testthat/test-set_time_window.R

@@ -148,7 +148,7 @@ test_that("set_time_window argument works in pareto_frontier()", {
     draws_per_minute = 1,
     percentiles = percentiles,
     fare_structure = fare_structure,
-    monetary_cost_cutoffs = c(0, 5, 10),
+    fare_cutoffs = c(0, 5, 10),
     max_trip_duration = 60
   )
 

### 2022-11-24 r-package/tests/testthat/test-set_verbose.R

@@ -26,7 +26,7 @@ test_that("verbose argument works in routing functions", {
         pois[1:5],
         departure_datetime = departure_datetime,
         fare_structure = fare_structure,
-        monetary_cost_cutoffs = c(0, 5, 10),
+        fare_cutoffs = c(0, 5, 10),
         max_trip_duration = 60
       )"
     } else if (identical(f, accessibility)) {

### 2022-11-24 r-package/tests/tests_rafa/test_rafa.R
0 errors v | 0 warnings

@@ -348,10 +348,7 @@ ttm <- travel_time_matrix(r5r_core,
  mode <- c('car', 'BUS')
  mode <- c('car', 'BUS', 'walk', 'BICYCLE')
 
- bu kike === walk
- se só tem direct mode, outros vazios
 
- mode_list <- select_mode(mode)
 
 
 
@@ -554,8 +551,7 @@ zeroCov <- covr::zero_coverage(a)
 x <- as.data.frame(r5r_cov)
 covr::codecov( coverage = r5r_cov, token ='2a7013e9-6562-4011-beb9-168e922c4c84' )
 
-Error (test-street_network_to_sf.R:14:3): street_network_to_sf - expected behavior
-Error (test-detailed_itineraries.R:169:3): detailed_itineraries output is correct
+
 
 
 
@@ -583,16 +579,6 @@ system("R CMD Rd2pdf --title=Package gtfs2gps --output=./gtfs2gps/manual.pdf")
 
 
 
-1. Failure: detailed_itineraries output is correct (@test-detailed_itineraries.R#182)
-2. Failure: detailed_itineraries output is correct (@test-detailed_itineraries.R#202)
-
-pdflatex
-
-
-
-path <- 'E:/Dropbox/other_projects/0_jean_capability/opentripplanner/otp_got'
-options(java.parameters = "-Xmx16G")
-r5r_core <- setup_r5(path)
 
 ### CMD Check ----------------
 # Check package errors

### 2022-12-03 r-package/tests/tests_rafa/test_rafa.R
test ttmatrix vignette

@@ -26,31 +26,198 @@ mapviewOptions(platform = 'leafgl')
 
 
 
-##### INPUT  ------------------------
+############## issue 281 - fare calculator not working with Sao Paulo example ------------------------------
 
-utils::remove.packages('r5r')
-devtools::install_github("ipeaGIT/r5r", subdir = "r-package")
+# https://github.com/ipeaGIT/r5r/blob/8101578e5178333cceb969ab5efd040755c48afa/r-package/R/detailed_itineraries.R#L134
 
 
+
+##################################################### POA
+library(data.table)
+library(r5r)
+library(dplyr)
 
 # build transport network
-path <- system.file("extdata/poa", package = "r5r")
-list.files(path)
-spo <- system.file("extdata/spo", package = "r5r")
-
-# r5r::download_r5(force_update = T)
-r5r::download_r5()
-r5r_core <- setup_r5(data_path = path, verbose = F)
+data_path <- system.file("extdata/poa", package = "r5r")
+r5r_core <- setup_r5(data_path = data_path)
 
 # load origin/destination points
-points <- read.csv(system.file("extdata/poa/poa_hexgrid.csv", package = "r5r"))
-points_sf <- sfheaders::sf_multipoint(points, x='lon', y='lat', multipoint_id = 'id')
+points <- read.csv(file.path(data_path, "poa_hexgrid.csv"))[1:5,]
 
-# remove files
-# file.remove( file.path(path, "network.dat") )
-# file.remove( file.path(.libPaths()[1], "r5r", "jar", "r5r_v4.9.0.jar") )
-# list.files(path)
-# list.files(file.path(.libPaths()[1], "r5r", "jar"))
+# load fare structure object
+fare_structure_path <- system.file(
+  "extdata/poa/fares/fares_poa.zip",
+  package = "r5r"
+)
+fare_structure <- read_fare_structure(fare_structure_path)
+
+departure_datetime <- as.POSIXct(
+  "13-05-2019 14:00:00",
+  format = "%d-%m-%Y %H:%M:%S"
+)
+
+pf_10 <- pareto_frontier(
+  r5r_core,
+  origins = points,
+  destinations = points,
+  mode = c("WALK", "TRANSIT"),
+  time_window = 30,
+  departure_datetime = departure_datetime,
+  fare_structure = fare_structure,
+  fare_cutoffs = c(4.5, 4.8, 9, 9.3, 9.6),
+  progress = T,
[…]

### 2022-12-22 r-package/R/r5r_sitrep.R
update NEWs

@@ -1,4 +1,4 @@
-#' Generate an r5r situation report
+#' Generate an r5r situation report to help debug errors
 #'
 #' @description
 #' The function reports a list with the following information:

### 2022-12-22 r-package/tests/testthat/test-set_monte_carlo_draws.R
vignette image

@@ -283,7 +283,7 @@ test_that("draws_per_minute arg works in pareto_frontier() - poa", {
   expect_false(identical(small_draws, big_draws))
 })
 
-# this test is currently failing
+# # this test is currently failing
 # test_that("draws_per_minute arg works in pareto_frontier() - spo", {
 #   basic_expr <- call(
 #     "pareto_frontier",

### 2023-01-27 r-package/man/roxygen/templates/lts_section.R
update documentation and fix unicode characters

@@ -21,10 +21,10 @@
 #' - **LTS 2**: Tolerable for the mainstream adult population. This includes
 #' streets where cyclists have dedicated lanes and only have to interact with
 #' traffic at formal crossing.
-#' - **LTS 3**: Tolerable for “enthused and confident” cyclists. This includes
+#' - **LTS 3**: Tolerable for "enthused and confident" cyclists. This includes
 #' streets which may involve close proximity to moderate- or high-speed
 #' vehicular traffic.
-#' - **LTS 4**: Tolerable only for “strong and fearless” cyclists. This
+#' - **LTS 4**: Tolerable only for "strong and fearless" cyclists. This
 #' includes streets where cyclists are required to mix with moderate- to
 #' high-speed vehicular traffic.
 #'

### 2023-01-27 r-package/tests/tests_rafa/test_rafa.R

@@ -653,7 +653,17 @@ system("R CMD Rd2pdf --title=Package gtfs2gps --output=./gtfs2gps/manual.pdf")
 # system("R CMD Rd2pdf gtfs2gps")
 
 
+# find unicode errors
+tools::showNonASCIIfile(file = './man/accessibility.Rd')
+tools::showNonASCIIfile(file = './man/roxygen/templates/mcraptor_algorithm_section.R')
+tools::showNonASCIIfile(file = './man/roxygen/templates/raptor_algorithm_section.R')
+tools::showNonASCIIfile(file = './vignettes/references.json')
 
+functions <- list.files(path = './R', all.files = T, recursive = T, full.names = T)
+lapply(X=functions, FUN = tools::showNonASCIIfile)
+
+docs <- list.files(path = './man', all.files = T, recursive = T, full.names = T)
+lapply(X=docs, FUN = tools::showNonASCIIfile)
 
 
 ### CMD Check ----------------
@@ -693,6 +703,16 @@ tictoc::toc()
 
 
 
+# submit to CRAN -----------------
+usethis::use_cran_comments('teste 2222, , asdadsad')
+
+urlchecker::url_check()
+devtools::check(remote = TRUE, manual = TRUE)
+devtools::check_win_release()
+rhub::check_for_cran(show_status = FALSE)
+
+
+devtools::submit_cran()
 
 
 # build binary -----------------

### 2023-09-08 r-package/R/accessibility.R
Closes #348

@@ -177,6 +177,14 @@ accessibility <- function(r5r_core,
   opportunities <- assign_opportunities(destinations, opportunities_colnames)
   mode_list <- assign_mode(mode, mode_egress)
   departure <- assign_departure(departure_datetime)
+
+  # cap trip duration with cutoffs
+  if(!is.null(cutoffs)){
+    max_trip_duration <- ifelse(max_trip_duration > max(cutoffs), max(cutoffs), max_trip_duration)
+
+    if(max_trip_duration < max(cutoffs)){stop("'max_trip_duration' cannot be shorter than 'max(cutoffs)'")}
+  }
+
   max_walk_time <- assign_max_street_time(
     max_walk_time,
     walk_speed,
@@ -201,6 +209,10 @@ accessibility <- function(r5r_core,
     max_walk_time,
     max_bike_time
   )
+
+
+
+
   decay_list <- assign_decay_function(decay_function, decay_value)
 
   set_time_window(r5r_core, time_window)

### 2023-12-05 r-package/R/download_r5.R
using tools::R_user_dir to store r5r jar file

@@ -48,8 +48,9 @@ download_r5 <- function(version = "7.0.0",
   destfile <- data.table::fifelse(
     temp_dir,
     file.path(tempdir(), filename),
-    file.path(system.file("jar", package = "r5r"), filename)
+    file.path(tools::R_user_dir("r5r", which = "cache"), filename)
   )
+
 
   # check if the file exists, and returns its path if it does. otherwise,
   # download it from IPEA's server - if there's no internet connection "fail
@@ -70,7 +71,7 @@ download_r5 <- function(version = "7.0.0",
   }
 
   # create dir
-  jar_dir <- system.file("jar", package = "r5r")
+  jar_dir <- tools::R_user_dir("r5r", which = "cache")
   if (!dir.exists(jar_dir)) dir.create(jar_dir)
 
   # download JAR

### 2023-12-05 r-package/R/r5r_sitrep.R

@@ -19,7 +19,7 @@
 r5r_sitrep <- function() {
   r5r_package_version <- utils::packageVersion("r5r")
 
-  jar_dir <- system.file("jar", package = "r5r")
+  jar_dir <- tools::R_user_dir("r5r", which = "cache")
   jar_dir_files <- list.files(jar_dir)
   jar_dir_files_full_names <- list.files(jar_dir, full.names = TRUE)
 

### 2023-12-05 r-package/R/setup_r5.R

@@ -116,7 +116,7 @@ setup_r5 <- function(data_path,
   jar_file <- data.table::fifelse(
     temp_dir,
     file.path(tempdir(), filename),
-    file.path(system.file("jar", package = "r5r"), filename)
+    file.path(tools::R_user_dir("r5r", which = "cache"), filename)
   )
 
   # If there isn't a JAR already, download it

### 2023-12-05 r-package/tests/tests_rafa/test_rafa.R

@@ -919,3 +919,4 @@ ggplot(data=df, aes(x=cost, y=time, label = modes)) +
         theme_classic()
 
 
+

### 2023-12-05 r-package/tests/testthat/test-write_fare_structure.R

@@ -29,3 +29,7 @@ test_that("written structure is identical to original", {
   written_struc <- read_fare_structure(tmpfile)
   expect_identical(struc, written_struc)
 })
+
+# clean cache
+jar_dir <- tools::R_user_dir("r5r", which = "cache")
+list.files(jar_dir, full.names = TRUE) |> file.remove()

### 2023-12-05 r-package/tests/testthat/test-write_fare_structure.R
uses unlink to clean chaced jar file

@@ -32,4 +32,5 @@ test_that("written structure is identical to original", {
 
 # clean cache
 jar_dir <- tools::R_user_dir("r5r", which = "cache")
-list.files(jar_dir, full.names = TRUE) |> file.remove()
+f <- list.files(jar_dir, full.names = TRUE)
+unlink(f, recursive = TRUE)

### 2024-01-13 r-package/R/download_r5.R
more robust cache of R5 jar

@@ -40,7 +40,9 @@ download_r5 <- function(version = "7.0.0",
   options(timeout = max(600, getOption("timeout")))
 
 
-  # download R5's jar -----------------------------------------------------
+  # download R5 jar -----------------------------------------------------
+
+  if (!dir.exists(r5r_env$cache_dir)) dir.create(r5r_env$cache_dir, recursive = TRUE)
 
   file_url <- fileurl_from_metadata(version)
   filename <- basename(file_url)
@@ -48,7 +50,7 @@ download_r5 <- function(version = "7.0.0",
   jar_file <- data.table::fifelse(
     temp_dir,
     file.path(tempdir(), filename),
-    file.path(tools::R_user_dir("r5r", which = "cache"), filename)
+    file.path( r5r_env$cache_dir , filename)
   )
 
 

### 2024-01-13 r-package/R/onLoad.R

@@ -1,0 +1,32 @@
+# initial message about ram memory
+.onAttach <- function(lib, pkg) {
+  packageStartupMessage(
+    "Please make sure you have already allocated ",
+    "some memory to Java by running:\n",
+    "  options(java.parameters = '-Xmx2G').\n",
+    "You should replace '2G' by the amount of memory you'll require. ",
+    "Currently, Java memory is set to ", getOption("java.parameters")
+    )
+  }
+
+# package global variables
+r5r_env <- new.env(parent = emptyenv()) # nocov start
+
+.onLoad <- function(lib, pkg) {
+
+  # JAR version
+  r5r_env$r5_jar_version <- "7.0.0"
+
+  # create dir to store R5 Jar
+  cache_d <- paste0('r5r/r5_jar_v', r5r_env$r5_jar_version)
+  r5r_env$cache_dir <- tools::R_user_dir(cache_d, which = 'cache')
+  if (!dir.exists(r5r_env$cache_dir)) dir.create(r5r_env$cache_dir, recursive = TRUE)
+  # gsub("\\\\", "/", r5r_env$cache_dir)
+
+  ## delete any JAR files from old releases
+  dir_above <- dirname(r5r_env$cache_dir)
+  all_cache <- list.files(dir_above, pattern = 'r5',full.names = TRUE)
+  old_cache <- all_cache[!grepl(r5r_env$r5_jar_version, all_cache)]
+  if(length(old_cache)>0){ unlink(old_cache, recursive = TRUE) }
+
+}

### 2024-01-13 r-package/R/r5r.R

@@ -83,18 +83,3 @@ if (getRversion() >= "2.15.1") {
     )
   )
 }
-
-.onAttach <- function(lib, pkg) {
-  packageStartupMessage(
-    "Please make sure you have already allocated ",
-    "some memory to Java by running:\n",
-    "  options(java.parameters = '-Xmx2G').\n",
-    "You should replace '2G' by the amount of memory you'll require. ",
-    "Currently, Java memory is set to ", getOption("java.parameters")
-  )
-
-  # create dir to store R5 Jar
-  jar_dir <- tools::R_user_dir("r5r", which = "cache")
-  if (!dir.exists(jar_dir)) dir.create(jar_dir, recursive = TRUE)
-
-}

### 2024-01-13 r-package/R/r5r_sitrep.R

@@ -19,13 +19,13 @@
 r5r_sitrep <- function() {
   r5r_package_version <- utils::packageVersion("r5r")
 
-  jar_dir <- tools::R_user_dir("r5r", which = "cache")
+  jar_dir <- r5r_env$cache_dir
   jar_dir_files <- list.files(jar_dir)
   jar_dir_files_full_names <- list.files(jar_dir, full.names = TRUE)
 
   r5r_jar <- jar_dir_files[grepl("r5r_\\d_\\d_\\d.*\\.jar", jar_dir_files)]
-  r5r_jar_version <- sub("\\.jar", "", sub("r5r_", "", r5r_jar))
-  r5r_jar_version <- gsub("_", "\\.", r5r_jar_version)
+
+  r5r_jar_version <- r5r_env$r5_jar_version
   r5r_jar_path <- jar_dir_files_full_names[
     grepl("\\/r5r_\\d_\\d_\\d*\\.jar", jar_dir_files_full_names)
   ]

### 2024-01-13 r-package/R/setup_r5.R

@@ -116,7 +116,7 @@ setup_r5 <- function(data_path,
   jar_file <- data.table::fifelse(
     temp_dir,
     file.path(tempdir(), filename),
-    file.path(tools::R_user_dir("r5r", which = "cache"), filename)
+    file.path( r5r_env$cache_dir, filename)
   )
 
   # If there isn't a JAR already, download it

### 2024-01-13 r-package/tests/testthat/test-write_fare_structure.R

@@ -31,6 +31,4 @@ test_that("written structure is identical to original", {
 })
 
 # clean cache
-jar_dir <- tools::R_user_dir("r5r", which = "cache")
-f <- list.files(jar_dir, full.names = TRUE)
-unlink(f, recursive = TRUE)
+r5r::r5r_cache(delete_file = 'all')

### 2024-06-24 r-package/man/roxygen/templates/draws_per_minute.R
improve documentation on behavior of time_window() without frequency.txt

@@ -2,4 +2,6 @@
 #'   perform per time window minute when calculating travel time matrices and
 #'   when estimating accessibility. Defaults to 5. This would mean 300 draws in
 #'   a 60-minute time window, for example. This parameter only affects the
-#'   results when the GTFS feeds contain a `frequencies.txt` table.
+#'   results when the GTFS feeds contain a `frequencies.txt` table. If the GTFS
+#'   feed does not have a frequency table, r5r still allow for multiple runs
+#'   over the set `time_window` but in a deterministic way.

### 2024-10-08 r-package/tests/tests_rafa/test_rafa.R
message o how to set number of cores

@@ -1,3 +1,8 @@
+#> Travel-time results are currently limited to 4 million destinations,
+#> and 16 million total origin-destination pairs. Path detail results are currently
+#> limited to 5000 destinations
+
+
 
 
 link <- 'https://ftp.ibge.gov.br/Trabalho_e_Rendimento/Pesquisa_Nacional_por_Amostra_de_Domicilios_continua/Trimestral/Microdados/2023/PNADC_032023.zip'

### 2025-06-11 r-package/R/arrival_travel_time_matrix.R
update docs od arrival_travel_time_matrix

@@ -2,12 +2,14 @@
 #' time of arrival
 #'
 #' Computation of travel time estimates between one or multiple origin
-#' destination pairs. This function considers a time of arrival set by the user.
-#' If you want to calculate travel times considering a departure time, have a
-#' look at the [travel_time_matrix()] function. This function is a wrapper around
-#' [expanded_travel_time_matrix()]. On one hand, this means this the output of
-#' this function has more columns (more info) compared the output of
-#' [travel_time_matrix()]. On the other hand, this function can be very memory
+#' destination pairs considering a time of arrival. This function considers a
+#' time of arrival set by the user. The function returns the travel time of the
+#' trip with the latest departure time that arrives before the arrival time set
+#' by the user. If you want to calculate travel times considering a departure
+#' time, have a' look at the [travel_time_matrix()] function. This function is a
+#' wrapper around [expanded_travel_time_matrix()]. On one hand, this means this
+#' the output of this function has more columns (more info) compared the output
+#' of [travel_time_matrix()]. On the other hand, this function can be very memory
 #' intensive if the user allows for really long max trip duration.
 #'
 #' @inheritParams expanded_travel_time_matrix

### 2025-06-23 r-package/R/fare_structure.R
text fixes to fare documentation

@@ -4,9 +4,7 @@
 #' calculated in [travel_time_matrix()], [expanded_travel_time_matrix()],
 #' [accessibility()] and [pareto_frontier()]. This fare structure can be
 #' manually edited and adjusted to the existing rules in your study area, as
-#' long as they stick to some basic premises. Please see fare structure
-#' vignette for more information on how the fare structure works:
-#' `vignette("fare_structure", package = "r5r")`.
+#' long as they stick to some basic premises. Please see the \link[vignette:fare_structure]{fare structure vignette} for more information.
 #'
 #' @template r5r_core
 #' @param base_fare A numeric. A base value used to populate the fare
@@ -341,7 +339,7 @@ assert_fare_structure <- function(fare_structure) {
     return(invisible(TRUE))
   } else {
     # This is an R5R fare structure object
-    
+
     element_names <- c(
       "max_discounted_transfers",
       "transfer_time_allowance",

### 2025-06-26 r-package/R/modify_osm_carspeeds.R
docs

@@ -1,7 +1,8 @@
-#' Create a transport network used for routing in R5 with modified car speeds
+#' Create a transport network used for routing in R5 with modified OSM car speeds
 #'
-#' The function builds a transport network after changing the car speeds from a
-#' `.pbf` file based on the new speeds indicated in a roadspeeds `.csv` file.
+#' The function builds a transport network with modified OSM car speeds. The new
+#' speed factors must me passed from a `.csv` file indication the new max speed
+#' of each OSM edge id.
 #'
 #' @param pbf_path Character. Path to the `.pbf` file of the OSM network.
 #' @param csv_path Character. Path to the CSV file witha a table specifying the

### 2025-07-15 java-r5rcore/src/org/ipea/r5r/Fares/FareStructureBuilder.java
rename fare structure first_leg to alight_leg and second_leg to board_leg

@@ -90,8 +90,8 @@ public class FareStructureBuilder {
             String[] legs = transfer.split("&");
 
             FarePerTransfer newTransfer = new FarePerTransfer();
-            newTransfer.setFirstLeg(legs[0]);
-            newTransfer.setSecondLeg(legs[1]);
+            newTransfer.setAlightLeg(legs[0]);
+            newTransfer.setBoardLeg(legs[1]);
             newTransfer.setFare(fare);
 
             faresPerTransfer.add(newTransfer);

### 2025-08-09 r-package/tests/tests_rafa/sample_data_lts_lines.R
more realistic example changing LTS with sf

@@ -1,0 +1,40 @@
+#' create lines with new LTS for sample data - Porto Alegre
+#'
+#' Here we simulate that local authorities would have build dedicated lanes
+#' along all secondary roads in the city.
+
+library(sf)
+library(dplyr)
+library(mapview)
+
+# data path
+data_path <- system.file("extdata/poa", package = "r5r")
+
+# path to OSM pbf
+pbf_path <- paste0(data_path, "/poa_osm.pbf")
+
+# read layer of lines from pbf
+allroads <- sf::st_read(
+  pbf_path,
+  layer = 'lines',
+  quiet = TRUE
+)
+
+# Filter only road types of interest
+roads <- allroads |>
+  select(osm_id, highway) |>
+  filter(highway %in% "secondary")
+
+head(roads)
+
+# Suppose your sf object is called `lines`
+lines_multi <- st_sf(
+  line_id = '1',
+  lts = 2L,
+  priority = 1L,
+  geometry = st_combine(roads)
+)
+
+mapview::mapview(lines_multi)
+
+saveRDS(lines_multi, 'poa_mls_lts.rds')

### 2025-08-10 r-package/tests/testthat/test-scenarios_LTS.R
changed lts settings to use lines instead of multilines

@@ -11,7 +11,7 @@ edge_lts <- data.frame(
   )
 
 # sf with new LTS
-lts_lines <- readRDS(file.path(data_path, "poa_mls_lts.rds"))
+lts_lines <- readRDS(file.path(data_path, "poa_ls_lts.rds"))
 
 
 meta_fun <- function(

### 2025-08-21 r-package/tests/tests_rafa/osm_ids.R
fix url in vignettes

@@ -8,7 +8,7 @@ RcppParallel::setThreadOptions (numThreads = 4L) # or desired number
 
 # data path where the .pbf file is located
 data_path <- system.file("extdata/poa", package = "r5r")
-pbf_path <- paste0(data_path,'/poa_osm.pbf')
+pbf_path <- file.path(data_path,'/poa_osm.pbf')
 
 # read layer of lines from pbf
 edges <- sf::st_read(
@@ -23,8 +23,13 @@ nodes <- sf::st_read(
   quiet = TRUE
 )
 
+edges$oneway <- ifelse(
+  edges$other_tags %like% "\"oneway\"=>\"yes\"",
+  "yes",
+  NA)
+
 # keep only osm ids
-edges <- edges[,c("osm_id","highway")]
+edges <- edges[,c("osm_id","highway", "oneway")]
 nodes <- nodes[,c("osm_id","highway")]
 
 
@@ -40,65 +45,75 @@ data.table::setDT(car_network)
 car_network[, from_id := as.numeric(from_id)]
 car_network[, to_id := as.numeric(to_id)]
 
-# create a unique edge id with the first and last node of the edge
-car_network <- car_network |>
-  group_by(way_id)  |>
-  mutate(unique_id_forward = paste0(min(from_id,to_id), '-', max(from_id,to_id)),
-         unique_id_backward = paste0(max(from_id,to_id), '-', min(from_id,to_id))
+# unique id, mas a edge_id ja eh uma unique id de cada edge
+car_network[,
+            unique_id := paste(way_id, geom_num, edge_id, from_id, to_id, sep = "_"),
+            by = .I
+            ]
+
+head(car_network)
+
+
+data.table::setDT(edges)
+car_network[edges, on = .(way_id=osm_id), oneway := i.oneway]
+car_network$oneway
+
+# 26786730 one way
+# 27126547 two ways
+car_network2 <- car_network[way_id %in% c(26786730, 27126547), ]
+car_network2 <- car_network
+
+# edges_sf <- sf::st_sf(edges)
+# mapview::mapview(
+#   filter(edges_sf, osm_id==27126547)
+# )
+
+one_way_net <- car_network2 |>
+  filter(oneway=="yes") |>
+  rename(osm_id = way_id) |>
+  group_by(osm_id, oneway)  |>
+  summarise(from_id = min(from_id),
+            from_lat = from_lat[which.min(from_id)],
+            from_lon = from_lon[which.min(from_id)],
+            to_id =  max(to_id),
[…]

### 2025-09-04 r-package/R/accessibility.R
fixed errors due to checking transit availability

@@ -188,8 +188,6 @@ accessibility <- function(r5r_network,
   # check inputs and set r5r options --------------------------------------
 
   checkmate::assert_class(r5r_network, "r5r_network")
-  r5r_network <- r5r_network@jcore
-
   origins <- assign_points_input(origins, "origins")
   destinations <- assign_points_input(destinations, "destinations")
   opportunities <- assign_opportunities(destinations, opportunities_colnames)
@@ -200,6 +198,8 @@ accessibility <- function(r5r_network,
   if (mode_list$transit_mode %like% 'TRANSIT|TRAM|SUBWAY|RAIL|BUS|CABLE_CAR|GONDOLA|FUNICULAR') {
     check_transit_availability_on_date(r5r_network, departure_date = departure$date)
   }
+
+  r5r_network <- r5r_network@jcore
 
   # cap trip duration with cutoffs
   set_cutoffs(r5r_network, cutoffs, decay_function)

### 2025-09-04 r-package/R/expanded_travel_time_matrix.R

@@ -137,24 +137,24 @@ expanded_travel_time_matrix <- function(r5r_network,
 
   # check inputs and set r5r options --------------------------------------
 
-  checkmate::assert_class(r5r_network, "r5r_network")
-  r5r_network <- r5r_network@jcore
-
   origins <- assign_points_input(origins, "origins")
   destinations <- assign_points_input(destinations, "destinations")
   mode_list <- assign_mode(mode, mode_egress)
   departure <- assign_departure(departure_datetime)
+
+  # check availability of transit services on the selected date
+  if (mode_list$transit_mode %like% 'TRANSIT|TRAM|SUBWAY|RAIL|BUS|CABLE_CAR|GONDOLA|FUNICULAR') {
+    check_transit_availability_on_date(r5r_network, departure_date = departure$date)
+  }
+
+  checkmate::assert_class(r5r_network, "r5r_network")
+  r5r_network <- r5r_network@jcore
 
   # in direct modes reverse origin/destination to take advantage of R5's One to Many algorithm
   data_path <- r5r_network$getDataPath()
   res <- reverse_if_direct_mode(origins, destinations, mode_list, data_path)
   origins <- res$origins
   destinations <- res$destinations
-
-  # check availability of transit services on the selected date
-  if (mode_list$transit_mode %like% 'TRANSIT|TRAM|SUBWAY|RAIL|BUS|CABLE_CAR|GONDOLA|FUNICULAR') {
-    check_transit_availability_on_date(r5r_network, departure_date = departure$date)
-  }
 
   max_walk_time <- assign_max_street_time(
     max_walk_time,

### 2025-09-04 r-package/R/travel_time_matrix.R

@@ -160,13 +160,18 @@ travel_time_matrix <- function(r5r_network,
 
   # check inputs and set r5r options --------------------------------------
 
-  checkmate::assert_class(r5r_network, "r5r_network")
-  r5r_network <- r5r_network@jcore
-
   origins <- assign_points_input(origins, "origins")
   destinations <- assign_points_input(destinations, "destinations")
   mode_list <- assign_mode(mode, mode_egress)
   departure <- assign_departure(departure_datetime)
+
+  # check availability of transit services on the selected date
+  if (mode_list$transit_mode %like% 'TRANSIT|TRAM|SUBWAY|RAIL|BUS|CABLE_CAR|GONDOLA|FUNICULAR') {
+    check_transit_availability_on_date(r5r_network, departure_date = departure$date)
+  }
+
+  checkmate::assert_class(r5r_network, "r5r_network")
+  r5r_network <- r5r_network@jcore
 
   # in direct modes reverse origin/destination to take advantage of R5's One to Many algorithm
   data_path <- r5r_network$getDataPath()
@@ -174,10 +179,6 @@ travel_time_matrix <- function(r5r_network,
   origins <- res$origins
   destinations <- res$destinations
 
-  # check availability of transit services on the selected date
-  if (mode_list$transit_mode %like% 'TRANSIT|TRAM|SUBWAY|RAIL|BUS|CABLE_CAR|GONDOLA|FUNICULAR') {
-    check_transit_availability_on_date(r5r_network, departure_date = departure$date)
-  }
 
   max_walk_time <- assign_max_street_time(
     max_walk_time,

### 2025-09-04 r-package/R/utils.R

@@ -45,7 +45,7 @@ check_transit_availability_on_date <- function(r5r_network,
                                                departure_date){ # nocov start
 
   # check services Available on the departure date
-  services <- check_transit_availability(r5r_network, dates = departure$date)
+  services <- check_transit_availability(r5r_network, dates = departure_date)
   services_available <- services$pct_active
 
   if (services_available == 0 | is.na(services_available)) {
